### 2021-03-19 README.Rmd
methods and results ☀️

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![Hay meadow from the Cantabrian Mountains](meadow.JPG)

# Seed germination ecology of species-rich mesic grasslands (*Arrhenatheretalia*)

This repository stores all the information related to the manuscript *Seed germination ecology of species-rich mesic grasslands (Arrhenatheretalia)*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

## Contents

This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.

The following materials are available in the folders of this repository:

* `data` Data files including the seed germination dataset, the species traits, and the phylogenetic tree.
* `doc` Files to create the manuscript using `Rmarkdown`.
* `results` Output of the `R` scripts, including figures and models.
* `src` Scripts in `R` language used to clean the raw data and perform the analyses of the manuscript.

## Citation

Please cite the dataset as `CITATION`

### 2021-03-19 doc/manuscript.Rmd
methods and results ☀️

---
title: "Seed germination ecology of species-rich mesic grasslands (*Arrhenatheretalia*)"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: journal-of-vegetation-science.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse)
source("../src/numbers.R")
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

Eduardo Fernández-Pascual^1^ ^+^, Elías Afif Khouri^1^, Joaquín Ascaso^2^, Beatriz Morais^1^, Ramón Reiné^2^, Madalena Vaz^3^

^1^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^2^ Departamento de Ciencias Agrarias y Medio Natural, Universidad de Zaragoza, Huesca, Spain; ^3^ Banco Português de Germoplasma Vegetal, Instituto Nacional de Investigação Agrária e Veterinária (INIAV), Braga, Portugal

^+^ Correspondence: Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, C/ Catedrático Rodrigo Uría, 33006 Oviedo/Uviéu, Spain. Email: efernandezpascual@gmail.com. Telephone: +34985104787.

# Abstract

# Keywords

*Arrhenatherion*, plant regeneration, seed germination, seed morphology

# Introduction

Iberian hay meadows [@RN2980] are seminatural grassland communities which developed in clearings of a formerly wooded landscape. They are species-rich mesophilous habitats, rarely fertilized, and occupy deep soils from clays to sands.  Mowing for haymaking thus is a predictable perturbation that occurs approximately at the same time every year, and which is expected to contribute to the high diversity of this habitats [@RN4887; @RN4888]. Hay meadow plants must establish, grow and reproduce in the time frame determined by this perturbation, and we can expect that seed morphology and germination timing will play a functional role in community assembly.

# Materials and methods

## Selection of study species

To select species for inclusion in our analysis, we used a dataset of 118 vegetation plots from three Iberian regions with well-maintained mesic grasslands: 43 plots from Northern Portugal (Trás-os-Montes), 25 from the Cantabrian Mountains of Spain (Parque Nacional Picos de Europa) and 50 from the Spanish Pyrenees (Aragón). The dataset contains a relatively high diversity of mesic grassland species, as the plots were situated along a major stress gradient related to drought [@RN2980]: two Spanish regions with a Temperate macroclimate, and Northern Portugal which is transitional between the Temperate and Mediterranean macroclimates [@RN4893]. Furthermore, grasslands from the Pyrenees are closest to the central European optimum of mesic grassland vegetation (*Arrhenatheretalia elatioris* Tx. 1931) while Portuguese grasslands are in suboptimal areas at the limit of the distribution [@RN4893].

Each plot was square (25-100 m^2 area) and was situated in the central part of a meadow avoiding the margins. The chosen meadows are maintained by traditional agricultural practices (mowing for hay-making once or twice per year). Vegetation in the plots was sampled in 2016-2017, at the peak of plant development, just before mowing. All vascular plant species in the plots were recorded and assigned an abundance value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%). All plant names in this manuscript follow the nomenclature of the Euro+Med plant database (https://www.emplantbase.org/home.html). Using all the plots, we calculated the cumulative abundance of each species in the entire area. We removed 208 species with cumulative abundance values below 2%, considering them transient species that might have been recorded by chance and do not represent the core mesic grasslands flora. We used the remaining 117 species as the core list to retrieve seed germination data for this article.

## Species collection and morphology

From the vegetation plots described above, we collected dispersal units (hereafter called seeds) during the dispersal seasons of 2016-2018. To describe seed morphology, we acquired images of 100-seed samples of each species using a flatbed scanner (Brother LC985) with a resolution of 200 dpi and a scanning area of 1024 x 1024 pixels [@RN2678]. We distributed the seeds on the scanner transparent glass, in a 10 x 10 grid. For each sample, and without moving the seeds, we repeated the scans with black and white backgrounds. In the case of the black background, we covered samples with a black box to avoid interference from environmental light. For the white background, we used the scanner cover. We digitized the obtained images and stored them in JPEG format (Joint Photographic Experts Group). We processed the scanned images using *ImageJ*, an open-source image processing program designed for scientific multidimensional images [@RN4708]. The program calculates several biometric parameters for each seed on the sample, among these, seed length and width. Using these values, we calculated seed shape as seed length divided by seed width. Additionally, we obtained seed mass for the study species from Kew's Seed Information Database [@RN2987].

## Seed germination data

To describe seed germination requirements, we germinated seeds of the different species following a protocol that accounts for the effect of germination temperature as a main physiological driver of germination phenology [@RN2249]. We used three germination treatments to determine the effect of temperatures representative for the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as the optimal summer temperature, and 30/20 ºC as sun-heated soil, e.g. soil exposed to sun after hay cutting. Additionally, we compared, at each of these temperature regimes, the germination of fresh seeds versus seeds subjected to a dormancy-breaking treatment. In the case of the Fabaceae, the treatment consisted in scarification by chipping the seed coat with a scalpel, to remove physical dormancy [@RN3214]. For the rest of the families, we used an incubation in gibberellic acid GA~3~ (0,0645 mM) in darkness during 24h, as a treatment to remove any potential physiological seed dormancy [@RN4976]. For each species and treatment, we sowed four Petri dishes with 25 seeds each. The germination substrate was 1% distilled water - agar. We sealed dishes with Parafilm to prevent desiccation. Trials took place in a germination chamber (KBW 400, Binder GmbH, Tuttlingen, Germany) with a 12/12 hour photoperiod (the light period corresponding to the higher temperature). Experiments lasted for four weeks, with germination scoring once per week. The germination criterion was 2 mm radicle emergence. After four weeks, we cut the seeds that failed to germinate and examined them under a magnifying glass. We classified them as normal when the embryo was visible and firm, empty when they lacked an embryo, and contaminated when they were moldy. We only considered normal seeds to calculate germination proportions and conduct subsequent analyses.

In addition to this experimental data, we retrieved seed germination records from ENSCOBASE, the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968]. The combined dataset, including our own experimental data and the records from ENSCOBASE, contained `r MSrecords` records (i.e. germination proportions for a given seed lot of a species, recorded in a set of laboratory experimental conditions). There were `r MSspecies` species in the dataset. Overall, `r MSseeds` seeds had been used in the experiments. The range of experimental germination temperatures (weighted average of the daily thermoperiod) which had been used in the experiments went from `r MSminT` to `r MSmaxT` ºC, with `r MSaltN` records of constant temperatures (i.e. experiments that used the same temperature during all their duration) and `r MSaltY` of alternating temperatures (i.e. experiments where different temperatures were applied during the day and the night, in diurnal cycles). Seeds had been exposed to light during some part of the diurnal cycle in `r MSlightY` records and kept in total darkness in `r MSlightN` records. Experiments had been performed with unstratified seeds (i.e. not subjected to a previous dormancy-breaking incubation) in `r MSstratN` records and with stratified seeds (i.e. subjected to previous incubation in dormancy-breaking conditions, including treatments of wet incubation under cold, warm and combinations of cold and warm conditions) in `r MSstratY` records. Finally, there were `r MSgaY` records where GA~3~ had been applied, and `r MSscarY` records where seeds had been scarified.

## Statistical analysis

To test the effect of the germination treatments on seed germination proportions, we performed a meta-analysis of the germination dataset by fitting binomial generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) using the R package *MCMCglmm* [@RN4755]. We fitted the analysis separately for the three major groups of mesic grassland plants: Poaceae, Fabaceae, and the minor families. To account for the effect of a shared phylogeny in species traits, models included as a random effect a reconstructed phylogenetic tree for the study species. To create the phylogeny we used the R package *V.PhyloMaker* [@RN4753] which contains an updated mega-tree of the seed plants based on Smith & Brown [-@RN4754]. We placed taxa absent from the mega-tree at the genus-level basal node. The phylogenetic tree is available in the data folder of the GitHub repository (see Data Availability Statement). Germination proportion was the response variable. Models had, as fixed effects, the experimental conditions (scarification, stratification, GA~3~, temperature, alternating temperature and light). Random effects included the phylogenetic tree, species identity, seed lot and source of the data. In all models, response variables were scaled so their contribution to the effect sizes could be compared. We used weakly informative priors in all models, with parameter-expanded priors for the random effects. Each model was run for 500,000 MCMC steps, with an initial burn-in phase of 50,000 and a thinning interval of 50 [@RN4756], resulting, on average, in 9,000 posterior distributions. From the resulting posterior distributions, we calculated mean parameter estimates and 95% Highest Posterior Density (HPD) and Credible Intervals (CI). We estimated the significance of model parameters by examining CIs, considering parameters with CIs overlapping with zero as non-significant. To estimate the phylogenetic signal of seed germination over all variables, we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by De Villemereuil et al. [-@RN4756]. When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in trait evolution that does not fully follow a Brownian motion model [@RN4757]. Detailed results of the MCMCglmms are available in the results folder of the GitHub repository (see Data Availability Statement).

To visualize the main patterns in the seed germination spectrum of mesic grassland plants, we did a Principal Component Analysis (PCA) as implemented in the package *FactoMineR* [@RN3166]. We performed the PCA ordination at the species level, i.e. calculating a series of continuous traits for each species. We transformed the final germination proportions to create a continuous variable for the germination cues (i.e. stratification, scarification, temperature, alternating temperatures and light). To do so, for each cue and species, we calculated a weighted average of the cue levels (in the case of temperature, cue levels were the temperature treatments; for the other cues the levels were 0 = absence and 1 = presence), weighting by the germination proportion at each level. This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of the response to the germination cues when visualized across the whole dataset; it must be stressed that this stage of the analysis serves only for visualization and not for hypothesis testing. We also included seed mass in the ordination. We left GA~3~ out of the PCA because its ecological interpretation is subordinated to stratification (as both cues break physiological seed dormancy). We also left seed length, width and shape out because these values were not available for all species.

# Results

## Seed morphology

The dispersal unit in the dominant grass family (Poaceae) is generally the floret (**Fig. 1a**), a composite structure made up of the caryopsis (a dry, indehiscent and monospermic fruit) and several surrounding bracts (i.e. modified leaves): the lemma, the palea and, in many species, two glumes. This structure is often accompanied by awns or hairs that aid in dispersal (**Fig. 1b**). The mass of these dispersal units in grassland grasses ranges from `r MSpoaMin` to `r MSpoaMax` mg, and they tend to be longer (`r MSpoaLength` mm on average) than they are wide (`r MSpoaWidth` mm). In the subordinate family of the legumes (Fabaceae), the dispersal unit of most species is the seed itself (**Fig. 1c**). Fabaceae generally are hard-seeded, i.e. they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e. physical seed dormancy). In some Fabaceae, dispersal units are more complex, including indehiscent monospermic fruits (e.g. *Onobrychis*, **Fig. 1d**) or indehiscent monospermic fruit fragments, i.e. loments (e.g. *Ornithopus*, **Fig. 1e**). In mesic grasslands, the dispersal units of the Fabaceae are generally heavier than those of the Poaceae, ranging from `r MSlegMin` to `r MSlegMax` mg; and they are also rounder (`r MSlegLength` mm long vs. `r MSlegWidth` mm wide). In many of the most frequent minor families, such as the Asteraceae, the Ranunculaceae, the Lamiaceae and the Dipsacaceae, the dispersal unit is the achene, i.e. a dry, indehiscent and monospermic fruit. The achenes of the grassland Asteraceae tend to be cypselas, wearing a pappus that aids in dispersal by wind (**Fig. 1f**). In the Ranunculaceae, the achenes are usually flattened and sometimes beaked (**Fig. 1g**). The achenes of the Lamiaceae are hardened nutlets (**Fig. 1h**). In the genus *Knautia* (Dipsacaceae), the achene is covered by hairs (**Fig. 1i**). In the Cyperaceae, the dispersal unit is an achene surrounded by a perigynium which aids in dispersal by water (**Fig. 1j**). In the genus *Sanguisorba* (Rosaceae), the dispersal unit is the urn-shaped receptacle containing one to three achenes (**Fig. 1k**). The dispersal unit of the Apiaceae is the mericarp (**Fig. 1l**), an indehiscent monospermic fragment of the fruit. In some other minor families the dispersal unit is the seed itself, such as in the Caryophyllaceae (**Fig. 1m**) or the Plantaginaceae (**Fig. 1n**). The hemiparasitic species of *Rhinanthus* (Orobanchaceae) disperse by winged seeds (**Fig. 1ñ**). When comparing the morphology of all species (**Fig. 2**), the Poaceae appear as having relatively lighter and more elongated dispersal units, in contrast with the Fabaceae, which tend to have rounder and heavier seeds.

## Germination responses

When considering the full set of mesic grassland species, all of the six studied germination cues had a significant effect on final germination proportions (**Fig. 3**). The germination of grassland seeds was increased by scarification, stratification, GA~3~, alternating temperatures, and light. Temperature had a negative effect, indicating a trend towards higher germination at lower temperatures. Averaging the whole dataset, the highest germination proportions were achieved at 20ºC. Between 0ºC and 20ºC, germination proportions increased steadily with increasing temperatures. Above 20ºC, germination declined more sharply, and grassland seeds rarely germinated at 30ºC.

Some differences became apparent when dividing the dataset in the major floristic groups that compose mesic grassland vegetation. In the dominant group of the Poaceae (**Fig. 3**), no effect was found for stratification, GA~3~ or temperature. Stratification had a negative effect on germination. The major drivers of Poaceae germination appeared to be alternating temperatures and light, with both having a positive effect. In the subordinate group of the Fabaceae (**Fig. 3**), the largest positive effect on germination was produced by scarification, with no effect of stratification, GA~3~ or light. Temperature had a negative effect, and in fact Fabaceae species had higher germination proportion at temperatures under 20ºC than Poaceae species did. Fabaceae seed germination also responded negatively to alternating temperatures. Finally, for the group of the minor families (**Fig. 3**), the main cues having a positive effect on germination were stratification, GA~3~. alternating temperatures and light. These species did not respond to scarification nor to temperature.

To describe the effects of the random factors, we will refer only to the full model that included all species in the dataset. The strongest effect was that of the phylogeny (mean = `r MSphylo`, CI = `r MSphylol` - `r MSphyloh`), followed by the source of the data (mean = `r MSsource`, CI = `r MSsourcel` - `r MSsourceh`), the seed lot (mean = `r MSpopulation`, CI = `r MSpopulationl` - `r MSpopulationh`), and species identity (mean = `r MSspecies`, CI = `r MSspeciesl` - `r MSspeciesh`). The phylogenetic signal in the germination responses was relatively high (lambda = `r MSlambda`, CI = `r MSlambdal` - `r MSlambdah`).

## Seed germination spectrum

The first PCA axis explained `r round(pca$eig[1, 2], 0)`% of the variation (**Fig. 4**). The variables with the largest contribution to this first axis were the germination responses to scarification and alternating temperatures, and seed mass. This horizontal axis separated (i, left) species of the Poaceae and the minor families which responded to alternating temperatures and had smaller seeds from (ii, right) species of the Fabaceae which responded to scarification and had heavier seeds. Axis 2 explained `r round(pca$eig[2, 2], 0)`% of the variability. The main contributing variables were the germination responses to light and stratification. This axis separated (iii, bottom) species from all groups that germinated better in light from (iv, top) species of the minor families which needed stratification to germinate and had higher germination in the dark.

# Discussion

# Acknowledgements

E.F.P. received financial support from the Government of Asturias and the FP7 – Marie Curie - COFUND programme of the European Commission (Grant ‘Clarín’ ACB17-19). This research was supported by project Interreg SUDOE “SOS PRADERAS” SOE1/P5/E037 (https://www.sospraderas.eu/en/).

# Author contributions

E.F.P. conceived the study, performed the analyses and wrote the manuscript. All authors contributed data, revised the manuscript, and approved the final version.

# Data availability

The original data, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/meadows. Upon publication, a version of record of the repository will be deposited in Zenodo.

# References

::: {#refs}
:::

# Figures

```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Diversity of dispersal units in mesic grasslands: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* subsp. *despectus* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadencis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (ñ) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
knitr::include_graphics("../results/figures/seeds.png")
```

```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Seed size and shape in mesic grasslands. Kernel probability densities (violin plots) for seed mass, seed length and seed shape (length / width).  The three horizontal lines within the probability densities represent the first quartile, the median and the third quartile of the data."}
knitr::include_graphics("../results/figures/violins.png")
```

```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on seed germination of mesic grassland species. Dots indicate the posterior mean of the effect size, and whiskers its 95 % credible interval. The line of zero effect is shown. When the credible intervals overlap with the zero-effect line, the effect is not significant. The figure shows the results of a model including all species, and of models including only the three main floristic groups of mesic grasslands"}
knitr::include_graphics("../results/figures/mcmc.png")
```

```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Principal Component Analysis ordination of the seed germination spectrum of mesic grassland species. Each symbol is a species, different symbols indicate the three major floristic groups of mesic grasslands Labels indicate the contribution of the variables to the axes."}
knitr::include_graphics("../results/figures/pca.png")
```

### 2021-03-19 src/cleaning.R

@@ -1,0 +1,357 @@
+library(tidyverse); library(openxlsx)
+
+# Clean species data
+
+read.xlsx("../#data/plots/data/sospraderas/Inventarios SOS PRADERAS privacidad.xlsx", sheet = 1) %>% # These are the relevés in the DB
+  merge(read.xlsx("../#data/plots/data/sospraderas/Inventarios SOS PRADERAS privacidad.xlsx", sheet = 2), # These are the species names in the DB
+        by.x = "codigo_especie", by.y = "ID") %>%
+  merge(read.csv("../#data/plots/data/sospraderas/Especies revisadas 20210108.csv")) %>% # Revised spp names
+  mutate(valor = fct_recode(valor, # Braun-Blanquet to cover scale
+                            "87.5" = "5",
+                            "62.5" = "4",
+                            "37.5" = "3",
+                            "15" = "2",
+                            "5" = "1",
+                            "1" = "+", 
+                            "0.1" = "r"),
+         valor = as.character(valor), 
+         valor = as.numeric(valor)) %>%
+  rename(Plot = Cod_Sitio,
+         Cover = valor) %>%
+  dplyr::select(Plot, Species, Cover) %>% 
+  arrange(Plot) %>%
+  group_by(Plot, Species) %>%
+  summarise(Cover = max(Cover)) -> # remove species that may have been merged by taxonomy
+  species # clean species tibble
+
+# Clean header data
+
+read.xlsx("../#data/plots/data/sospraderas/Inventarios SOS PRADERAS privacidad.xlsx", sheet = 3) %>% # This is the header data in the database
+  merge(read.csv("../#data/plots/data/sospraderas/Coordenadas ETRS.csv")) %>%
+  mutate(Date = convertToDate(Data)) %>%
+  filter(Date > 2001-01-01 & !is.na(Latitude)) %>% # Only 21st Century, with coordinates
+  mutate(`SIC.(LIC)` = as.character(`SIC.(LIC)`), 
+         Region = ifelse(`SIC.(LIC)` %in% "PNPE", "Cantabrian Mountains", Pais),
+         Region = fct_recode(Region, 
+                             "French Pyrenees" = "França", 
+                             "Spanish Pyrenees" = "Espanha",
+                             "Northern Portugal" = "Portugal")) %>%
[…]

### 2021-03-19 src/figures.R

@@ -1,0 +1,183 @@
+library(tidyverse)
+
+# Figure 1 - Seed morphology
+
+## Plot figure
+
+read.csv("data/traits.csv") %>%
+  select(Taxon, Family,
+         Seed.mass, Length, Width) %>%
+  na.omit %>%
+  mutate(Family = ifelse(Family %in% c("Leguminosae", "Poaceae"), Family, "Others"),
+         Family = ifelse(Family %in% "Leguminosae", "Fabaceae", Family),
+         Shape = Length / Width) %>%
+  select(-Width) %>%
+  gather(Trait, Value, Seed.mass:Shape) %>%
+  mutate(Family = fct_relevel(Family, c("Poaceae", "Fabaceae", "Others"))) %>%
+  mutate(Trait = fct_relevel(Trait, c("Seed.mass", "Length", "Shape"))) %>%
+  mutate(Trait = fct_recode(Trait, 
+                            "Seed mass (mg)" = "Seed.mass",
+                            "Seed length (mm)" = "Length",
+                            "Seed shape (length / width)" = "Shape")) %>%
+  ggplot(aes(Family, Value, color = Family, fill = Family)) + 
+  geom_violin(alpha = 0.5, draw_quantiles = c(0.25, 0.5, 0.75)) +
+  facet_wrap( ~ Trait, strip.position = "left", scales = "free_y") +
+  scale_y_continuous(labels = scales::number_format(accuracy = 1)) +
+  ggthemes::theme_tufte() +
+  theme(panel.background = element_rect(color = "grey96", fill = "grey96"),
+        strip.text = element_text(size = 12, color = "black"),
+        axis.text = element_text(size = 11, color = "black"),
+        axis.title = element_blank(),
+        legend.position = "none", strip.placement = "outside") +
+  geom_jitter(shape = 16, position = position_jitter(0.05), alpha = 0.6) +
+  scale_color_manual(values = c("yellowgreen",  "gold", "darkorchid")) +
+  scale_fill_manual(values = c("yellowgreen",  "gold", "darkorchid")) -> f1
+
+## Export
+
+ggsave(f1, file = "results/figures/violins.png", 
+       path = NULL, scale = 1, width = 180, height = 65, units = "mm", dpi = 600)
+
+# Figure 2 - MCMC
+
+## Load models
+
+load("results/models/allspp.RData")
+load("results/models/leguminosae.RData")
+load("results/models/poaceae.RData")
+load("results/models/rest.RData")
+
+## Bind effect sizes data frame
+
[…]

### 2021-03-19 src/models.R

@@ -1,0 +1,99 @@
+library(tidyverse)
+
+# Read tree
+
+phangorn::nnls.tree(cophenetic(ape::read.tree("data/meadowstree.tree")), 
+                    ape::read.tree("data/meadowstree.tree"), rooted = TRUE) -> 
+  nnls_orig
+
+# Set number of iterations
+
+nite = 500000
+nthi = 50
+nbur = 50000
+
+# Set priors from https://doi.org/10.1017/S0960258517000332
+
+list(R = list(V = 1, nu = 50), 
+     G = list(G1 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 500), 
+              G2 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 500),
+              G3 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 500),
+              G4 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 500)
+     )) -> 
+  priors
+
+# Read data
+
+read.csv("data/germination.csv") %>%
+  merge(read.csv("data/traits.csv"), by = "Taxon") %>%
+  mutate(ID = gsub(" ", "_", Taxon), animal = ID) %>% 
+  select(ID, animal, Family,
+         Reference:Population, Scarification:GA3,
+         Tmean:Light, Germinated:Germinable) -> 
+  data
+
+# All species
+
+MCMCglmm::MCMCglmm(cbind(Germinated, Germinable - Germinated) ~ scale(Scarification) + 
+                     scale(Stratification) + scale(GA3) + scale(Tmean) + 
+                     scale(Alternating) + scale(Light), 
+                   random = ~ animal + ID + Reference + Population:Reference, 
+                   family = "multinomial2", pedigree = nnls_orig, prior = priors, data = data,
+                   nitt = nite, thin = nthi, burnin = nbur, 
+                   verbose = FALSE, saveX = FALSE, saveZ = FALSE, saveXL = FALSE, 
+                   pr = FALSE, pl = FALSE) -> allspp
+
+summary(allspp)
+save(allspp, file = "results/models/allspp.Rdata")
+rm(allspp)
+
+# Poaceae
+
+data %>% filter(Family == "Poaceae") -> poaceaedf
+
+MCMCglmm::MCMCglmm(cbind(Germinated, Germinable - Germinated) ~ scale(Scarification) + 
+                     scale(Stratification) + scale(GA3) + scale(Tmean) + 
+                     scale(Alternating) + scale(Light), 
[…]

### 2021-03-19 src/numbers.R

@@ -1,0 +1,90 @@
+# Read data
+
+read.csv("../data/germination.csv") -> germination
+read.csv("../data/traits.csv") -> traits
+load("../results/models/allspp.RData")
+
+# Calculate numbers
+
+germination %>% pull(Taxon) %>% length -> MSrecords
+germination %>% pull(Taxon) %>% unique %>% length -> MSspecies
+sum(germination$Number_seeds) -> MSseeds
+min(germination$Tmean) -> MSminT
+max(germination$Tmean) -> MSmaxT
+germination %>% filter(Alternating == 1) %>% pull(Taxon) %>% length -> MSaltY
+germination %>% filter(Alternating == 0) %>% pull(Taxon) %>% length -> MSaltN
+germination %>% filter(Light == 1) %>% pull(Taxon) %>% length -> MSlightY
+germination %>% filter(Light == 0) %>% pull(Taxon) %>% length -> MSlightN
+germination %>% filter(Stratification == 0) %>% pull(Taxon) %>% length -> MSstratN
+germination %>% filter(Stratification == 1) %>% pull(Taxon) %>% length -> MSstratY
+germination %>% filter(Scarification == 0) %>% pull(Taxon) %>% length -> MSscarN
+germination %>% filter(Scarification == 1) %>% pull(Taxon) %>% length -> MSscarY
+germination %>% filter(GA3 == 0) %>% pull(Taxon) %>% length -> MSgaN
+germination %>% filter(GA3 == 1) %>% pull(Taxon) %>% length -> MSgaY
+
+# Seed morpho
+
+traits %>%
+  group_by(Family) %>%
+  summarise(Min = min(Seed.mass, na.rm = TRUE), 
+            Max = max(Seed.mass, na.rm = TRUE),
+            Length = mean(Length, na.rm = TRUE), 
+            Width = mean(Width, na.rm = TRUE)) -> morpho
+
+morpho %>% filter(Family == "Poaceae") %>% pull(Min) %>% round(1) -> MSpoaMin
+morpho %>% filter(Family == "Poaceae") %>% pull(Max) %>% round(1) -> MSpoaMax
+morpho %>% filter(Family == "Poaceae") %>% pull(Length) %>% round(1) -> MSpoaLength
+morpho %>% filter(Family == "Poaceae") %>% pull(Width) %>% round(1) -> MSpoaWidth
[…]

### 2021-04-30 README.Rmd
intro 🌧️

@@
 
-# Seed germination ecology of species-rich mesic grasslands (*Arrhenatheretalia*)
+# Seed germination ecology of European mesic grasslands
 
-This repository stores all the information related to the manuscript *Seed germination ecology of species-rich mesic grasslands (Arrhenatheretalia)*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
+This repository stores all the information related to the manuscript *Seed germination ecology of European mesic grasslands*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
 

### 2021-04-30 doc/manuscript.Rmd
intro 🌧️

@@
 ---
-title: "Seed germination ecology of species-rich mesic grasslands (*Arrhenatheretalia*)"
+title: "Seed germination ecology of European mesic grasslands"
 output: 
@@
 
-Eduardo Fernández-Pascual^1^ ^+^, Elías Afif Khouri^1^, Joaquín Ascaso^2^, Beatriz Morais^1^, Ramón Reiné^2^, Madalena Vaz^3^
+Eduardo Fernández-Pascual^1^ ^+^, Madalena Vaz^2^, Beatriz Morais^1^, Ramón Reiné^3^, Joaquín Ascaso^3^, Elías Afif Khouri^1^, Angelino Carta^4^
 
-^1^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^2^ Departamento de Ciencias Agrarias y Medio Natural, Universidad de Zaragoza, Huesca, Spain; ^3^ Banco Português de Germoplasma Vegetal, Instituto Nacional de Investigação Agrária e Veterinária (INIAV), Braga, Portugal
+^1^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^2^ Banco Português de Germoplasma Vegetal, Instituto Nacional de Investigação Agrária e Veterinária (INIAV), Braga, Portugal; ^3^ Departamento de Ciencias Agrarias y Medio Natural, Universidad de Zaragoza, Huesca, Spain; ^4^ Dipartimento di Biologia, Botany Unit, University of Pisa, Pisa, Italy
 
@@
 
-*Arrhenatherion*, plant regeneration, seed germination, seed morphology
+*Arrhenatheretalia*, plant regeneration, seed germination, seed morphology, species-rich meadows, hay meadows, mesic pastures
 
@@
 
-Iberian hay meadows [@RN2980] are seminatural grassland communities which developed in clearings of a formerly wooded landscape.
-They are species-rich mesophilous habitats, rarely fertilized, and occupy deep soils from clays to sands.
-Mowing for haymaking thus is a predictable perturbation that occurs approximately at the same time every year, and which is expected to contribute to the high diversity of this habitats [@RN4887; @RN4888].
-Hay meadow plants must establish, grow and reproduce in the time frame determined by this perturbation, and we can expect that seed morphology and germination timing will play a functional role in community assembly.
+European mesic grasslands are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698].
+These plant communities developed by anthropogenic intervention in clearings of a formerly wooded landscape and are maintained by agricultural practices, contributing to the diversity and conservation value of European landscapes [@RN4996].
+Traditional phytosociological classification of mesic grasslands put emphasis on the separation between mesic meadows used for hay making versus mesic pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893].
+Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic grasslands in large parts of Europe [@RN4997].
+For this reason the European Habitats Directive (92/43/EEC) has included in its lists of habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.
+
+One of the reasons for the high conservation interest of traditional mesic grasslands is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed materials to be used in ecological restoration, rewilding and urban greening [@RN4894].
+However, a lack of knowledge about species germination traits has already been identified as a bottleneck [@RN2241] hampering the development of a competitive native seed industry [@RN4665] in Europe.
+Understanding seed germination is part of the practical scientific framework needed to tackle large-scale ecological restoration challenges [@RN3507] and to maintain ex situ collections of plant genetic resources [@RN3763].
+When such a knowledge framework is available, ambitious regional schemes of seed-based landscape restoration can be designed [@RN4666].
+
+In the case of mesic grasslands, a concerted investigation of their seed germination ecology is missing, despite their conservation and restoration importance.
+Most of the available information is scattered in single-species studies [@RN2236] or using a limited range of germination conditions [@RN4894].
+In this article, we present for the first time a synthesis of germination ecology in European mesic grasslands.
+To do so, we combined a newly-generated dataset with germination records from ENSCOBASE, the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968].
+To the resulting dataset containing `r MSrecords` germination records from `r MScountries` European countries we applied Bayesian meta-analysis [@RN5000] to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light.
+In addition, we also analyzed seed morphology (mass and shape) and the differences between the main functional groups of mesic grasslands: grasses (Poaceae), legumes (Fabaceae) and the other families.
 
@@
 
-To select species for inclusion in our analysis, we used a dataset of 118 vegetation plots from three Iberian regions with well-maintained mesic grasslands: 43 plots from Northern Portugal (Trás-os-Montes), 25 from the Cantabrian Mountains of Spain (Parque Nacional Picos de Europa) and 50 from the Spanish Pyrenees (Aragón).
-The dataset contains a relatively high diversity of mesic grassland species, as the plots were situated along a major stress gradient related to drought [@RN2980]: two Spanish regions with a Temperate macroclimate, and Northern Portugal which is transitional between the Temperate and Mediterranean macroclimates [@RN4893].
+To create a species list of representative mesic grassland species for inclusion in our analysis, we used a dataset of 118 vegetation relevés from three Western European regions with well-maintained mesic grasslands: 43 from Northern Portugal (Trás-os-Montes), 25 from the Cantabrian Mountains of Spain (Parque Nacional Picos de Europa) and 50 from the Pyrenees.
+This dataset represents a sample of mesic grassland diversity, as the plots were situated along a major stress gradient related to drought [@RN2980]: two regions with a Temperate macroclimate, and Northern Portugal which is transitional between the Temperate and Mediterranean macroclimates [@RN4893].
 Furthermore, grasslands from the Pyrenees are closest to the central European optimum of mesic grassland vegetation (*Arrhenatheretalia elatioris* Tx.
 1931) while Portuguese grasslands are in suboptimal areas at the limit of the distribution [@RN4893].
+Finally, traditional management of mesic grasslands is relatively well preserved in this area within the European context [@RN4998; @RN4999].
 
-Each plot was square (25-100 m^2 area) and was situated in the central part of a meadow avoiding the margins.
-The chosen meadows are maintained by traditional agricultural practices (mowing for hay-making once or twice per year).
-Vegetation in the plots was sampled in 2016-2017, at the peak of plant development, just before mowing.
-All vascular plant species in the plots were recorded and assigned an abundance value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%).
+Each relevé had been conducted in a square area (25-100 m^2^ area) situated in the central part of a meadow avoiding the margins.
+The chosen meadows are maintained by traditional agricultural practices (mowing for hay-making once or twice per year and light grazing).
+Vegetation in the plots had been sampled in 2016-2017, at the peak of plant development, just before mowing.
+All vascular plant species in the plots had been recorded and assigned an abundance value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%).
 All plant names in this manuscript follow the nomenclature of the Euro+Med plant database (https://www.emplantbase.org/home.html).
-Using all the plots, we calculated the cumulative abundance of each species in the entire area.
+Using all the relevés, we calculated the cumulative abundance of each species in the entire area.
 We removed 208 species with cumulative abundance values below 2%, considering them transient species that might have been recorded by chance and do not represent the core mesic grasslands flora.
@@
 
-## Species collection and morphology
+## Seed collection and morphology
 
-From the vegetation plots described above, we collected dispersal units (hereafter called seeds) during the dispersal seasons of 2016-2018.
+From the vegetation plots described above, we collected dispersal units (hereafter called seeds) during the dispersal seasons of 2016, 2017 and 2018.
 To describe seed morphology, we acquired images of 100-seed samples of each species using a flatbed scanner (Brother LC985) with a resolution of 200 dpi and a scanning area of 1024 x 1024 pixels [@RN2678].
@@
 
-To describe seed germination requirements, we germinated seeds of the different species following a protocol that accounts for the effect of germination temperature as a main physiological driver of germination phenology [@RN2249].
-We used three germination treatments to determine the effect of temperatures representative for the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as the optimal summer temperature, and 30/20 ºC as sun-heated soil, e.g.
+We germinated seeds of the different species using three germination treatments to determine the effect of temperatures representative for the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as the optimal summer temperature, and 30/20 ºC as sun-heated soil, e.g.
 soil exposed to sun after hay cutting.
@@
 In the case of the Fabaceae, the treatment consisted in scarification by chipping the seed coat with a scalpel, to remove physical dormancy [@RN3214].
-For the rest of the families, we used an incubation in gibberellic acid GA~3~ (0,0645 mM) in darkness during 24h, as a treatment to remove any potential physiological seed dormancy [@RN4976].
+For the rest of the families, we used gibberellic acid GA~3~ (0,0645 mM) in darkness during 24h, as a treatment to remove potential physiological seed dormancy [@RN4976].
 For each species and treatment, we sowed four Petri dishes with 25 seeds each.
@@
 In addition to this experimental data, we retrieved seed germination records from ENSCOBASE, the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968].
-The combined dataset, including our own experimental data and the records from ENSCOBASE, contained `r MSrecords` records (i.e.
-germination proportions for a given seed lot of a species, recorded in a set of laboratory experimental conditions).
-There were `r MSspecies` species in the dataset.
-Overall, `r MSseeds` seeds had been used in the experiments.
-The range of experimental germination temperatures (weighted average of the daily thermoperiod) which had been used in the experiments went from `r MSminT` to `r MSmaxT` ºC, with `r MSaltN` records of constant temperatures (i.e.
+The combined dataset, including our own experimental data and the records from ENSCOBASE and *SylvanSeeds*, contained `r MSrecords` germination records (i.e.
+germination proportions for a given seed lot of a species, recorded in a set of laboratory experimental conditions) from `r MScountries` European countries.
+There were `r MSspeciesN` species in the dataset.
+Overall, `r MSseedsN` seeds had been used in the experiments.
+The range of experimental germination temperatures (weighted average of the daily thermoperiod) which had been used in the experiments spans from `r MSminT` to `r MSmaxT` ºC, with `r MSaltN` records of constant temperatures (i.e.
 experiments that used the same temperature during all their duration) and `r MSaltY` of alternating temperatures (i.e.
 experiments where different temperatures were applied during the day and the night, in diurnal cycles).
-Seeds had been exposed to light during some part of the diurnal cycle in `r MSlightY` records and kept in total darkness in `r MSlightN` records.
+Seeds had been exposed to light during some part of the diurnal cycle in `r MSlightY` records or kept in total darkness in `r MSlightN` records.
 Experiments had been performed with unstratified seeds (i.e.
@@
 a dry, indehiscent and monospermic fruit.
-The achenes of the grassland Asteraceae tend to be cypselas, wearing a pappus that aids in dispersal by wind (**Fig.
+In the case of the Asteraceae, these achenes tend to be cypselas, wearing a pappus that aids in dispersal by wind (**Fig.
 1f**).
@@
 The hemiparasitic species of *Rhinanthus* (Orobanchaceae) disperse by winged seeds (**Fig.
-1ñ**).
+1o**).
 When comparing the morphology of all species (**Fig.
@@
 3**).
-The germination of grassland seeds was increased by scarification, stratification, GA~3~, alternating temperatures, and light.
-Temperature had a negative effect, indicating a trend towards higher germination at lower temperatures.
-Averaging the whole dataset, the highest germination proportions were achieved at 20ºC.
-Between 0ºC and 20ºC, germination proportions increased steadily with increasing temperatures.
-Above 20ºC, germination declined more sharply, and grassland seeds rarely germinated at 30ºC.
+The germination of grassland seeds was positively associated with scarification, stratification, GA~3~, alternating temperatures, and light.
+Average temperature had a negative effect, indicating a trend towards higher germination at lower temperatures.
+Averaging the whole dataset, the highest germination proportions were achieved at 20 ºC.
+Between 0 ºC and 20 ºC, germination proportions increased steadily with increasing temperatures.
+Above 20 ºC, germination declined more sharply, and grassland seeds rarely germinated at 30 ºC.
 
-Some differences became apparent when dividing the dataset in the major floristic groups that compose mesic grassland vegetation.
+Some differences became apparent when dividing the dataset in the major floristic groups that compose mesic grassland vegetation (i.e.
+(i) the Poaceae as dominant and structural species, (ii) the Fabaceae as the second largest group, and (iii) the rest of the families).
 In the dominant group of the Poaceae (**Fig.
@@
 The major drivers of Poaceae germination appeared to be alternating temperatures and light, with both having a positive effect.
-In the subordinate group of the Fabaceae (**Fig.
+In the group of the Fabaceae (**Fig.
 3**), the largest positive effect on germination was produced by scarification, with no effect of stratification, GA~3~ or light.
-Temperature had a negative effect, and in fact Fabaceae species had higher germination proportion at temperatures under 20ºC than Poaceae species did.
+Temperature had a negative effect, and in fact Fabaceae species had higher germination proportion at temperatures under 20 ºC.
 Fabaceae seed germination also responded negatively to alternating temperatures.
-Finally, for the group of the minor families (**Fig.
+Finally, the remaining families (**Fig.
 3**), the main cues having a positive effect on germination were stratification, GA~3~.
 alternating temperatures and light.
-These species did not respond to scarification nor to temperature.
+These species did not respond to scarification nor to average temperature.
 
 To describe the effects of the random factors, we will refer only to the full model that included all species in the dataset.
-The strongest effect was that of the phylogeny (mean = `r MSphylo`, CI = `r MSphylol` - `r MSphyloh`), followed by the source of the data (mean = `r MSsource`, CI = `r MSsourcel` - `r MSsourceh`), the seed lot (mean = `r MSpopulation`, CI = `r MSpopulationl` - `r MSpopulationh`), and species identity (mean = `r MSspecies`, CI = `r MSspeciesl` - `r MSspeciesh`).
+The strongest effect was that of the phylogeny (mean = `r MSphylo`, CI = `r MSphylol` - `r MSphyloh`), followed by the source of the data (mean = `r MSsource`, CI = `r MSsourcel` - `r MSsourceh`) and the seed lot (mean = `r MSpopulation`, CI = `r MSpopulationl` - `r MSpopulationh`).
 The phylogenetic signal in the germination responses was relatively high (lambda = `r MSlambda`, CI = `r MSlambdal` - `r MSlambdah`).
@@
 
+* Mowing for haymaking thus is a predictable perturbation that occurs approximately at the same time every year, and which is expected to contribute to the high diversity of this habitats [@RN4887; @RN4888].
+Hay meadow plants must establish, grow and reproduce in the time frame determined by this perturbation.
+
+* Question of clonal regeneration.
+
+* Discuss second axis = geophytes vs hemicriptófitos?
+
 # Acknowledgements
@@
 
-E.F.P. conceived the study, performed the analyses and wrote the manuscript. All authors contributed data, revised the manuscript, and approved the final version.
+All authors contributed data. E.F.P. conceived the study and performed the analyses. E.F.P. wrote the manuscript with help from A.C. All authors revised the manuscript, and approved the final version.
 
@@
 ```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Diversity of dispersal units in mesic grasslands: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* subsp.
-*despectus* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadencis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (ñ) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
+*despectus* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadencis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (o) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
 knitr::include_graphics("../results/figures/seeds.png")
@@
 When the credible intervals overlap with the zero-effect line, the effect is not significant.
-The figure shows the results of a model including all species, and of models including only the three main floristic groups of mesic grasslands"}
+The figure shows the results of a model including all species, and of models including only the three main floristic groups of mesic grasslands (i.e.
+the Poacaea, the Fabaceae and the remaining families)."}
 knitr::include_graphics("../results/figures/mcmc.png")

### 2021-04-30 src/figures.R

@@ -16,13 +16,13 @@ read.csv("data/traits.csv") %>%
   mutate(Family = fct_relevel(Family, c("Poaceae", "Fabaceae", "Others"))) %>%
   mutate(Trait = fct_relevel(Trait, c("Seed.mass", "Length", "Shape"))) %>%
   mutate(Trait = fct_recode(Trait, 
-                            "Seed mass (mg)" = "Seed.mass",
-                            "Seed length (mm)" = "Length",
-                            "Seed shape (length / width)" = "Shape")) %>%
-  ggplot(aes(Family, Value, color = Family, fill = Family)) + 
+                            "Seed mass (log)" = "Seed.mass",
+                            "Seed length (log)" = "Length",
+                            "Seed shape (length / width, log)" = "Shape")) %>%
+  ggplot(aes(Family, log(Value), color = Family, fill = Family)) + 
   geom_violin(alpha = 0.5, draw_quantiles = c(0.25, 0.5, 0.75)) +
   facet_wrap( ~ Trait, strip.position = "left", scales = "free_y") +
-  scale_y_continuous(labels = scales::number_format(accuracy = 1)) +
+  scale_y_continuous(labels = scales::number_format(accuracy = .1)) +
   ggthemes::theme_tufte() +
   theme(panel.background = element_rect(color = "grey96", fill = "grey96"),
         strip.text = element_text(size = 12, color = "black"),
@@ -142,6 +142,9 @@ read.csv("data/germination.csv") %>%
 ## Do PCA
 
 traits %>%
+  gather(Trait, Value, Temperature:Seed.mass) %>%
+  mutate(Value = log(0.01+ Value)) %>%
+  spread(Trait, Value) %>%
   select(-c(Taxon, Family)) %>%
   FactoMineR::PCA(graph = FALSE) -> pca
 

### 2021-04-30 src/models.R

@@ -17,8 +17,7 @@ nbur = 50000
 list(R = list(V = 1, nu = 50), 
      G = list(G1 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 500), 
               G2 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 500),
-              G3 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 500),
-              G4 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 500)
+              G3 = list(V = 1, nu = 1, alpha.mu = 0, alpha.V = 500)
      )) -> 
   priors
 
@@ -37,7 +36,7 @@ read.csv("data/germination.csv") %>%
 MCMCglmm::MCMCglmm(cbind(Germinated, Germinable - Germinated) ~ scale(Scarification) + 
                      scale(Stratification) + scale(GA3) + scale(Tmean) + 
                      scale(Alternating) + scale(Light), 
-                   random = ~ animal + ID + Reference + Population:Reference, 
+                   random = ~ animal + Reference + Population:Reference, 
                    family = "multinomial2", pedigree = nnls_orig, prior = priors, data = data,
                    nitt = nite, thin = nthi, burnin = nbur, 
                    verbose = FALSE, saveX = FALSE, saveZ = FALSE, saveXL = FALSE, 
@@ -54,7 +53,7 @@ data %>% filter(Family == "Poaceae") -> poaceaedf
 MCMCglmm::MCMCglmm(cbind(Germinated, Germinable - Germinated) ~ scale(Scarification) + 
                      scale(Stratification) + scale(GA3) + scale(Tmean) + 
                      scale(Alternating) + scale(Light), 
-                   random = ~ animal + ID + Reference + Population:Reference, 
+                   random = ~ animal + Reference + Population:Reference, 
                    family = "multinomial2", pedigree = nnls_orig, prior = priors, data = poaceaedf,
                    nitt = nite, thin = nthi, burnin = nbur, 
                    verbose = FALSE, saveX = FALSE, saveZ = FALSE, saveXL = FALSE, 
@@ -71,7 +70,7 @@ data %>% filter(Family == "Leguminosae") -> leguminosaedf
 MCMCglmm::MCMCglmm(cbind(Germinated, Germinable - Germinated) ~ scale(Scarification) + 
                      scale(Stratification) + scale(GA3) + scale(Tmean) + 
                      scale(Alternating) + scale(Light), 
[…]

### 2021-04-30 src/numbers.R

@@ -7,8 +7,9 @@ load("../results/models/allspp.RData")
 # Calculate numbers
 
 germination %>% pull(Taxon) %>% length -> MSrecords
-germination %>% pull(Taxon) %>% unique %>% length -> MSspecies
-sum(germination$Number_seeds) -> MSseeds
+germination %>% pull(Taxon) %>% unique %>% length -> MSspeciesN
+germination %>% pull(Country) %>% unique %>% length -> MScountries
+sum(germination$Germinable) -> MSseedsN
 min(germination$Tmean) -> MSminT
 max(germination$Tmean) -> MSmaxT
 germination %>% filter(Alternating == 1) %>% pull(Taxon) %>% length -> MSaltY
@@ -54,17 +55,13 @@ summary(allspp)$Gcovariances[1, 1] %>% round(2) -> MSphylo
 summary(allspp)$Gcovariances[1, 2] %>% round(2) -> MSphylol
 summary(allspp)$Gcovariances[1, 3] %>% round(2) -> MSphyloh
 
-summary(allspp)$Gcovariances[2, 1] %>% round(2) -> MSspecies
-summary(allspp)$Gcovariances[2, 2] %>% round(2) -> MSspeciesl
-summary(allspp)$Gcovariances[2, 3] %>% round(2) -> MSspeciesh
+summary(allspp)$Gcovariances[3, 1] %>% round(2) -> MSpopulation
+summary(allspp)$Gcovariances[3, 2] %>% round(2) -> MSpopulationl
+summary(allspp)$Gcovariances[3, 3] %>% round(2) -> MSpopulationh
 
-summary(allspp)$Gcovariances[4, 1] %>% round(2) -> MSpopulation
-summary(allspp)$Gcovariances[4, 2] %>% round(2) -> MSpopulationl
-summary(allspp)$Gcovariances[4, 3] %>% round(2) -> MSpopulationh
-
-summary(allspp)$Gcovariances[3, 1] %>% round(2) -> MSsource
-summary(allspp)$Gcovariances[3, 2] %>% round(2) -> MSsourcel
-summary(allspp)$Gcovariances[3, 3] %>% round(2) -> MSsourceh
+summary(allspp)$Gcovariances[2, 1] %>% round(2) -> MSsource
+summary(allspp)$Gcovariances[2, 2] %>% round(2) -> MSsourcel
+summary(allspp)$Gcovariances[2, 3] %>% round(2) -> MSsourceh
 
 # PCA numbers
 

### 2021-05-20 doc/manuscript.Rmd
new intro, discu, new figs 🌿

@@
 
-*Arrhenatheretalia*, plant regeneration, seed germination, seed morphology, species-rich meadows, hay meadows, mesic pastures
+*Arrhenatheretalia*, *Arrhenatherion*, plant regeneration, seed germination, seed morphology, species-rich meadows, hay meadows, mesic pastures
 
@@
 European mesic grasslands are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698].
-These plant communities developed by anthropogenic intervention in clearings of a formerly wooded landscape and are maintained by agricultural practices, contributing to the diversity and conservation value of European landscapes [@RN4996].
-Traditional phytosociological classification of mesic grasslands put emphasis on the separation between mesic meadows used for hay making versus mesic pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893].
+These plant communities developed in clearings of a formerly temperate woodland landscape as a consequence of anthropogenic intervention [@RN5033].
+At a time, they were novel plant communities with a composite flora made up of species from different origins: grasslands plants that existed in non-forested "islands" of the landscape, plus plants that evolved in parallel to human intervention by hybridisation or polyploidization [@RN5033].
+During historical times the botanical composition of mesic grasslands evolved following changes in human activities and management practices [@RN5035].
+They are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034].
+Nowadays, mesic grasslands maintained by agricultural practices are a definitory aspect of European landscapes, both real [@RN4996] and imagined [@RN5029].
+
+Traditional phytosociological classification of mesic grasslands puts emphasis on the separation between mesic meadows used for hay making versus mesic pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893].
 Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic grasslands in large parts of Europe [@RN4997].
@@
 
-One of the reasons for the high conservation interest of traditional mesic grasslands is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed materials to be used in ecological restoration, rewilding and urban greening [@RN4894].
+One of the reasons for the high conservation interest of traditional mesic grasslands is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed materials to be used in ecological restoration, rewilding and urban greening [@RN5030; @RN4894; @RN5031].
 However, a lack of knowledge about species germination traits has already been identified as a bottleneck [@RN2241] hampering the development of a competitive native seed industry [@RN4665] in Europe.
-Understanding seed germination is part of the practical scientific framework needed to tackle large-scale ecological restoration challenges [@RN3507] and to maintain ex situ collections of plant genetic resources [@RN3763].
+Understanding seed germination is part of the practical scientific framework needed to tackle large-scale ecological restoration challenges [@RN3507] and to maintain *ex situ* collections of plant genetic resources [@RN3763].
 When such a knowledge framework is available, ambitious regional schemes of seed-based landscape restoration can be designed [@RN4666].
 
-In the case of mesic grasslands, a concerted investigation of their seed germination ecology is missing, despite their conservation and restoration importance.
-Most of the available information is scattered in single-species studies [@RN2236] or using a limited range of germination conditions [@RN4894].
+Mesic grasslands are dominated by mesophilous grasses of the Poaceae family, which make up most of the biomass and define the structure of the vegetation.
+The dispersal unit in the grasses is generally the floret (**Fig.
+1a**), a composite structure made up of the caryopsis or grain (a dry, indehiscent and monospermic fruit) and its surrounding bracts (i.e.
+modified leaves): the lemma and the palea.
+Removal of the lemma and palea, as well as puncturing the pericarp, can aid in seed germination [@RN1248].
+This structure is often accompanied by awns or hairs (**Fig.
+1b**) that are assumed to aid dispersal, although it is not always the case that they do [@RN1290].
+Grass seeds have been reported as having physiological seed dormancy [@RN5018; @RN3214; @RN5024] in various degrees, but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273].
+Freshly harvested seeds are comparatively more dormant, but dormancy tends to disappear quickly in dry storage [@RN5024; @RN5018].
+Germination has been reported to occur at temperatures ranging from 5 to 30 ºC [@RN3273; @RN1229; @RN1261; @RN1230; @RN1292; @RN5024].
+Most Poaceae species have also been reported to germinate better in light than in darkness [@RN1271; @RN1248; @RN1230; @RN1221; @RN1172; @RN1292; @RN1280; @RN5024] and even to be inhibited to germinate by darkness [@RN1229].
+Nonetheless, species of *Bromus* have been reported as germinating better in darkness [@RN1172] and *Cynosurus cristatus* as being indifferent to light/darkness [@RN1292; @RN1280].
+In *Poa trivialis*, germination was promoted by light but not by alternating temperatures [@RN1205], although alternating temperatures did encourage some germination in darkness [@RN1229].
+Wild grass ecotypes usually have a germination response to alternating temperatures [@RN1248; @RN1261; @RN1230; @RN1221; @RN1290; @RN1172; @RN1292; @RN1280], although this response is missing in some wild ecotypes and in the domesticated cultivars [@RN1261], and some species such as *Lolium perenne* have been reported as insensitive to temperature alternation [@RN1280; @RN1380].
+
+Next in abundance to the grasses are the legumes of the Fabaceae family.
+Legumes contribute to the nutritional value of grassland fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040].
+The dispersal unit of most legumes is the seed itself (**Fig.
+1c**) but in some species dispersal units are more complex, including indehiscent monospermic fruits (e.g.
+*Onobrychis*, **Fig.
+1d**) or indehiscent monospermic fruit fragments, i.e.
+loments (e.g.
+*Ornithopus*, **Fig.
+1e**).
+Legumes are generally hard-seeded, i.e.
+they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e.
+physical seed dormancy) [@RN3214; @RN3273; @RN1215; @RN5016; @RN3073].
+In *Medicago*, seeds that have not reached full maturity can germinate and are not yet impermeable, but the completion of maturation imposed coat impermeability, and thereafter the seed coat must be scarified to allow water imbibition and germination [@RN652].
+Buried legume seeds are able to track the seasons, and in some species germination seems to be promoted by cold stratification and alternating temperatures [@RN780], some of these have been described as having combinational dormancy (i.e.
+physical + physiological) [@RN5037].
+However, seeds have also been reported to germinate without any previous treatment [@RN881; @RN632; @RN460; @RN372], and to lose dormancy during storage [@RN5037].
+As in the grasses, legume seeds have been reported to germinate in high numbers across a range of temperatures from 5 to 25 ºC [@RN652; @RN3273].
+Legume seeds have been described as not responsive to light and capable of germinating in darkness [@RN1339; @RN3273].
+
+Grasses and legumes are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the aesthetic value of the meadows [@RN5038; @RN5039].
+They also add nutritional quality, being richer than grasses and legumes in specific elements [@RN5028; @RN5040].
+In many of the most frequent minor families (e.g.
+in the Asteraceae, Ranunculaceae, Polygonaceae, Lamiaceae, Dipsacaceae, Cyperaceae) the dispersal unit is the achene: dry, indehiscent and monospermic fruits.
+The morphology of these achenes is varied: cypselas with a hairy pappus in *Centaurea* (Asteraceae, **Fig.
+1f**); beaked in *Ranunculus* (Ranunculaceae, **Fig.
+1g**); hardened nutlets in *Prunella* (Lamiaceae, **Fig.
+1h**); hairy in *Knautia* (Dipsacaceae, **Fig.
+1i**); surrounded by a perigynium which aids in dispersal by water in *Carex* (Cyperaceae, **Fig.
+1j**).
+In the genus *Sanguisorba* (Rosaceae), the dispersal unit is the urn-shaped receptacle containing one to three achenes (**Fig.
+1k**).
+In the Apiaceae, it is the mericarp (**Fig.
+1l**), an indehiscent monospermic fragment of the fruit.
+In some other minor families the dispersal unit is the seed itself, such as in the Caryophyllaceae (**Fig.
+1m**), the Plantaginaceae (**Fig.
+1n**), the Juncaceae or the hemiparasitic species of *Rhinanthus* (Orobanchaceae) (**Fig.
+1o**).
+Our understanding of the germination ecology of these seeds is as varied as their morphology.
+In the Asteraceae, high germination without pretreatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; RN120; @RN963; @RN1428] and *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403].
+*Taraxacum officinale* germinated between 5 and 30 ºC [@RN120; RN1428; @RN1276], and had higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428].
+In *Stachys officinalis* (Lamiaceae), seeds need either cold stratification, light or alternating temperatures to germinate [@RN416; @RN128].
+Underdevelopped embryos that need to grow inside the seed before germination (i.e.
+morphological dormancy) are widespread in the Ranunculaceae and the Apiaceae [@RN3214; @RN1112].
+*Ranunculus repens* (Ranunculaceae) germinated between 10 and 25 ºC, but the germination percentages were low [@RN931]; the same species has been reported to respond to alternating temperatures, which can promote its germination even in darkness [@RN3180].
+In the Polygonaceae, *Rumex acetosa* can germinate inmediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not, this differences is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814].
+In *Heracleum sphondylium* (Apiaceae), growth of the embryo only occurs below 10 ºC, in moist conditions [@RN1112].
+*Sanguisorba minor* (Rosaceae) increased its germination after abrasion of the seeds with bleach [@RN194; @RN226], although germination without pre-treatment has also been reported [@RN298; @RN226].
+Seeds of the hemiparasitic species *Rhinanthus angustifolius* and *Rhinanthus minor* (Orobanchaceae) required relatively long periods of cold stratification to germinate [@RN4014], and were able to germinate in the dark [@RN2236].
+
+Although a wealth of studies has accumulated, a synthesis of the seed germination ecology of mesic grasslands is missing, despite their conservation, restoration and cultural significance.
 In this article, we present for the first time a synthesis of germination ecology in European mesic grasslands.
-To do so, we combined a newly-generated dataset with germination records from ENSCOBASE, the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968].
+To do so, we combined a newly-generated dataset with germination records from ENSCOBASE [@RN5036], the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968].
 To the resulting dataset containing `r MSrecords` germination records from `r MScountries` European countries we applied Bayesian meta-analysis [@RN5000] to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light.
@@
 
-In addition to this experimental data, we retrieved seed germination records from ENSCOBASE, the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968].
+In addition to this experimental data, we retrieved seed germination records from ENSCOBASE [@RN5036], the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968].
 The combined dataset, including our own experimental data and the records from ENSCOBASE and *SylvanSeeds*, contained `r MSrecords` germination records (i.e.
@@
 
-To test the effect of the germination treatments on seed germination proportions, we performed a meta-analysis of the germination dataset by fitting binomial generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) using the R package *MCMCglmm* [@RN4755].
+To test the effect of the germination treatments on seed germination proportions, we performed a meta-analysis of the germination dataset by fitting binomial generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) [@RN5036; @RN4965] using the R package *MCMCglmm* [@RN4755].
 We fitted the analysis separately for the three major groups of mesic grassland plants: Poaceae, Fabaceae, and the minor families.
@@
 
-The dispersal unit in the dominant grass family (Poaceae) is generally the floret (**Fig.
-1a**), a composite structure made up of the caryopsis (a dry, indehiscent and monospermic fruit) and several surrounding bracts (i.e.
-modified leaves): the lemma, the palea and, in many species, two glumes.
-This structure is often accompanied by awns or hairs that aid in dispersal (**Fig.
-1b**).
 The mass of these dispersal units in grassland grasses ranges from `r MSpoaMin` to `r MSpoaMax` mg, and they tend to be longer (`r MSpoaLength` mm on average) than they are wide (`r MSpoaWidth` mm).
-In the subordinate family of the legumes (Fabaceae), the dispersal unit of most species is the seed itself (**Fig.
-1c**).
-Fabaceae generally are hard-seeded, i.e.
-they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e.
-physical seed dormancy).
-In some Fabaceae, dispersal units are more complex, including indehiscent monospermic fruits (e.g.
-*Onobrychis*, **Fig.
-1d**) or indehiscent monospermic fruit fragments, i.e.
-loments (e.g.
-*Ornithopus*, **Fig.
-1e**).
+
 In mesic grasslands, the dispersal units of the Fabaceae are generally heavier than those of the Poaceae, ranging from `r MSlegMin` to `r MSlegMax` mg; and they are also rounder (`r MSlegLength` mm long vs.
 `r MSlegWidth` mm wide).
-In many of the most frequent minor families, such as the Asteraceae, the Ranunculaceae, the Lamiaceae and the Dipsacaceae, the dispersal unit is the achene, i.e.
-a dry, indehiscent and monospermic fruit.
-In the case of the Asteraceae, these achenes tend to be cypselas, wearing a pappus that aids in dispersal by wind (**Fig.
-1f**).
-In the Ranunculaceae, the achenes are usually flattened and sometimes beaked (**Fig.
-1g**).
-The achenes of the Lamiaceae are hardened nutlets (**Fig.
-1h**).
-In the genus *Knautia* (Dipsacaceae), the achene is covered by hairs (**Fig.
-1i**).
-In the Cyperaceae, the dispersal unit is an achene surrounded by a perigynium which aids in dispersal by water (**Fig.
-1j**).
-In the genus *Sanguisorba* (Rosaceae), the dispersal unit is the urn-shaped receptacle containing one to three achenes (**Fig.
-1k**).
-The dispersal unit of the Apiaceae is the mericarp (**Fig.
-1l**), an indehiscent monospermic fragment of the fruit.
-In some other minor families the dispersal unit is the seed itself, such as in the Caryophyllaceae (**Fig.
-1m**) or the Plantaginaceae (**Fig.
-1n**).
-The hemiparasitic species of *Rhinanthus* (Orobanchaceae) disperse by winged seeds (**Fig.
-1o**).
+
 When comparing the morphology of all species (**Fig.
@@
 
-* Mowing for haymaking thus is a predictable perturbation that occurs approximately at the same time every year, and which is expected to contribute to the high diversity of this habitats [@RN4887; @RN4888].
-Hay meadow plants must establish, grow and reproduce in the time frame determined by this perturbation.
+Our meta-analysis of germination records showed that, overall, the environmental factors with the strongest effect to control seed germination in mesic grassland species were alternating temperatures, light and scarification.
+Alternating temperatures and light are usually considered to be micro-environmental cues that indicate, at a fine scale, the existence of safe sites for regeneration [@RN4719].
+The diurnal alternation of temperatures decreases with burial depth in the soil, and the alternation is also reduced by vegetation cover [@RN1380; @RN4910; @RN4952].
+Thus, a positive germination response to alternating temperatures could detect the depth of seed burial, and prevent the germination of seeds that are buried too deep for seedlings to survive before emerging from the ground [@RN4898].
+Perhaps more importantly in this habitat, alternating temperatures could also indicate that the vegetation cover has been diminished by either sowing or grazing, and therefore mark an appropriate time for seedling regeneration, when the competition by the established plants is lessened.
+Similar functions can be ascribed to the germination response to light, as physiologically-active light in significant amounts only travels the uppermost milliliters of soil [@RN4909] and the quality of light will be affected by vegetation cover [@RN604].
+Therefore, the germination response to alternating temperatures and light indicate conditions that are found immediately after mowing for haymaking, a predictable perturbation that occurs approximately at the same time every year, and which forces meadow plants to establish, grow and reproduce in the time frame determined by this perturbation [@RN4887; @RN4888].
+In addition, this traits would also promote the formation of a persistent soil seed bank [@RN4718; @RN4929; @RN1292].
+On the other hand, it is worth mentioning that the Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate grasslands [RN5014; @RN5013].
+
+The strong effect of scarification was specifically related to the hard-seeded Fabaceae.
+In this family, physical dormancy as a result of an unpermeable seed coat works as a mechanism to detect seasonal cycles of temperature and humidity [@RN780].
+It has been also proposed that hardseedeness can avoid seed predation, by preventing the emission of olfactory cues that are elicited by seed imbibition and that can be perceived by seed predators [@RN2568].
+In the case of this study, Fabaceae seeds are the heaviest, and they clearly follow a separate regeneration strategy from that of the other families.
+While other groups rely on alternating temperatures and light as gap-detecting mechanisms, Fabaceae seeds do not respond to these factors, and instead they are mainly regulated by scarification, and they also respond to cooler temperatures than the rest of species.
+This would indicate that Fabaceae seeds tend to germinate when temperatures are cooler, and thus before or after the summer hay-making season.
+Their larger size, and the related larger reserves, could allow Fabaceae seedlings to emerge at times when the competition from the established vegetation would be higher.
+
+The germination response to average temperatures and stratification is understood to detect cues related to seasonal cycling and macroclimatic variation [@RN3063, @RN5036].
+Apart from the aforementioned response to cool temperatures in the Fabaceae, the rest of the families did not show a response to average temperatures, further highlighting their reliance on micro-habitat cueing.
+Moreover, the dominant family of the Poaceae showed a negative response to cold stratification.
+The minor families, however, showed a divergence from the Poaceae strategy in their positive response to stratification.
+These subordinate species appear to require a period of cold stratification, indicating that overwintering has occured, before they can germinate [@RN3214].
+These minor families are also the only group that shows a positive response to GA~3~, a phytohormone which can work as a substitute of cold stratification to overcome physiological dormancy [@RN3368].
+
+The positive phylogenetic signal found in the models highlights the phylogenetic clustering of regeneration strategies that we have described so far: (1) the dominant family of the Poaceae, showing a lack of responses to seasonal cues (average temperature and cold stratification), and relying instead on micro-niche cues (alternating temperatures, light) that can be associated to the yearly perturbation of mowing, (2) the second most-dominant family, the Fabaceae, which do not respond to the micro-niche cues, and regulate their germination timing through scarification and cooler germination temperatures, possibly because their larger size and reserves allow their seeds to decouple their emergence timing from the mowing perturbance, and finally (3) the minor families that respond to micro-cues in a similar way than the Poaceae, but which are differentieted from them by showing a positive response to cold stratification and G~3~, indicating that they rely on physiological seed dormancy.
+
+* Less dormancy in grassland vs.
+arable populations [@RN1229].
+Herbaceous plants adapted to open habitats needed less cold stratification than congeneric species adapted to forest understories [@RN3570].
+
+* Better germination without treatments [@RN4894].
+
+* In Poland soil bank dominated by forbs of arable and weedy species, low representation of grasses and legumes, confirms the typical Arrhenatherion species do not form a seed bank [@RN5032].
+Soil bank is transient in hay meadows and related grasslands [@RN5026; @RN5041; @RN5042].
+Regeneration limited by dispersal limitation [@RN3073].
+
+* Mowing promotes germination and emergence [@RN3073]. Mowing promotes autumn germination [@RN5025].
 
@@
 * Discuss second axis = geophytes vs hemicriptófitos?
+
+* Conclusions/applications
+
+* Discuss dormancy in freshly collected grasses, limitations of stored seed data [@RN5018]
 

### 2021-05-20 src/cleaning.R

@@ -148,7 +148,13 @@ read.csv("../#data/germination/data/sospraderas/Accesiones morfo.csv") %>% # See
   merge(seedlots, by = "Seedlot") %>%
   rename(Taxon = Species) %>%
   group_by(Taxon) %>%
-  summarise(Length = mean(Height), Width = mean(Width)) -> morphometrics
+  select(Taxon, Country, Seedlot, Height, Width) %>%
+  rename(Population = Seedlot, Length = Height) %>%
+  filter(Width < 2*Length) -> morpho
+  
+morpho %>% 
+  group_by(Taxon) %>%
+  summarise(Length = mean(Length), Width = mean(Width)) -> morphometrics
 
 # Clean seed mass
 
@@ -189,18 +195,24 @@ rbind(dormancy1, dormancy2) -> dormancy
 
 rm(dormancy1, dormancy2)
 
+# Lifeforms
+
+read.csv("../#data/lifeforms/results/lifeforms.csv") %>%
+  rename(Taxon = TPLName) -> lifeforms
+
 # Merge species traits
 
 dominances %>%
   filter(Dominance != "Transient") %>%
+  merge(lifeforms, all.x = TRUE) %>%
   merge(sncs, all.x = TRUE) %>%
   merge(mass, all.x = TRUE) %>%
   merge(morphometrics, all.x = TRUE) %>%
   merge(dormancy, all.x = TRUE) %>%
   merge(read.csv("../#data/tpl/results/TPLNames.csv")) %>%
-  select(Taxon, Family, Abundance, Dominance, bio06, bio14, pH, Sand, Seed.mass, Length, Width, Dormancy) -> traits0
+  select(Taxon, Family, Lifeform, Abundance, Dominance, bio06, bio14, pH, Sand, Seed.mass, Length, Width, Dormancy) -> traits0
 
-rm(dominances, dormancy, mass, morphometrics, sncs, species, header)
+rm(dominances, dormancy, morphometrics, mass, sncs, lifeforms)
 
 # Extract higher taxonomical ranks
 
@@ -219,8 +231,24 @@ traits0 %>%
   merge(orders, all.x = TRUE) %>%
   merge(read.csv("../#data/germination/data/sospraderas/clades.csv"), all.x = TRUE) %>%
   rename(Order = order, Clade = clade) %>%
-  select(Taxon, Family, Order, Clade,
[…]

### 2021-05-20 src/figures.R

@@ -12,6 +12,7 @@ read.csv("data/traits.csv") %>%
          Family = ifelse(Family %in% "Leguminosae", "Fabaceae", Family),
          Shape = Length / Width) %>%
   select(-Width) %>%
+  select(-Length) %>%
   gather(Trait, Value, Seed.mass:Shape) %>%
   mutate(Family = fct_relevel(Family, c("Poaceae", "Fabaceae", "Others"))) %>%
   mutate(Trait = fct_relevel(Trait, c("Seed.mass", "Length", "Shape"))) %>%
@@ -184,3 +185,160 @@ ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
 
 ggsave(f3, file = "results/figures/pca.png", 
        path = NULL, scale = 1, width = 100, height = 100, units = "mm", dpi = 600)
+
+# Figure other habitats
+
+read.csv("data/others.csv") %>%
+  select(-c(C14, C22, C30)) -> others
+
+read.csv("data/germination.csv") %>%
+  filter(Reference == "SOSPRADERAS") %>%
+  filter(GA3 == 0) %>%
+  merge(read.csv("data/traits.csv")) %>%
+  filter((Family == "Leguminosae" & Scarification == 1) | Family != "Leguminosae") %>%
+  group_by(Taxon, Seed.mass, Tmean) %>%
+  summarise(G = sum(Germinated) / sum(Germinable)) %>%
+  spread(Tmean, G) %>%
+  group_by(Taxon) %>%
+  mutate(Habitat = "Mesic grasslands") %>%
+  rename(F14 = `9`, F22 = `17`, F30 = `25`) %>%
+  select(Taxon, Habitat, Seed.mass:F30) %>%
+  rbind(others) %>%
+  group_by() -> df1
+
+## Do PCA
+
+df1 %>%
+  select(-c(Taxon, Habitat, Seed.mass)) %>%
+  FactoMineR::PCA(graph = FALSE, scale = FALSE) -> pca
+
+df1 %>%
+  select(c(Taxon, Habitat)) %>%
+  cbind(pca$ind$coord[, 1:2]) -> pcaInds
+
+pca$var$coord[, 1:2] %>%
+  data.frame %>%
+  rownames_to_column(var = "Variable") %>%
+  mutate(Variable = fct_recode(Variable, `22/12 ºC` = "F22", 
[…]

### 2021-05-27 doc/manuscript.Rmd
1st full draft 🌱

@@
 ---
-title: "Seed germination ecology of European mesic grasslands"
+title: "Seed ecology of European mesic meadows: the domestication of plant regeneration strategies"
 output: 
@@
 
-*Arrhenatheretalia*, *Arrhenatherion*, plant regeneration, seed germination, seed morphology, species-rich meadows, hay meadows, mesic pastures
+*Arrhenatheretalia*, *Arrhenatherion*, plant regeneration, seed germination, seed morphology, species-rich meadows, hay meadows, mesic grasslands
 
@@
 
-European mesic grasslands are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698].
-These plant communities developed in clearings of a formerly temperate woodland landscape as a consequence of anthropogenic intervention [@RN5033].
-At a time, they were novel plant communities with a composite flora made up of species from different origins: grasslands plants that existed in non-forested "islands" of the landscape, plus plants that evolved in parallel to human intervention by hybridisation or polyploidization [@RN5033].
-During historical times the botanical composition of mesic grasslands evolved following changes in human activities and management practices [@RN5035].
+European mesic meadows are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698].
+These plant communities developed in clearings of a temperate woodland landscape as a consequence of anthropogenic intervention [@RN5033; @RN5045].
+At a time, they were novel plant communities with a composite flora made up of species from different origins: grasslands plants that existed in non-forested "islands" of the landscape, plus plants that evolved in parallel to human intervention by hybridization or polyploidization [@RN5033].
+During historical times the botanical composition of mesic meadows evolved following changes in human activities and management practices [@RN5035; @RN5045].
 They are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034].
-Nowadays, mesic grasslands maintained by agricultural practices are a definitory aspect of European landscapes, both real [@RN4996] and imagined [@RN5029].
+Nowadays, mesic meadows maintained by agricultural practices are an essential aspect of European landscapes, both real [@RN4996] and imagined [@RN5029].
 
-Traditional phytosociological classification of mesic grasslands puts emphasis on the separation between mesic meadows used for hay making versus mesic pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893].
-Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic grasslands in large parts of Europe [@RN4997].
-For this reason the European Habitats Directive (92/43/EEC) has included in its lists of habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.
+Traditional phytosociological classification of mesic grasslands puts emphasis on the separation between meadows used for hay making versus pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893].
+Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic meadows in large parts of Europe [@RN4997].
+For this reason, the European Habitats Directive (92/43/EEC) has included in its lists of habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.
 
-One of the reasons for the high conservation interest of traditional mesic grasslands is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed materials to be used in ecological restoration, rewilding and urban greening [@RN5030; @RN4894; @RN5031].
+One of the reasons for the high conservation interest of traditional mesic meadows is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed materials to be used in ecological restoration, rewilding and urban greening [@RN5030; @RN4894; @RN5031].
 However, a lack of knowledge about species germination traits has already been identified as a bottleneck [@RN2241] hampering the development of a competitive native seed industry [@RN4665] in Europe.
@@
 
-Mesic grasslands are dominated by mesophilous grasses of the Poaceae family, which make up most of the biomass and define the structure of the vegetation.
+Mesic meadows are dominated by mesophilous grasses of the Poaceae family, which make up most of the biomass and define the structure of the vegetation.
 The dispersal unit in the grasses is generally the floret (**Fig.
@@
 Removal of the lemma and palea, as well as puncturing the pericarp, can aid in seed germination [@RN1248].
-This structure is often accompanied by awns or hairs (**Fig.
+The dispersal unit is often accompanied by awns or hairs (**Fig.
 1b**) that are assumed to aid dispersal, although it is not always the case that they do [@RN1290].
-Grass seeds have been reported as having physiological seed dormancy [@RN5018; @RN3214; @RN5024] in various degrees, but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273].
+Meadow grass seeds have been reported as having physiological seed dormancy [@RN5018; @RN3214; @RN5024] in various degrees, but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273].
 Freshly harvested seeds are comparatively more dormant, but dormancy tends to disappear quickly in dry storage [@RN5024; @RN5018].
@@
 Next in abundance to the grasses are the legumes of the Fabaceae family.
-Legumes contribute to the nutritional value of grassland fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040].
+Legumes contribute to the nutritional value of meadow fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040].
 The dispersal unit of most legumes is the seed itself (**Fig.
@@
 1e**).
-Legumes are generally hard-seeded, i.e.
+Legume seeds are generally hard-seeded, i.e.
 they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e.
 physical seed dormancy) [@RN3214; @RN3273; @RN1215; @RN5016; @RN3073].
-In *Medicago*, seeds that have not reached full maturity can germinate and are not yet impermeable, but the completion of maturation imposed coat impermeability, and thereafter the seed coat must be scarified to allow water imbibition and germination [@RN652].
-Buried legume seeds are able to track the seasons, and in some species germination seems to be promoted by cold stratification and alternating temperatures [@RN780], some of these have been described as having combinational dormancy (i.e.
+In *Medicago*, seeds that have not reached full maturity can germinate before they become impermeable, but the completion of maturation imposes coat impermeability, and thereafter the seed must be scarified to allow water imbibition and germination [@RN652].
+Buried legume seeds are able to track the seasons, and in some species germination seems to be promoted by cold stratification and alternating temperatures [@RN780]; some of these species have been described as having combinational dormancy (i.e.
 physical + physiological) [@RN5037].
@@
 Grasses and legumes are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the aesthetic value of the meadows [@RN5038; @RN5039].
-They also add nutritional quality, being richer than grasses and legumes in specific elements [@RN5028; @RN5040].
+They also add nutritional scope, being richer than grasses and legumes in specific elements [@RN5028; @RN5040].
 In many of the most frequent minor families (e.g.
 in the Asteraceae, Ranunculaceae, Polygonaceae, Lamiaceae, Dipsacaceae, Cyperaceae) the dispersal unit is the achene: dry, indehiscent and monospermic fruits.
-The morphology of these achenes is varied: cypselas with a hairy pappus in *Centaurea* (Asteraceae, **Fig.
+The morphology of these achenes is varied: cypselae with a hairy pappus in *Centaurea* (Asteraceae, **Fig.
 1f**); beaked in *Ranunculus* (Ranunculaceae, **Fig.
@@
 1o**).
-Our understanding of the germination ecology of these seeds is as varied as their morphology.
-In the Asteraceae, high germination without pretreatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; RN120; @RN963; @RN1428] and *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403].
-*Taraxacum officinale* germinated between 5 and 30 ºC [@RN120; RN1428; @RN1276], and had higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428].
+In the Asteraceae, high germination without pretreatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; @RN120; @RN963; @RN1428], *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403].
+*Taraxacum officinale* germinates between 5 and 30 ºC [@RN120; RN1428; @RN1276], and has higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428].
 In *Stachys officinalis* (Lamiaceae), seeds need either cold stratification, light or alternating temperatures to germinate [@RN416; @RN128].
-Underdevelopped embryos that need to grow inside the seed before germination (i.e.
+Underdeveloped embryos that need to grow inside the seed before germination (i.e.
 morphological dormancy) are widespread in the Ranunculaceae and the Apiaceae [@RN3214; @RN1112].
-*Ranunculus repens* (Ranunculaceae) germinated between 10 and 25 ºC, but the germination percentages were low [@RN931]; the same species has been reported to respond to alternating temperatures, which can promote its germination even in darkness [@RN3180].
-In the Polygonaceae, *Rumex acetosa* can germinate inmediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not, this differences is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814].
+*Ranunculus repens* (Ranunculaceae) germinates between 10 and 25 ºC, but the germination percentages were reported to be low [@RN931]; the same species has been reported to respond to alternating temperatures, which can promote its germination even in darkness [@RN3180].
+In the Polygonaceae, *Rumex acetosa* can germinate immediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not, this difference is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814].
 In *Heracleum sphondylium* (Apiaceae), growth of the embryo only occurs below 10 ºC, in moist conditions [@RN1112].
-*Sanguisorba minor* (Rosaceae) increased its germination after abrasion of the seeds with bleach [@RN194; @RN226], although germination without pre-treatment has also been reported [@RN298; @RN226].
-Seeds of the hemiparasitic species *Rhinanthus angustifolius* and *Rhinanthus minor* (Orobanchaceae) required relatively long periods of cold stratification to germinate [@RN4014], and were able to germinate in the dark [@RN2236].
+*Sanguisorba minor* (Rosaceae) increases its germination after abrasion of the seeds with bleach [@RN194; @RN226], although germination without pre-treatment has also been reported [@RN298; @RN226].
+Seeds of the hemiparasitic species *Rhinanthus angustifolius* and *Rhinanthus minor* (Orobanchaceae) require relatively long periods of cold stratification to germinate [@RN4014], and are able to germinate in the dark [@RN2236].
 
-Although a wealth of studies has accumulated, a synthesis of the seed germination ecology of mesic grasslands is missing, despite their conservation, restoration and cultural significance.
-In this article, we present for the first time a synthesis of germination ecology in European mesic grasslands.
-To do so, we combined a newly-generated dataset with germination records from ENSCOBASE [@RN5036], the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968].
-To the resulting dataset containing `r MSrecords` germination records from `r MScountries` European countries we applied Bayesian meta-analysis [@RN5000] to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light.
-In addition, we also analyzed seed morphology (mass and shape) and the differences between the main functional groups of mesic grasslands: grasses (Poaceae), legumes (Fabaceae) and the other families.
+Although a wealth of studies has accumulated, a synthesis of the seed ecology of mesic meadows is missing, despite their conservation, restoration and cultural significance.
+In this article, we present for the first time a synthesis of regeneration by seed in European mesic meadows.
+To do so, we combine newly-generated data on seed morphology and germination with records from databases [@RN5036; @RN4968; @RN2987; @RN2319].
+To the resulting dataset containing `r MSrecords` germination records from `r MScountries` European countries we apply Bayesian meta-analysis [@RN5000] to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light.
+We analyze the relationship between seed traits and species environmental preferences, and compare the germination ecology of mesic meadows with that of other herbaceous plant communities.
 
@@
 
-## Selection of study species
+## Selection of mesic meadow species
 
-To create a species list of representative mesic grassland species for inclusion in our analysis, we used a dataset of 118 vegetation relevés from three Western European regions with well-maintained mesic grasslands: 43 from Northern Portugal (Trás-os-Montes), 25 from the Cantabrian Mountains of Spain (Parque Nacional Picos de Europa) and 50 from the Pyrenees.
-This dataset represents a sample of mesic grassland diversity, as the plots were situated along a major stress gradient related to drought [@RN2980]: two regions with a Temperate macroclimate, and Northern Portugal which is transitional between the Temperate and Mediterranean macroclimates [@RN4893].
-Furthermore, grasslands from the Pyrenees are closest to the central European optimum of mesic grassland vegetation (*Arrhenatheretalia elatioris* Tx.
-1931) while Portuguese grasslands are in suboptimal areas at the limit of the distribution [@RN4893].
-Finally, traditional management of mesic grasslands is relatively well preserved in this area within the European context [@RN4998; @RN4999].
+To create a list of representative mesic meadow species for inclusion in our analysis, we used a dataset comprising 118 vegetation relevés (i.e.
+records of plants species co-occurring in sampling plots) from three Western European regions with well-maintained mesic meadows: 43 relevés from Northern Portugal, 25 from the Cantabrian Mountains of Spain and 50 from the Pyrenees.
+These relevés contain a sample of mesic meadow diversity, as they were recorded along a major stress gradient related to summer drought [@RN2980]: the Pyrenees and Cantabrian Mountains have a temperate macroclimate, whereas Northern Portugal is transitional between the temperate and Mediterranean macroclimates.
+Furthermore, meadows from the Pyrenees are closest to the European optimum of mesic meadow vegetation while the Portuguese ones are in suboptimal areas at the limit of the European distribution of temperate meadows [@RN4893].
+Finally, the traditional management of meadows is relatively well preserved in these three regions compared to their European context [@RN4998; @RN4999].
 
-Each relevé had been conducted in a square area (25-100 m^2^ area) situated in the central part of a meadow avoiding the margins.
-The chosen meadows are maintained by traditional agricultural practices (mowing for hay-making once or twice per year and light grazing).
-Vegetation in the plots had been sampled in 2016-2017, at the peak of plant development, just before mowing.
-All vascular plant species in the plots had been recorded and assigned an abundance value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%).
-All plant names in this manuscript follow the nomenclature of the Euro+Med plant database (https://www.emplantbase.org/home.html).
-Using all the relevés, we calculated the cumulative abundance of each species in the entire area.
-We removed 208 species with cumulative abundance values below 2%, considering them transient species that might have been recorded by chance and do not represent the core mesic grasslands flora.
-We used the remaining 117 species as the core list to retrieve seed germination data for this article.
+Sampled meadows had been maintained by traditional agricultural practices: mowing for hay-making once or twice per year plus light grazing.
+Each sampling plot was placed in a square area (25-100 m^2^ area) situated in the central part of a meadow, avoiding the margins.
+Sampling took place in 2016-2017, at the peak of plant development, just before mowing.
+All vascular plant species in the plots were recorded and given a cover value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%).
+All plant names were assigned following the nomenclature of Euro+Med [-@RN5044], which is used throughout this article.
+As expected, the vegetation of the sampled meadows was dominated by the Poaceae and the Fabaceae: these two families represented 47% and 17%, respectively, of the total plant cover recorded in all the plots.
+Other 20 families were recorded, the largest of which was the Asteraceae; each of these other families represented less than 10% of the total cover.
+These numbers refer only to the seed plants, pteridophytes being a marginal part of the records.
 
-## Seed collection and morphology
+Using all the relevés, we calculated the cumulative cover of each species in the entire area.
+To perform the calculation, first we standardized the cover values of the plots by dividing the cover of each species in each plot by the total plant cover in that plot.
+Then, for each species, we calculated its total cover in the dataset, by summing its standardized cover values from all the plots.
+Finally, we rescaled the values of all species to a 1-100 scale to obtain the cumulative cover values.
+We removed 208 species with cumulative cover values below 2%, considering them to be transient species that might have been recorded by chance and may not represent the core mesic meadows flora.
+We used the remaining 116 species as the core list to retrieve seed germination data for this article.
+The majority of these species were hemicryptophytes (78%), with some therophytes (16%) and a few chamaephytes and geophytes (3% of each).
+The family with more species in the list was the Poaceae (22%), with another 17% belonging to the Fabaceae, 15% to the Asteraceae, 8% to the Apiaceae, and the rest of the families representing less than 5% each.
+
+We also used the relevés as a basis to characterize the preferences of the selected species for three ecological drivers of mesic meadow diversity: cold, summer drought and soil reaction [@RN2980].
+For cold and drought, we used the coordinates of the plots to retrieve from CHELSA [@RN4766] the bioclimatic variables bio06 (minimum temperature of the coldest month) and bio14 (precipitation of the driest month).
+and represents the limiting water availability along the Iberian gradient of summer drought.
+For soil reaction, we took from each plot five soil samples from between 0 and 20 cm depth with a Dutch auger and combined them to make a bulk soil horizon, which we subsequently air-dried, crumbled, finely crushed and sieved with a 2 mm screen, to finally measure the pH in H~2~O with a glass electrode in a suspension of soil and water (1:2.5).
+For each of these three environmental variables (bio06, bio14, pH) measured at the plot level we calculated the species niche centroids (SNCs).
+The SNC for any given species and variable is the mean of the environmental variable in all the plots were the species occurs, weighted by species cover in each plot [@RN4709].
+The list of core species, with their cumulative covers and SNCs, is available at GitHub (see Data Availability Statement).
+
+## Seed morphology and germination dataset
 
 From the vegetation plots described above, we collected dispersal units (hereafter called seeds) during the dispersal seasons of 2016, 2017 and 2018.
+Seed collection followed the methodology of ENSCONET [-@RN2907].
 To describe seed morphology, we acquired images of 100-seed samples of each species using a flatbed scanner (Brother LC985) with a resolution of 200 dpi and a scanning area of 1024 x 1024 pixels [@RN2678].
-We distributed the seeds on the scanner transparent glass, in a 10 x 10 grid.
+We distributed the seeds on the transparent glass of the scanner, in a 10 x 10 grid.
 For each sample, and without moving the seeds, we repeated the scans with black and white backgrounds.
@@
 We processed the scanned images using *ImageJ*, an open-source image processing program designed for scientific multidimensional images [@RN4708].
-The program calculates several biometric parameters for each seed on the sample, among these, seed length and width.
-Using these values, we calculated seed shape as seed length divided by seed width.
-Additionally, we obtained seed mass for the study species from Kew's Seed Information Database [@RN2987].
+The program calculates several biometric parameters for each seed on the sample, and among these we choose seed length and width.
+Additionally, we retrieved species values of seed mass from the Seed Information Database [@RN2987] and of seed number at the individual/ramet level from the LEDA database [@RN2319].
+The dataset with the length and width measures is available at GitHub (see Data Availability Statement).
 
-## Seed germination data
-
-We germinated seeds of the different species using three germination treatments to determine the effect of temperatures representative for the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as the optimal summer temperature, and 30/20 ºC as sun-heated soil, e.g.
+We germinated the collected seeds using three germination treatments to determine the germination response to temperatures that are representative of the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as the summer temperature, and 30/20 ºC as sun-heated soil, e.g.
 soil exposed to sun after hay cutting.
 Additionally, we compared, at each of these temperature regimes, the germination of fresh seeds versus seeds subjected to a dormancy-breaking treatment.
-In the case of the Fabaceae, the treatment consisted in scarification by chipping the seed coat with a scalpel, to remove physical dormancy [@RN3214].
+In the case of the Fabaceae and other families that might present physical dormancy [@RN3214], the treatment consisted in scarification by chipping the seed coat with a scalpel.
 For the rest of the families, we used gibberellic acid GA~3~ (0,0645 mM) in darkness during 24h, as a treatment to remove potential physiological seed dormancy [@RN4976].
@@
 We sealed dishes with Parafilm to prevent desiccation.
-Trials took place in a germination chamber (KBW 400, Binder GmbH, Tuttlingen, Germany) with a 12/12 hour photoperiod (the light period corresponding to the higher temperature).
+Trials took place in a germination chamber (KBW 400, Binder GmbH, Tuttlingen, Germany) with a 12/12-hour photoperiod (the light period corresponding to the higher temperature).
 Experiments lasted for four weeks, with germination scoring once per week.
@@
 
-In addition to this experimental data, we retrieved seed germination records from ENSCOBASE [@RN5036], the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968].
+In addition to this experimental data, we retrieved seed germination records from ENSCOBASE [@RN5036], the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for the nemoral biome [@RN4968].
 The combined dataset, including our own experimental data and the records from ENSCOBASE and *SylvanSeeds*, contained `r MSrecords` germination records (i.e.
 germination proportions for a given seed lot of a species, recorded in a set of laboratory experimental conditions) from `r MScountries` European countries.
-There were `r MSspeciesN` species in the dataset.
+There were `r MSspeciesN` species in the dataset (i.e.
+90% of the core list of meadow species was covered).
 Overall, `r MSseedsN` seeds had been used in the experiments.
-The range of experimental germination temperatures (weighted average of the daily thermoperiod) which had been used in the experiments spans from `r MSminT` to `r MSmaxT` ºC, with `r MSaltN` records of constant temperatures (i.e.
+The range of experimental germination temperatures (weighted average of the daily thermoperiod) which had been used in the experiments spanned from `r MSminT` to `r MSmaxT` ºC, with `r MSaltN` records of constant temperatures (i.e.
 experiments that used the same temperature during all their duration) and `r MSaltY` of alternating temperatures (i.e.
@@
 Finally, there were `r MSgaY` records where GA~3~ had been applied, and `r MSscarY` records where seeds had been scarified.
+The full germination dataset is available at GitHub (see Data Availability Statement).
 
@@
 
-To test the effect of the germination treatments on seed germination proportions, we performed a meta-analysis of the germination dataset by fitting binomial generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) [@RN5036; @RN4965] using the R package *MCMCglmm* [@RN4755].
-We fitted the analysis separately for the three major groups of mesic grassland plants: Poaceae, Fabaceae, and the minor families.
+We conducted all analyses in R [@RN2315], and the code for analysis and creation of the figures and manuscript is available at GitHub (see Data Availability Statement)
+
+To test the effect of germination cues on seed germination proportions, we performed a meta-analysis [@RN5000] of the germination dataset by fitting binomial generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) [@RN5036; @RN4965] using the R package *MCMCglmm* [@RN4755].
+We fitted models to the entire dataset, and also separately for each of the three botanical groups of mesic meadows: Poaceae, Fabaceae, and the other families.
 To account for the effect of a shared phylogeny in species traits, models included as a random effect a reconstructed phylogenetic tree for the study species.
-To create the phylogeny we used the R package *V.PhyloMaker* [@RN4753] which contains an updated mega-tree of the seed plants based on Smith & Brown [-@RN4754].
+To create the phylogeny, we used the R package *V.PhyloMaker* [@RN4753] which contains an updated mega-tree of the seed plants based on Smith & Brown [-@RN4754].
 We placed taxa absent from the mega-tree at the genus-level basal node.
-The phylogenetic tree is available in the data folder of the GitHub repository (see Data Availability Statement).
+The phylogenetic tree is available at GitHub (see Data Availability Statement).
 Germination proportion was the response variable.
@@
 When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in trait evolution that does not fully follow a Brownian motion model [@RN4757].
-Detailed results of the MCMCglmms are available in the results folder of the GitHub repository (see Data Availability Statement).
+Detailed results of the MCMCglmms are available at GitHub (see Data Availability Statement).
 
-To visualize the main patterns in the seed germination spectrum of mesic grassland plants, we did a Principal Component Analysis (PCA) as implemented in the package *FactoMineR* [@RN3166].
+To check whether seed traits and plant ecological preferences were related, we did a Principal Component Analysis (PCA) of seed traits and species SNCs for cold, drought and pH.
 We performed the PCA ordination at the species level, i.e.
-calculating a series of continuous traits for each species.
+calculating a series of continuous seed traits for each species.
 We transformed the final germination proportions to create a continuous variable for the germination cues (i.e.
@@
 To do so, for each cue and species, we calculated a weighted average of the cue levels (in the case of temperature, cue levels were the temperature treatments; for the other cues the levels were 0 = absence and 1 = presence), weighting by the germination proportion at each level.
-This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of the response to the germination cues when visualized across the whole dataset; it must be stressed that this stage of the analysis serves only for visualization and not for hypothesis testing.
-We also included seed mass in the ordination.
+This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of the response to the germination cues when visualized across the whole dataset.
+We also included seed mass and seed number in the ordination.
 We left GA~3~ out of the PCA because its ecological interpretation is subordinated to stratification (as both cues break physiological seed dormancy).
-We also left seed length, width and shape out because these values were not available for all species.
+We also left seed length and width out because these values were not available for enough species.
+We calculated the PCA with the package *FactoMineR* [@RN3166].
+
+Finally, to compare the germination of mesic meadow species with other herbaceous communities, we retrieved data from previous works on the seed germination ecology of bogs and fens [@RN3480; @RN3212], alpine and subalpine grasslands [@RN2371], and coastal plant communities of rocky cliffs and sand dunes [@RN2865].
+These additional germination records corresponded to the same experimental methodology as the one employed for some of the germination experiments of this study: recently-collected seeds, untreated for physiological dormancy (but scarified in the cases of families known of having physical dormancy), had been subjected to three germination thermoperiods (14/4 ºC, 22/12 ºC, 30/20 ºC) in the same conditions as described previously.
+All seeds had been collected in the Cantabrian Mountains of Spain and the neighboring coast.
+We added these records and the records with matching experimental conditions from the meadows germination dataset, and performed a PCA of the resulting dataset.
+The germination records for the other plant communities are available at GitHub (see Data Availability Statement).
 
@@
 
-The mass of these dispersal units in grassland grasses ranges from `r MSpoaMin` to `r MSpoaMax` mg, and they tend to be longer (`r MSpoaLength` mm on average) than they are wide (`r MSpoaWidth` mm).
+The Poaceae had lower values of seed mass and higher values of seed number, while the Fabaceae had heavier but fewer seeds (**Fig.
+2**).
+The other families covered the range of values showed by the Poaceae and the Fabaceae, but their median values were high for both traits: their median seed mass was close to that of the Fabaceae, while their median seed number was higher than that of the Poaceae (**Fig.
+2**).
+Seed shape also showed a divergence between the Poaceae and Fabaceae, with seeds being elongated in the former and rounder in the latter (**Fig.
+2**).
+The other families covered the full range of variation, with both elongated and round seeds (**Fig.
+2**).
 
-In mesic grasslands, the dispersal units of the Fabaceae are generally heavier than those of the Poaceae, ranging from `r MSlegMin` to `r MSlegMax` mg; and they are also rounder (`r MSlegLength` mm long vs.
-`r MSlegWidth` mm wide).
+## Seed germination
 
-When comparing the morphology of all species (**Fig.
-2**), the Poaceae appear as having relatively lighter and more elongated dispersal units, in contrast with the Fabaceae, which tend to have rounder and heavier seeds.
-
-## Germination responses
-
-When considering the full set of mesic grassland species, all of the six studied germination cues had a significant effect on final germination proportions (**Fig.
+When considering the full set of mesic meadow species, all of the six studied germination cues had a significant effect on final germination proportions (**Fig.
 3**).
-The germination of grassland seeds was positively associated with scarification, stratification, GA~3~, alternating temperatures, and light.
+The germination of meadow seeds was positively associated with scarification, stratification, GA~3~, alternating temperatures, and light.
 Average temperature had a negative effect, indicating a trend towards higher germination at lower temperatures.
@@
 Between 0 ºC and 20 ºC, germination proportions increased steadily with increasing temperatures.
-Above 20 ºC, germination declined more sharply, and grassland seeds rarely germinated at 30 ºC.
+Above 20 ºC, germination declined more sharply, and meadow seeds rarely germinated at 30 ºC.
 
-Some differences became apparent when dividing the dataset in the major floristic groups that compose mesic grassland vegetation (i.e.
-(i) the Poaceae as dominant and structural species, (ii) the Fabaceae as the second largest group, and (iii) the rest of the families).
+Some differences became apparent when dividing the dataset in the three floristic groups that compose mesic meadow vegetation.
 In the dominant group of the Poaceae (**Fig.
@@
 Fabaceae seed germination also responded negatively to alternating temperatures.
-Finally, the remaining families (**Fig.
-3**), the main cues having a positive effect on germination were stratification, GA~3~.
-alternating temperatures and light.
+Finally, in the remaining families (**Fig.
+3**), the main cues having a positive effect on germination were stratification, GA~3~, alternating temperatures and light.
 These species did not respond to scarification nor to average temperature.
@@
 
-## Seed germination spectrum
+## Seed traits and ecological preferences
 
-The first PCA axis explained `r round(pca$eig[1, 2], 0)`% of the variation (**Fig.
+PCA indicated a clear separation between environmental preferences and seed traits, with each set of variables contributing to different axes (**Fig.
 4**).
-The variables with the largest contribution to this first axis were the germination responses to scarification and alternating temperatures, and seed mass.
-This horizontal axis separated (i, left) species of the Poaceae and the minor families which responded to alternating temperatures and had smaller seeds from (ii, right) species of the Fabaceae which responded to scarification and had heavier seeds.
-Axis 2 explained `r round(pca$eig[2, 2], 0)`% of the variability.
-The main contributing variables were the germination responses to light and stratification.
-This axis separated (iii, bottom) species from all groups that germinated better in light from (iv, top) species of the minor families which needed stratification to germinate and had higher germination in the dark.
+The first PCA axis explained `r round(pca$eig[1, 2], 0)`% of the variation and was related to environmental preferences.
+The variables with the largest contribution to this first axis were soil reaction (pH), winter cold (bio06) and summer drought (bio14).
+This horizontal axis separated (i, left) species with preferences for sites with warm winter temperatures from (ii, right) species with preferences for sites with high summer rainfall and alkaline soils.
+Axis 2 explained `r round(pca$eig[2, 2], 0)`% of the variability and related to seed traits.
+The main contributing variables were seed mass, seed number and the germination response to scarification.
+This axis separated (iii, bottom) species of the Poaceae that produced many seeds with a positive germination response to alternating temperatures from (iv, top) species of the Fabaceae that produced heavy seeds with a positive response to scarification.
+
+## Comparison with other habitats
+
+The second ordination resulted in a separation of mesic meadow species from species belonging to natural herbaceous communities of the same region (**Fig.
+5**).
+Axis 1 explained 78% of the variance and was positively associated to germination at all three temperature treatments.
+This horizontal axis separated (i, left) species with low germination across treatments from (ii, right) species with high germinability across treatments.
+The second axis explained 17% of the variation and separated (iii, bottom) species that responded more to the cool temperature germination treatment from (iv, top) species that responded more positively to the warm germination treatment.
+Meadow species tended to be situated to the right in the horizontal axis, indicating high germinability; and at the center of the vertical axis, indicating a neutral response to temperature.
+On the other hand, species from the other communities were positioned at the left of the horizontal axis, indicating a lower germinability; and were more separated along the vertical axis, indicating a preference for either warmer (bogs and fens) or cooler (alpine grasslands, coastal communities) germination treatments.
 
@@
 
-Our meta-analysis of germination records showed that, overall, the environmental factors with the strongest effect to control seed germination in mesic grassland species were alternating temperatures, light and scarification.
+Our meta-analysis of germination records showed that, overall, the regeneration of mesic meadows is characterized by (1) high germinability when compared to other herbaceous plant communities; (2) low variation in relation to natural environmental drivers of mesic meadow diversity; and (3) a strong phylogenetic separation between the two major families, the Poaceae and the Fabaceae.
+
 Alternating temperatures and light are usually considered to be micro-environmental cues that indicate, at a fine scale, the existence of safe sites for regeneration [@RN4719].
@@
 Therefore, the germination response to alternating temperatures and light indicate conditions that are found immediately after mowing for haymaking, a predictable perturbation that occurs approximately at the same time every year, and which forces meadow plants to establish, grow and reproduce in the time frame determined by this perturbation [@RN4887; @RN4888].
-In addition, this traits would also promote the formation of a persistent soil seed bank [@RN4718; @RN4929; @RN1292].
-On the other hand, it is worth mentioning that the Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate grasslands [RN5014; @RN5013].
+In addition, these traits would also promote the formation of a persistent soil seed bank [@RN4718; @RN4929; @RN1292].
+On the other hand, it is worth mentioning that the Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate meadows [@RN5014; @RN5013].
 
 The strong effect of scarification was specifically related to the hard-seeded Fabaceae.
-In this family, physical dormancy as a result of an unpermeable seed coat works as a mechanism to detect seasonal cycles of temperature and humidity [@RN780].
-It has been also proposed that hardseedeness can avoid seed predation, by preventing the emission of olfactory cues that are elicited by seed imbibition and that can be perceived by seed predators [@RN2568].
-In the case of this study, Fabaceae seeds are the heaviest, and they clearly follow a separate regeneration strategy from that of the other families.
+In this family, physical dormancy as a result of an impermeable seed coat works as a mechanism to detect seasonal cycles of temperature and humidity [@RN780].
+It has been also proposed that hardseedness can avoid seed predation, by preventing the emission of olfactory cues that are elicited by seed imbibition and that can be perceived by seed predators [@RN2568].
+In the case of this study, Fabaceae seeds are amongst the heaviest, and they clearly follow a separate regeneration strategy from that of the other families.
 While other groups rely on alternating temperatures and light as gap-detecting mechanisms, Fabaceae seeds do not respond to these factors, and instead they are mainly regulated by scarification, and they also respond to cooler temperatures than the rest of species.
@@
 
-The germination response to average temperatures and stratification is understood to detect cues related to seasonal cycling and macroclimatic variation [@RN3063, @RN5036].
+The germination response to average temperatures and stratification is understood to detect cues related to seasonal cycling and macroclimatic variation [@RN3063; @RN5036].
 Apart from the aforementioned response to cool temperatures in the Fabaceae, the rest of the families did not show a response to average temperatures, further highlighting their reliance on micro-habitat cueing.
@@
 The minor families, however, showed a divergence from the Poaceae strategy in their positive response to stratification.
-These subordinate species appear to require a period of cold stratification, indicating that overwintering has occured, before they can germinate [@RN3214].
+These subordinate species appear to require a period of cold stratification, indicating that overwintering has occurred, before they can germinate [@RN3214].
 These minor families are also the only group that shows a positive response to GA~3~, a phytohormone which can work as a substitute of cold stratification to overcome physiological dormancy [@RN3368].
 
-The positive phylogenetic signal found in the models highlights the phylogenetic clustering of regeneration strategies that we have described so far: (1) the dominant family of the Poaceae, showing a lack of responses to seasonal cues (average temperature and cold stratification), and relying instead on micro-niche cues (alternating temperatures, light) that can be associated to the yearly perturbation of mowing, (2) the second most-dominant family, the Fabaceae, which do not respond to the micro-niche cues, and regulate their germination timing through scarification and cooler germination temperatures, possibly because their larger size and reserves allow their seeds to decouple their emergence timing from the mowing perturbance, and finally (3) the minor families that respond to micro-cues in a similar way than the Poaceae, but which are differentieted from them by showing a positive response to cold stratification and G~3~, indicating that they rely on physiological seed dormancy.
+The positive phylogenetic signal found in the models highlights the phylogenetic clustering of regeneration strategies that we have described so far: (1) the dominant family of the Poaceae, showing a lack of responses to seasonal cues (average temperature and cold stratification), and relying instead on micro-niche cues (alternating temperatures, light) that can be associated to the yearly perturbation of mowing, (2) the second most-dominant family, the Fabaceae, which do not respond to the micro-niche cues, and regulate their germination timing through scarification and cooler germination temperatures, possibly because their larger size and reserves allow their seeds to decouple their emergence timing from the mowing disturbance, and finally (3) the minor families that respond to micro-cues in a similar way than the Poaceae, but which are differentiated from them by showing a positive response to cold stratification and G~3~, indicating that they rely on physiological seed dormancy.
+The phylogenetic clustering is also apparent regarding seed shape, mass and number, with the Poaceae seeds producing many elongated and light seeds, and the Fabaceae producing fewer, rounder and heavier seeds.
+The shape of the Poaceae seeds could make them particularly successful in being dispersed attached to hay, and grass seeds tend to be overrepresented in seed mixes obtained via hay-making [@RN5046; @RN4894].
 
-* Less dormancy in grassland vs.
-arable populations [@RN1229].
-Herbaceous plants adapted to open habitats needed less cold stratification than congeneric species adapted to forest understories [@RN3570].
-
-* Better germination without treatments [@RN4894].
-
-* In Poland soil bank dominated by forbs of arable and weedy species, low representation of grasses and legumes, confirms the typical Arrhenatherion species do not form a seed bank [@RN5032].
-Soil bank is transient in hay meadows and related grasslands [@RN5026; @RN5041; @RN5042].
-Regeneration limited by dispersal limitation [@RN3073].
-
-* Mowing promotes germination and emergence [@RN3073]. Mowing promotes autumn germination [@RN5025].
-
-* Question of clonal regeneration.
-
-* Discuss second axis = geophytes vs hemicriptófitos?
-
-* Conclusions/applications
-
-* Discuss dormancy in freshly collected grasses, limitations of stored seed data [@RN5018]
+As we have seen, this regeneration strategy of high germinability appears to be unrelated to environmental patterns in hay meadows, and seems more adapted to successful regeneration in face of the predictable perturbation that is yearly mowing.
+In an abandoned meadow of the Prealps, resuming mowing promoted germination and emergence, although natural regeneration was limited due to a lack of seed dispersal to the site [@RN3073].
+Mowing also promoted autumn germination in a dry grassland of northern Germany [@RN5025].
+These results agree with the lack of dormancy and the high germinability of meadows seeds that we have found, and which has also been highlighted by previous studies.
+For example, when comparing populations of *Poa trivialis*, grassland seeds were less dormant than seeds collected in an arable field [@RN1229].
+Ten Brink et al.
+[-@RN3570], when comparing congeneric herbaceous species from open and forest habitats found that species from open habitats needed less cold stratification.
+Similarly, seeds collected from a hay meadow germinated better when left untreated, rather than when being exposed to dormancy-breaking treatments [@RN4894].
+These high-germinability strategy has the consequence of greatly limiting the long-term seed bank of mesic meadows.
+Rather, the soil bank is transient in hay meadows and related grasslands [@RN5026; @RN5041; @RN5042].
+In Poland, the soil bank of a hay meadow was dominated by forbs of arable and weedy species, with low representation of grasses and legumes [@RN5032].
+Overall, this indicated that mesic meadows are quite sensitive to land use change, and their inability to form seed banks leads to the known dispersal limitation to natural regeneration [@RN3073], highlighting the importance of active actions of meadow restoration via seed supply.
+Human activities have shaped the regeneration of hay meadows, leading to a loss of seed dormancy and seasonal tracking, as has been found in many domesticated species [@RN4315].
+The same anthropic processes that have shaped semi-natural mesic meadows have left them dependent on continued human intervention for their regeneration.
 
@@
 
-```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Diversity of dispersal units in mesic grasslands: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* subsp.
-*despectus* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadencis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (o) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
-knitr::include_graphics("../results/figures/seeds.png")
+```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Diversity of dispersal units in mesic meadows: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadensis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (o) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
+knitr::include_graphics("../results/figures/Fig1.png")
 ```
 
-```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Seed size and shape in mesic grasslands.
-Kernel probability densities (violin plots) for seed mass, seed length and seed shape (length / width).
-The three horizontal lines within the probability densities represent the first quartile, the median and the third quartile of the data."}
-knitr::include_graphics("../results/figures/violins.png")
+```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Morphology of mesic meadow seeds.
+The two panels on the left show the probability densities of species values for seed mass and seed number.
+The three horizontal lines within the probability densities represent the first quartile, the median and the third quartile of the values.
+The panel on the right shows values of seed length and width obtained by image analysis, with each point being a seed.
+In all cases, data is divided between the grasses (Poaceae), legumes (Fabaceae) and the other botanical families."}
+knitr::include_graphics("../results/figures/Fig2.png")
 ```
 
-```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on seed germination of mesic grassland species.
-Dots indicate the posterior mean of the effect size, and whiskers its 95 % credible interval.
-The line of zero effect is shown.
-When the credible intervals overlap with the zero-effect line, the effect is not significant.
-The figure shows the results of a model including all species, and of models including only the three main floristic groups of mesic grasslands (i.e.
-the Poacaea, the Fabaceae and the remaining families)."}
-knitr::include_graphics("../results/figures/mcmc.png")
+```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Germination cues in mesic meadows.
+Effect of germination cues simulated in the laboratory over the final germination proportions of mesic meadow seeds.
+Dots indicate the posterior mean of the effect size for each cue, and whiskers the 95% credible interval of the effect size.
+The line of zero-effect is shown: when a credible interval overlaps with the zero-effect line, the effect can be regarded as non-significant.
+In separate panels, the figure shows the results of a general model including data for all species, plus specific models for the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families."}
+knitr::include_graphics("../results/figures/Fig3.png")
 ```
 
-```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Principal Component Analysis ordination of the seed germination spectrum of mesic grassland species.
-Each symbol is a species, different symbols indicate the three major floristic groups of mesic grasslands Labels indicate the contribution of the variables to the axes."}
-knitr::include_graphics("../results/figures/pca.png")
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Environment and seed traits are separate axis of variation in mesic meadows.
+Principal Component Analysis ordination of mesic meadow species considering their environmental preferences and their seed traits.
+Each point is a species, colored by the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families.
+Labels indicate the contribution of the variables to the axes: grey-background labels for environmental preferences, and white-background labels for seed traits.
+Environmental preferences were calculated as species niche centroids (SNCs) for the minimal temperature of the coldest month (bio06), precipitation of the driest month (bio14), and soil pH.
+To calculate the SNCs, a vegetation dataset of mesic meadows of the Iberian Peninsula was used.
+The seed traits are seed mass and the germination relative indices for the response to scarification, stratification, average germination temperature (temperature), alternating temperature (alternating) and light.
+All environmental preferences were aligned to the first axis, while the seed traits were aligned to the second axis, showing that environment and seed traits are separate axis of variation in mesic meadows."}
+knitr::include_graphics("../results/figures/Fig4.png")
 ```
+
+```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Higher germinability in mesic meadows compared to other herbaceous plant communities.
+Principal Component Analysis ordination of species responses to three seed germination treatments.
+Species are grouped by their habitat, with points showing the centroid for each habitat group, and rays linking the centroid with the position of each species in the group.
+Labels indicate the contribution of the germination treatments to the axes.
+All species had been collected in herbaceous vegetation types of the Cantabrian Mountains of Spain and the neighboring coast.
+All seeds were recently collected and untreated, except for scarification, which was applied routinely to all botanical families presenting it.
+Germination treatments consisted in 12/12 h periods with a warmer phase in light and a cooler phase in darkness.
+The position of mesic meadows compared to the other habitats indicates that their species tended to have higher germinability even if untreated (i.e.
+less seed dormancy) and were more neutral in their thermal requirements."}
+knitr::include_graphics("../results/figures/Fig5.png")
+```

### 2021-05-27 src/cleaning.R

@@ -52,16 +52,14 @@ read.xlsx("../#data/plots/data/sospraderas/Analisis_Suelos_SOSPraderas_16-02-201
   merge(read.csv("../#data/plots/data/sospraderas/plotIDs.csv"), by = "X2") %>% # This is a file I had to do manually to fix the mess of IDs in the extra soil data file
   dplyr::select_if(~ any(is.numeric(.))) %>% # Remove non-numeric variables
   rename(Plot = Cod_Sitio,
-         pH = `pH.H2O.(1:2.5)`,
-         Fertility = `CICE.(cmol+/kg)`,
-         Sand = `Arena.(%)`) %>%
-  select(Plot, pH, Sand) -> soils
+         pH = `pH.H2O.(1:2.5)`) %>%
+  select(Plot, pH) -> soils
 
 # Get CHELSA bioclimatic variables
 # CHELSA rasters are stored outside project folder, must download from https://chelsa-climate.org/bioclim/
 
-bio06 <- raster::raster("../#data/maps/CHELSA/CHELSA_bio10_06.tif") # Min temperature coldest month
-bio14 <- raster::raster("../#data/maps/CHELSA/CHELSA_bio10_14.tif") # Precipitation driest month
+bio06 <- raster::raster("../#data/maps/CHELSA/CHELSA_bio10_06.tif") # Min coldest month
+bio14 <- raster::raster("../#data/maps/CHELSA/CHELSA_bio10_14.tif") # P driest month
 
 pts <- header0
 sp::coordinates(pts) <- ~Longitude+Latitude # Make points a spatial object
@@ -83,8 +81,8 @@ rm(header0, Chelsa, soils, bio06, bio14, pts)
 
 header %>%
   merge(species) %>%
-  select(Species, bio06:Sand, Cover) %>%
-  gather(Trait, Value, bio06:Sand) %>%
+  select(Species, bio06:pH, Cover) %>%
+  gather(Trait, Value, bio06:pH) %>%
   group_by(Species, Trait) %>%
   summarise(Value = weighted.mean(Value, Cover)) %>%
   spread(Trait, Value) %>%
@@ -102,9 +100,7 @@ species %>%
   summarise(Frequency = length(Species), Abundance = sum(Cover)) %>% # Relative abundance
   mutate(Abundance = scales::rescale(Abundance, to = c(1, 100))) %>% # Cumulative relative abundance (rescaled 1-100)
   mutate(Frequency = 100 * Frequency / length(unique(species$Plot))) %>%
[…]

### 2021-05-27 src/figures.R

@@ -1,28 +1,24 @@
 library(tidyverse)
 
-# Figure 1 - Seed morphology
+# Figure Seed morphology
 
-## Plot figure
+## Plot figure A
 
 read.csv("data/traits.csv") %>%
   select(Taxon, Family,
-         Seed.mass, Length, Width) %>%
-  na.omit %>%
+         Seed.mass, Seed.number) %>%
   mutate(Family = ifelse(Family %in% c("Leguminosae", "Poaceae"), Family, "Others"),
-         Family = ifelse(Family %in% "Leguminosae", "Fabaceae", Family),
-         Shape = Length / Width) %>%
-  select(-Width) %>%
-  select(-Length) %>%
-  gather(Trait, Value, Seed.mass:Shape) %>%
+         Family = ifelse(Family %in% "Leguminosae", "Fabaceae", Family)) %>%
+  gather(Trait, Value, Seed.mass:Seed.number) %>%
   mutate(Family = fct_relevel(Family, c("Poaceae", "Fabaceae", "Others"))) %>%
-  mutate(Trait = fct_relevel(Trait, c("Seed.mass", "Length", "Shape"))) %>%
+  mutate(Trait = fct_relevel(Trait, c("Seed.mass", "Seed.number"))) %>%
   mutate(Trait = fct_recode(Trait, 
                             "Seed mass (log)" = "Seed.mass",
-                            "Seed length (log)" = "Length",
-                            "Seed shape (length / width, log)" = "Shape")) %>%
-  ggplot(aes(Family, log(Value), color = Family, fill = Family)) + 
+                            "Seed number (log)" = "Seed.number")) %>%
+  mutate(Value = ifelse(! Trait %in% c("Seed roundness"), log(Value), Value)) %>%
+  ggplot(aes(Family, Value, color = Family, fill = Family)) + 
   geom_violin(alpha = 0.5, draw_quantiles = c(0.25, 0.5, 0.75)) +
-  facet_wrap( ~ Trait, strip.position = "left", scales = "free_y") +
+  facet_wrap( ~ Trait, strip.position = "left", scales = "free_y", ncol = 1) +
   scale_y_continuous(labels = scales::number_format(accuracy = .1)) +
   ggthemes::theme_tufte() +
   theme(panel.background = element_rect(color = "grey96", fill = "grey96"),
@@ -30,16 +26,64 @@ read.csv("data/traits.csv") %>%
[…]

### 2021-05-27 src/numbers.R

@@ -23,24 +23,6 @@ germination %>% filter(Scarification == 1) %>% pull(Taxon) %>% length -> MSscarY
 germination %>% filter(GA3 == 0) %>% pull(Taxon) %>% length -> MSgaN
 germination %>% filter(GA3 == 1) %>% pull(Taxon) %>% length -> MSgaY
 
-# Seed morpho
-
-traits %>%
-  group_by(Family) %>%
-  summarise(Min = min(Seed.mass, na.rm = TRUE), 
-            Max = max(Seed.mass, na.rm = TRUE),
-            Length = mean(Length, na.rm = TRUE), 
-            Width = mean(Width, na.rm = TRUE)) -> morpho
-
-morpho %>% filter(Family == "Poaceae") %>% pull(Min) %>% round(1) -> MSpoaMin
-morpho %>% filter(Family == "Poaceae") %>% pull(Max) %>% round(1) -> MSpoaMax
-morpho %>% filter(Family == "Poaceae") %>% pull(Length) %>% round(1) -> MSpoaLength
-morpho %>% filter(Family == "Poaceae") %>% pull(Width) %>% round(1) -> MSpoaWidth
-morpho %>% filter(Family == "Leguminosae") %>% pull(Min) %>% round(1) -> MSlegMin
-morpho %>% filter(Family == "Leguminosae") %>% pull(Max) %>% round(1) -> MSlegMax
-morpho %>% filter(Family == "Leguminosae") %>% pull(Length) %>% round(1) -> MSlegLength
-morpho %>% filter(Family == "Leguminosae") %>% pull(Width) %>% round(1) -> MSlegWidth
-
 # Calculate lambda http://www.mpcm-evolution.com/practice/online-practical-material-chapter-11/chapter-11-1-simple-model-mcmcglmm
 
 lambda <- allspp$VCV[,"animal"]/(allspp$VCV[,"animal"] + allspp$VCV[,"units"])
@@ -85,3 +67,5 @@ traits %>%
   FactoMineR::PCA(graph = FALSE) -> pca
 
 pca$var$contrib
+
+

### 2021-06-30 doc/manuscript.Rmd
pc mig

@@
 
-European mesic meadows are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698].
+The mesic meadows of Europe are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698].
 These plant communities developed in clearings of a temperate woodland landscape as a consequence of anthropogenic intervention [@RN5033; @RN5045].
-At a time, they were novel plant communities with a composite flora made up of species from different origins: grasslands plants that existed in non-forested "islands" of the landscape, plus plants that evolved in parallel to human intervention by hybridization or polyploidization [@RN5033].
-During historical times the botanical composition of mesic meadows evolved following changes in human activities and management practices [@RN5035; @RN5045].
-They are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034].
+At a time they were novel plant communities with a composite flora made up of species from different ecological backgrounds, and during historical times their species composition evolved following changes in human activities and management practices [@RN5035; @RN5045].
+Mesic meadows are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034].
 Nowadays, mesic meadows maintained by agricultural practices are an essential aspect of European landscapes, both real [@RN4996] and imagined [@RN5029].
 
-Traditional phytosociological classification of mesic grasslands puts emphasis on the separation between meadows used for hay making versus pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893].
+Traditional classification of mesic grasslands puts emphasis on the separation between meadows used for hay making versus pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893].
 Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic meadows in large parts of Europe [@RN4997].
@@
 
-Mesic meadows are dominated by mesophilous grasses of the Poaceae family, which make up most of the biomass and define the structure of the vegetation.
-The dispersal unit in the grasses is generally the floret (**Fig.
+Mesic meadows are dominated by mesophilous grasses of the family Poaceae, which make up most of the biomass and define the structure of the vegetation.
+The dispersal unit in Poaceae is generally the floret (**Fig.
 1a**), a composite structure made up of the caryopsis or grain (a dry, indehiscent and monospermic fruit) and its surrounding bracts (i.e.
@@
 1b**) that are assumed to aid dispersal, although it is not always the case that they do [@RN1290].
-Meadow grass seeds have been reported as having physiological seed dormancy [@RN5018; @RN3214; @RN5024] in various degrees, but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273].
+Meadow Poaceae seeds have been reported as having physiological seed dormancy in various degrees [@RN5018; @RN3214; @RN5024], but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273].
 Freshly harvested seeds are comparatively more dormant, but dormancy tends to disappear quickly in dry storage [@RN5024; @RN5018].
@@
 In *Poa trivialis*, germination was promoted by light but not by alternating temperatures [@RN1205], although alternating temperatures did encourage some germination in darkness [@RN1229].
-Wild grass ecotypes usually have a germination response to alternating temperatures [@RN1248; @RN1261; @RN1230; @RN1221; @RN1290; @RN1172; @RN1292; @RN1280], although this response is missing in some wild ecotypes and in the domesticated cultivars [@RN1261], and some species such as *Lolium perenne* have been reported as insensitive to temperature alternation [@RN1280; @RN1380].
+Wild Poaceae ecotypes usually have a germination response to alternating temperatures [@RN1248; @RN1261; @RN1230; @RN1221; @RN1290; @RN1172; @RN1292; @RN1280], although this response is missing in some wild ecotypes and in the domesticated cultivars [@RN1261], and some species such as *Lolium perenne* have been reported as insensitive to temperature alternation [@RN1280; @RN1380].
 
-Next in abundance to the grasses are the legumes of the Fabaceae family.
-Legumes contribute to the nutritional value of meadow fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040].
-The dispersal unit of most legumes is the seed itself (**Fig.
+Next in abundance to Poaceae are the legumes of the family Fabaceae.
+Fabaceae contribute to the nutritional value of meadow fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040].
+The dispersal unit of most Fabaceae is the seed itself (**Fig.
 1c**) but in some species dispersal units are more complex, including indehiscent monospermic fruits (e.g.
@@
 1e**).
-Legume seeds are generally hard-seeded, i.e.
+Fabaceae are generally hard-seeded, i.e.
 they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e.
@@
 In *Medicago*, seeds that have not reached full maturity can germinate before they become impermeable, but the completion of maturation imposes coat impermeability, and thereafter the seed must be scarified to allow water imbibition and germination [@RN652].
-Buried legume seeds are able to track the seasons, and in some species germination seems to be promoted by cold stratification and alternating temperatures [@RN780]; some of these species have been described as having combinational dormancy (i.e.
+Buried Fabaceae seeds are able to track the seasons, and in some species germination seems to be promoted by cold stratification and alternating temperatures [@RN780]; some of these species have been described as having combinational dormancy (i.e.
 physical + physiological) [@RN5037].
-However, seeds have also been reported to germinate without any previous treatment [@RN881; @RN632; @RN460; @RN372], and to lose dormancy during storage [@RN5037].
-As in the grasses, legume seeds have been reported to germinate in high numbers across a range of temperatures from 5 to 25 ºC [@RN652; @RN3273].
-Legume seeds have been described as not responsive to light and capable of germinating in darkness [@RN1339; @RN3273].
+However, Fabaceae seeds have also been reported to germinate without any previous treatment [@RN881; @RN632; @RN460; @RN372], and to lose dormancy during storage [@RN5037].
+As in Poaceae, seeds of Fabaceae have been reported to germinate in high numbers across a range of temperatures from 5 to 25 ºC [@RN652; @RN3273].
+Fabaceae seeds have been described as not responsive to light and capable of germinating in darkness [@RN1339; @RN3273].
 
-Grasses and legumes are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the aesthetic value of the meadows [@RN5038; @RN5039].
-They also add nutritional scope, being richer than grasses and legumes in specific elements [@RN5028; @RN5040].
-In many of the most frequent minor families (e.g.
-in the Asteraceae, Ranunculaceae, Polygonaceae, Lamiaceae, Dipsacaceae, Cyperaceae) the dispersal unit is the achene: dry, indehiscent and monospermic fruits.
+Poaceae and Fabaceae are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the aesthetic value of the meadows [@RN5038; @RN5039].
+They also add nutritional scope, being richer than Poaceae and Fabaceae in specific elements [@RN5028; @RN5040].
+In many of these families (e.g.
+Asteraceae, Cyperaceae, Dipsacaceae, Lamiaceae, Polygonaceae, Ranunculaceae) the dispersal unit is the achene: dry, indehiscent and monospermic fruits.
 The morphology of these achenes is varied: cypselae with a hairy pappus in *Centaurea* (Asteraceae, **Fig.
@@
 1k**).
-In the Apiaceae, it is the mericarp (**Fig.
+In Apiaceae, it is the mericarp (**Fig.
 1l**), an indehiscent monospermic fragment of the fruit.
-In some other minor families the dispersal unit is the seed itself, such as in the Caryophyllaceae (**Fig.
-1m**), the Plantaginaceae (**Fig.
-1n**), the Juncaceae or the hemiparasitic species of *Rhinanthus* (Orobanchaceae) (**Fig.
+In some other minor families the dispersal unit is the seed itself, such as in Caryophyllaceae (**Fig.
+1m**), Plantaginaceae (**Fig.
+1n**), Juncaceae or the hemiparasitic species of *Rhinanthus* (Orobanchaceae) (**Fig.
 1o**).
-In the Asteraceae, high germination without pretreatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; @RN120; @RN963; @RN1428], *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403].
+In Asteraceae, high germination without pretreatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; @RN120; @RN963; @RN1428], *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403].
 *Taraxacum officinale* germinates between 5 and 30 ºC [@RN120; RN1428; @RN1276], and has higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428].
@@
 Underdeveloped embryos that need to grow inside the seed before germination (i.e.
-morphological dormancy) are widespread in the Ranunculaceae and the Apiaceae [@RN3214; @RN1112].
+morphological dormancy) are widespread in Ranunculaceae and Apiaceae [@RN3214; @RN1112].
 *Ranunculus repens* (Ranunculaceae) germinates between 10 and 25 ºC, but the germination percentages were reported to be low [@RN931]; the same species has been reported to respond to alternating temperatures, which can promote its germination even in darkness [@RN3180].
-In the Polygonaceae, *Rumex acetosa* can germinate immediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not, this difference is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814].
+In Polygonaceae, *Rumex acetosa* can germinate immediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not, this difference is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814].
 In *Heracleum sphondylium* (Apiaceae), growth of the embryo only occurs below 10 ºC, in moist conditions [@RN1112].
@@
 
-Although a wealth of studies has accumulated, a synthesis of the seed ecology of mesic meadows is missing, despite their conservation, restoration and cultural significance.
-In this article, we present for the first time a synthesis of regeneration by seed in European mesic meadows.
-To do so, we combine newly-generated data on seed morphology and germination with records from databases [@RN5036; @RN4968; @RN2987; @RN2319].
-To the resulting dataset containing `r MSrecords` germination records from `r MScountries` European countries we apply Bayesian meta-analysis [@RN5000] to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light.
-We analyze the relationship between seed traits and species environmental preferences, and compare the germination ecology of mesic meadows with that of other herbaceous plant communities.
+Although a wealth of studies has accumulated, a synthesis of seed ecology in European mesic meadows is missing.
+In this article, we review for the first time this topic, combining newly-generated data on seed morphology and germination with records from existing databases [@RN5036; @RN4968; @RN2987; @RN2319].
+The resulting dataset contains `r MSrecords` germination records of `r MSspeciesN` plant species from `r MScountries` European countries.
+To test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light we apply Bayesian meta-analysis [@RN5000].
+Further, using well-preserved meadows of the Iberian Peninsula as a study system, we analyze the covariation between seed traits and species environmental preferences, and compare the germination ecology of mesic meadows with that of other herbaceous plant communities from the same area.
 
@@
 
-Sampled meadows had been maintained by traditional agricultural practices: mowing for hay-making once or twice per year plus light grazing.
-Each sampling plot was placed in a square area (25-100 m^2^ area) situated in the central part of a meadow, avoiding the margins.
-Sampling took place in 2016-2017, at the peak of plant development, just before mowing.
+Study meadows had been maintained by traditional agricultural practices: mowing for hay-making once or twice per year plus light grazing.
+Each vegetation sampling plot was placed in a square area (25-100 m^2^ area) situated in the central part of a meadow, avoiding the margins.
+Vegetation sampling took place in 2016-2017, at the peak of plant development, just before mowing.
 All vascular plant species in the plots were recorded and given a cover value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%).
 All plant names were assigned following the nomenclature of Euro+Med [-@RN5044], which is used throughout this article.
-As expected, the vegetation of the sampled meadows was dominated by the Poaceae and the Fabaceae: these two families represented 47% and 17%, respectively, of the total plant cover recorded in all the plots.
-Other 20 families were recorded, the largest of which was the Asteraceae; each of these other families represented less than 10% of the total cover.
-These numbers refer only to the seed plants, pteridophytes being a marginal part of the records.
+As expected, the vegetation of the sampled meadows was dominated by Poaceae and Fabaceae: these two families represented 47% and 17%, respectively, of the total plant cover recorded in all the plots.
+Other 20 families were recorded, the largest of which was Asteraceae; each of these other families represented less than 10% of the total cover.
 
@@
 Finally, we rescaled the values of all species to a 1-100 scale to obtain the cumulative cover values.
-We removed 208 species with cumulative cover values below 2%, considering them to be transient species that might have been recorded by chance and may not represent the core mesic meadows flora.
+We removed 208 species with cumulative cover values below 2%, considering them to be rare species that might have been recorded by chance and may not represent the core mesic meadows flora.
 We used the remaining 116 species as the core list to retrieve seed germination data for this article.
 The majority of these species were hemicryptophytes (78%), with some therophytes (16%) and a few chamaephytes and geophytes (3% of each).
-The family with more species in the list was the Poaceae (22%), with another 17% belonging to the Fabaceae, 15% to the Asteraceae, 8% to the Apiaceae, and the rest of the families representing less than 5% each.
+The family with more species in the list was Poaceae (22%), with another 17% belonging to Fabaceae, 15% to Asteraceae, 8% to Apiaceae, and the rest of the families representing less than 5% each.
 
@@
 For cold and drought, we used the coordinates of the plots to retrieve from CHELSA [@RN4766] the bioclimatic variables bio06 (minimum temperature of the coldest month) and bio14 (precipitation of the driest month).
-and represents the limiting water availability along the Iberian gradient of summer drought.
 For soil reaction, we took from each plot five soil samples from between 0 and 20 cm depth with a Dutch auger and combined them to make a bulk soil horizon, which we subsequently air-dried, crumbled, finely crushed and sieved with a 2 mm screen, to finally measure the pH in H~2~O with a glass electrode in a suspension of soil and water (1:2.5).
@@
 Additionally, we compared, at each of these temperature regimes, the germination of fresh seeds versus seeds subjected to a dormancy-breaking treatment.
-In the case of the Fabaceae and other families that might present physical dormancy [@RN3214], the treatment consisted in scarification by chipping the seed coat with a scalpel.
-For the rest of the families, we used gibberellic acid GA~3~ (0,0645 mM) in darkness during 24h, as a treatment to remove potential physiological seed dormancy [@RN4976].
+In the case of Fabaceae and other families that might present physical dormancy [@RN3214], the treatment consisted in scarification by chipping the seed coat with a scalpel.
+For the rest of the families, we used gibberellic acid GA~3~ (0.0645 mM) in darkness during 24h, as a treatment to remove potential physiological seed dormancy [@RN4976].
 For each species and treatment, we sowed four Petri dishes with 25 seeds each.
@@
 
-The Poaceae had lower values of seed mass and higher values of seed number, while the Fabaceae had heavier but fewer seeds (**Fig.
+Poaceae had lower values of seed mass and higher values of seed number, while Fabaceae had heavier but fewer seeds (**Fig.
 2**).
-The other families covered the range of values showed by the Poaceae and the Fabaceae, but their median values were high for both traits: their median seed mass was close to that of the Fabaceae, while their median seed number was higher than that of the Poaceae (**Fig.
+The other families covered the range of values showed by Poaceae and Fabaceae, but their median values were high for both traits: their median seed mass was close to that of the Fabaceae, while their median seed number was higher than that of the Poaceae (**Fig.
 2**).
-Seed shape also showed a divergence between the Poaceae and Fabaceae, with seeds being elongated in the former and rounder in the latter (**Fig.
+Seed shape also showed a divergence between Poaceae and Fabaceae, with seeds being elongated in the former and rounder in the latter (**Fig.
 2**).
@@
 Some differences became apparent when dividing the dataset in the three floristic groups that compose mesic meadow vegetation.
-In the dominant group of the Poaceae (**Fig.
+In Poaceae (**Fig.
 3**), no effect was found for stratification, GA~3~ or temperature.
@@
 The major drivers of Poaceae germination appeared to be alternating temperatures and light, with both having a positive effect.
-In the group of the Fabaceae (**Fig.
+In Fabaceae (**Fig.
 3**), the largest positive effect on germination was produced by scarification, with no effect of stratification, GA~3~ or light.
@@
 This horizontal axis separated (i, left) species with preferences for sites with warm winter temperatures from (ii, right) species with preferences for sites with high summer rainfall and alkaline soils.
-Axis 2 explained `r round(pca$eig[2, 2], 0)`% of the variability and related to seed traits.
+Axis 2 explained `r round(pca$eig[2, 2], 0)`% of the variability, mostly related to seed traits.
 The main contributing variables were seed mass, seed number and the germination response to scarification.
-This axis separated (iii, bottom) species of the Poaceae that produced many seeds with a positive germination response to alternating temperatures from (iv, top) species of the Fabaceae that produced heavy seeds with a positive response to scarification.
+This axis separated (iii, bottom) Poaceae species that produced many seeds with a positive germination response to alternating temperatures from (iv, top) Fabaceae species that produced heavier seeds with a positive response to scarification.
 
@@
 
-The second ordination resulted in a separation of mesic meadow species from species belonging to natural herbaceous communities of the same region (**Fig.
+Ordination resulted in a separation of mesic meadow species from species belonging to natural herbaceous communities of the same region (**Fig.
 5**).
-Axis 1 explained 78% of the variance and was positively associated to germination at all three temperature treatments.
+Axis 1 explained 78% of the variance and was positively associated to high germination at all three temperature treatments.
 This horizontal axis separated (i, left) species with low germination across treatments from (ii, right) species with high germinability across treatments.
 The second axis explained 17% of the variation and separated (iii, bottom) species that responded more to the cool temperature germination treatment from (iv, top) species that responded more positively to the warm germination treatment.
-Meadow species tended to be situated to the right in the horizontal axis, indicating high germinability; and at the center of the vertical axis, indicating a neutral response to temperature.
+Meadow species tended to be situated in the right sector of the ordination, indicating high germinability; and at the center of the vertical axis, indicating a neutral response to temperature.
 On the other hand, species from the other communities were positioned at the left of the horizontal axis, indicating a lower germinability; and were more separated along the vertical axis, indicating a preference for either warmer (bogs and fens) or cooler (alpine grasslands, coastal communities) germination treatments.
@@
 
-Our meta-analysis of germination records showed that, overall, the regeneration of mesic meadows is characterized by (1) high germinability when compared to other herbaceous plant communities; (2) low variation in relation to natural environmental drivers of mesic meadow diversity; and (3) a strong phylogenetic separation between the two major families, the Poaceae and the Fabaceae.
+Our meta-analysis of germination records showed that, overall, the regeneration of mesic meadows is characterized by (1) high germinability when compared to other herbaceous plant communities; (2) low correspondence between seed trait variability and the natural environmental drivers of mesic meadow diversity; and (3) a strong phylogenetic separation between the two major families, Poaceae and Fabaceae.
 
+We found that alternating temperatures and light are the two most influential factors triggering germination, especially in species from the family Poaceae and the remaining ones, with the exception of Fabaceae.
 Alternating temperatures and light are usually considered to be micro-environmental cues that indicate, at a fine scale, the existence of safe sites for regeneration [@RN4719].
@@
 In addition, these traits would also promote the formation of a persistent soil seed bank [@RN4718; @RN4929; @RN1292].
-On the other hand, it is worth mentioning that the Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate meadows [@RN5014; @RN5013].
+On the other hand, it is worth mentioning that Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate meadows [@RN5014; @RN5013].
 
@@
 The germination response to average temperatures and stratification is understood to detect cues related to seasonal cycling and macroclimatic variation [@RN3063; @RN5036].
-Apart from the aforementioned response to cool temperatures in the Fabaceae, the rest of the families did not show a response to average temperatures, further highlighting their reliance on micro-habitat cueing.
-Moreover, the dominant family of the Poaceae showed a negative response to cold stratification.
-The minor families, however, showed a divergence from the Poaceae strategy in their positive response to stratification.
-These subordinate species appear to require a period of cold stratification, indicating that overwintering has occurred, before they can germinate [@RN3214].
-These minor families are also the only group that shows a positive response to GA~3~, a phytohormone which can work as a substitute of cold stratification to overcome physiological dormancy [@RN3368].
+Apart from the aforementioned response to cool temperatures in Fabaceae, the rest of the families did not show a response to average temperatures, further highlighting their reliance on micro-habitat cueing.
+Moreover, the dominant family Poaceae showed a negative response to cold stratification.
+The less frequent families, however, showed a divergence from the Poaceae strategy in their positive response to stratification.
+These non-dominant species appear to require a period of cold stratification, indicating that overwintering has occurred, before they can germinate [@RN3214].
+These minor families are also the only group of species that showed a positive response to GA~3~, a phytohormone which can work as a substitute of cold stratification to overcome physiological dormancy [@RN3368].
 
-The positive phylogenetic signal found in the models highlights the phylogenetic clustering of regeneration strategies that we have described so far: (1) the dominant family of the Poaceae, showing a lack of responses to seasonal cues (average temperature and cold stratification), and relying instead on micro-niche cues (alternating temperatures, light) that can be associated to the yearly perturbation of mowing, (2) the second most-dominant family, the Fabaceae, which do not respond to the micro-niche cues, and regulate their germination timing through scarification and cooler germination temperatures, possibly because their larger size and reserves allow their seeds to decouple their emergence timing from the mowing disturbance, and finally (3) the minor families that respond to micro-cues in a similar way than the Poaceae, but which are differentiated from them by showing a positive response to cold stratification and G~3~, indicating that they rely on physiological seed dormancy.
-The phylogenetic clustering is also apparent regarding seed shape, mass and number, with the Poaceae seeds producing many elongated and light seeds, and the Fabaceae producing fewer, rounder and heavier seeds.
-The shape of the Poaceae seeds could make them particularly successful in being dispersed attached to hay, and grass seeds tend to be overrepresented in seed mixes obtained via hay-making [@RN5046; @RN4894].
+The positive phylogenetic signal found in the models highlights the phylogenetic clustering of regeneration strategies that we have described so far: (1) the dominant family Poaceae, showing a lack of responses to seasonal cues (average temperature and cold stratification), and relying instead on large amounts of propagules and on detecting micro-niche cues (alternating temperatures, light) that can be associated to the yearly perturbation of mowing, (2) the second most-dominant family, Fabaceae, which do not respond to the micro-niche cues, and regulate their germination timing through scarification and cooler germination temperatures, possibly because their larger size and reserves allow their seeds to decouple their emergence timing from the mowing disturbance, and finally (3) the minor families that respond to micro-cues in a similar way than Poaceae, but which are differentiated from them by showing a positive response to cold stratification and G~3~, indicating that they rely on physiological seed dormancy.
+This third strategy would allow to fine-tune their germination with the local environmental conditions and the inter-annual climatic variation.
+Phylogenetic clustering is also apparent regarding seed shape, mass and number, with Poaceae producing many elongated and light seeds, and Fabaceae producing fewer, rounder and heavier seeds.
+The shape of the Poaceae seeds could make them particularly successful in being dispersed attached to hay, and Poaceae seeds tend to be overrepresented in seed mixes obtained via hay-making [@RN5046; @RN4894].
+Asteraceae and Apiaceae, some of the most abundant among the minor families, also tend to have shapes resembling those of Poaceae.
 
-As we have seen, this regeneration strategy of high germinability appears to be unrelated to environmental patterns in hay meadows, and seems more adapted to successful regeneration in face of the predictable perturbation that is yearly mowing.
+As we have seen, this regeneration strategy of high germinability appears to be unrelated to environmental patterns in hay meadows, and seems more adapted to the predictable perturbation that is yearly mowing.
 In an abandoned meadow of the Prealps, resuming mowing promoted germination and emergence, although natural regeneration was limited due to a lack of seed dispersal to the site [@RN3073].
@@
 Rather, the soil bank is transient in hay meadows and related grasslands [@RN5026; @RN5041; @RN5042].
-In Poland, the soil bank of a hay meadow was dominated by forbs of arable and weedy species, with low representation of grasses and legumes [@RN5032].
+In Poland, the soil bank of a hay meadow was dominated by forbs of arable and weedy species, with low representation of Poaceae and Fabaceae [@RN5032].
+
 Overall, this indicated that mesic meadows are quite sensitive to land use change, and their inability to form seed banks leads to the known dispersal limitation to natural regeneration [@RN3073], highlighting the importance of active actions of meadow restoration via seed supply.
@@
 The same anthropic processes that have shaped semi-natural mesic meadows have left them dependent on continued human intervention for their regeneration.
+Understanding the germination requirements of the different plant groups that coexist in mesic meadows can help to manage, conserve and restore their biological diversity.
 
@@
 
-```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Diversity of dispersal units in mesic meadows: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadensis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (o) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
+```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Diversity of dispersal units in plant species from mesic meadows: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadensis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (o) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
 knitr::include_graphics("../results/figures/Fig1.png")
@@
 ```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Morphology of mesic meadow seeds.
-The two panels on the left show the probability densities of species values for seed mass and seed number.
+The two panels on the left show the probability densities of species values for seed mass and seed number, log-transformed for ease of visualization.
 The three horizontal lines within the probability densities represent the first quartile, the median and the third quartile of the values.
 The panel on the right shows values of seed length and width obtained by image analysis, with each point being a seed.
-In all cases, data is divided between the grasses (Poaceae), legumes (Fabaceae) and the other botanical families."}
+In all cases, data is divided between the grasses (Poaceae), legumes (Fabaceae) and the other plant families."}
 knitr::include_graphics("../results/figures/Fig2.png")
@@
 The seed traits are seed mass and the germination relative indices for the response to scarification, stratification, average germination temperature (temperature), alternating temperature (alternating) and light.
-All environmental preferences were aligned to the first axis, while the seed traits were aligned to the second axis, showing that environment and seed traits are separate axis of variation in mesic meadows."}
+All environmental preferences were aligned to the first axis, while the seed traits were aligned to the second axis, showing that environment and seed traits are separate axis of variation in mesic meadows.
+Stratification and average germination temperatire showed very low variation, in accordance with their small effect on Figure 2."}
 knitr::include_graphics("../results/figures/Fig4.png")

### 2021-06-30 src/figures.R

@@ -194,7 +194,8 @@ traits %>%
 
 traits %>%
   select(Taxon, Family) %>%
-  cbind(pca$ind$coord[, 1:2]) -> pcaInds
+  cbind(pca$ind$coord[, 1:2]) %>%
+  filter(Dim.2 > -3) -> pcaInds
 
 pca$var$coord[, 1:2] %>%
   data.frame %>%
@@ -210,7 +211,7 @@ ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
   geom_vline(xintercept = 0, linetype = "dashed") +
   geom_point(aes(color = Family), alpha = 0.8, size = 8) +
   ggthemes::theme_tufte() + 
-  theme(legend.position = "top",
+  theme(legend.position = "none",
         legend.title = element_blank(),
         legend.text = element_text(size = 12, color = "black"),
         panel.background = element_rect(color = "grey96", fill = "grey96"),
@@ -268,6 +269,11 @@ ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
   annotate("text", x = -4.2, y = 2.1, label = "Hardseedness\nHeavy seeds\nFabaceae", color = "grey50") +
   annotate("text", x = -2.4, y = -3, label = "Warm-dry climes\nAcidic soils", color = "grey50") +
   annotate("text", x = 2.4, y = -3, label = "Cool-wet climes\nAlkaline soils", color = "grey50") +
+  
+  annotate("text", x = 4, y = 2.9, label = "Poaceae", color = "yellowgreen", hjust = 1) +
+  annotate("text", x = 4, y = 2.6, label = "Fabaceae", color = "gold", hjust = 1) +
+  annotate("text", x = 4, y = 2.3, label = "Other families", color = "darkorchid", hjust = 1) +
+  
   scale_color_manual(values = c("yellowgreen",  "gold", "darkorchid")) +
   scale_fill_manual(values = c("grey65",  "white"))-> f4;f4
 
@@ -324,7 +330,7 @@ ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
   ggthemes::theme_tufte() + 
   theme(legend.position = "top",
         panel.background = element_rect(color = "grey96", fill = "grey96"),
-        legend.text = element_text(size = 12, color = "black"),
+        legend.text = element_text(size = 11, color = "black"),
         legend.title = element_blank(),
         axis.text = element_text(size = 12, color = "black"),
         axis.title = element_text(size = 12, color = "black")) +

### 2021-07-16 README.Rmd
submitted AoB 🌱

@@
 
-# Seed germination ecology of European mesic grasslands
+# Seed ecology of European mesic meadows: the domestication of plant regeneration strategies
 
-This repository stores all the information related to the manuscript *Seed germination ecology of European mesic grasslands*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
+This repository stores all the information related to the manuscript *Seed ecology of European mesic meadows: the domestication of plant regeneration strategies*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
 

### 2021-07-16 doc/manuscript.Rmd
submitted AoB 🌱

@@
 ---
-title: "Seed ecology of European mesic meadows: the domestication of plant regeneration strategies"
+title: "RESEARCH IN CONTEXT: Seed ecology of European mesic meadows: the domestication of plant regeneration strategies"
 output: 
@@
 bibliography: refs.bib
-csl: journal-of-vegetation-science.csl
+csl: annals-of-botany.csl
 link-citations: yes
@@
 
-Eduardo Fernández-Pascual^1^ ^+^, Madalena Vaz^2^, Beatriz Morais^1^, Ramón Reiné^3^, Joaquín Ascaso^3^, Elías Afif Khouri^1^, Angelino Carta^4^
+Eduardo Fernández-Pascual^1,2^ ^+^, Madalena Vaz^3^, Beatriz Morais^2^, Ramón Reiné^4^, Joaquín Ascaso^4^, Elías Afif Khouri^5^, Angelino Carta^6^
 
-^1^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^2^ Banco Português de Germoplasma Vegetal, Instituto Nacional de Investigação Agrária e Veterinária (INIAV), Braga, Portugal; ^3^ Departamento de Ciencias Agrarias y Medio Natural, Universidad de Zaragoza, Huesca, Spain; ^4^ Dipartimento di Biologia, Botany Unit, University of Pisa, Pisa, Italy
+^1^ IMIB – Biodiversity Research Institute, University of Oviedo, Mieres, Spain; ^2^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^3^ Banco Português de Germoplasma Vegetal, Instituto Nacional de Investigação Agrária e Veterinária (INIAV), Braga, Portugal; ^4^ Departamento de Ciencias Agrarias y Medio Natural, Universidad de Zaragoza, Huesca, Spain; ^5^ Agroforestry Engineering Area, Department of Organisms and Systems Biology, Polytechnic School of Mieres, University of Oviedo, 33600 Mieres, Asturias, Spain; ^6^ Dipartimento di Biologia, Botany Unit, University of Pisa, Pisa, Italy
 
-^+^ Correspondence: Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, C/ Catedrático Rodrigo Uría, 33006 Oviedo/Uviéu, Spain.
-Email: efernandezpascual@gmail.com.
-Telephone: +34985104787.
+Running title: Seed ecology of European mesic meadows
 
-# Abstract
+^+^ Correspondence: IMIB – Biodiversity Research Institute, University of Oviedo, Edificio de Investigación – 5ª planta, C/ Gonzalo Gutiérrez Quirós s/n
+33600 Mieres, Spain. Email: fernandezpeduardo@uniovi.es. Telephone: +34 985103000.
 
-# Keywords
+# ABSTRACT (300 words)
 
-*Arrhenatheretalia*, *Arrhenatherion*, plant regeneration, seed germination, seed morphology, species-rich meadows, hay meadows, mesic grasslands
+* **Background and Aims** European mesic meadows are semi-natural open habitats of high biodiversity and an essential part of European landscapes.
+Their species-rich seed mixes can be a source of plant material for ecological restoration, urban greening and rewilding.
+Here, we make a synthesis of the seed ecology of mesic meadows.
+* **Methods** We combined our own experimental data with data obtained from databases to create a combined dataset containing `r MSrecords` germination records of `r MSspeciesN` plant species from `r MScountries` European countries.
+We performed a Bayesian meta-analysis of this dataset to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light.
+We also used multivariate ordination to check the relationship between seed traits (germination and morphology) and species ecological preferences, and to compare the seed ecology of mesic meadows with that of other herbaceous plant communities from the same area.
+* **Key Results** The seed ecology of mesic meadows is characterized by (1) high seed germinability when compared to other herbaceous plant communities; (2) low correspondence between seed traits and species ecological preferences; and (3) a deep phylogenetic separation between the two major families, Poaceae and Fabaceae.
+Poaceae produce many light seeds which respond to gap-detecting germination cues (alternating temperatures and light); Fabaceae produce fewer heavy seeds, which need scarification to break their physical dormancy.
+* **Conclusions** High germinability of meadow seeds will reduce their capacity to form persistent seed banks, resulting in dispersal limitations to passive regeneration.
+For centuries, human activities have shaped the regeneration of meadows, leading to a loss of seed dormancy and decoupling seeds from seasonal cycles, as has been found in many domesticated species.
+The same anthropic processes that have shaped semi-natural mesic meadows have left them dependent on continued human intervention for their regeneration, highlighting the importance of active restoration via seed supply.
 
-# Introduction
+# KEYWORDS
 
-The mesic meadows of Europe are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698].
-These plant communities developed in clearings of a temperate woodland landscape as a consequence of anthropogenic intervention [@RN5033; @RN5045].
-At a time they were novel plant communities with a composite flora made up of species from different ecological backgrounds, and during historical times their species composition evolved following changes in human activities and management practices [@RN5035; @RN5045].
+*Arrhenatheretalia*, *Arrhenatherion*, Asteraceae, Fabaceae, domestication, hay meadows, mesic grasslands, Poaceae, seed germination, seed morphology, semi-natural, species-rich meadows
+
+# INTRODUCTION
+
+The mesic meadows of Europe are semi-natural open habitats that occupy clearings created by human intervention over a temperate wooded landscape [@RN5033; @RN5045], in sites with moderately fertile and well-drained soils [@RN4698].
+Once, these were novel plant communities, with a composite flora made up of species from different ecological backgrounds; during historical times, their species composition further evolved following changes in human events and management practices [@RN5035; @RN5045].
 Mesic meadows are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034].
@@
 
-Traditional classification of mesic grasslands puts emphasis on the separation between meadows used for hay making versus pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893].
+Traditional classification of mesic grasslands emphasises the separation between meadows used for hay making versus pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893].
 Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic meadows in large parts of Europe [@RN4997].
-For this reason, the European Habitats Directive (92/43/EEC) has included in its lists of habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.
+For this reason, the European Habitats Directive (92/43/EEC) lists as habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.
 
-One of the reasons for the high conservation interest of traditional mesic meadows is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed materials to be used in ecological restoration, rewilding and urban greening [@RN5030; @RN4894; @RN5031].
+One of the reasons for the high conservation interest of traditional mesic meadows is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed material to be used in ecological restoration, rewilding and urban greening [@RN5030; @RN4894; @RN5031].
 However, a lack of knowledge about species germination traits has already been identified as a bottleneck [@RN2241] hampering the development of a competitive native seed industry [@RN4665] in Europe.
@@
 
-Mesic meadows are dominated by mesophilous grasses of the family Poaceae, which make up most of the biomass and define the structure of the vegetation.
+Mesic meadows are dominated by mesophilous grasses of the family Poaceae, which make up most of their biomass and define the structure of the vegetation.
 The dispersal unit in Poaceae is generally the floret (**Fig.
@@
 Removal of the lemma and palea, as well as puncturing the pericarp, can aid in seed germination [@RN1248].
-The dispersal unit is often accompanied by awns or hairs (**Fig.
+The dispersal unit often carries awns or hairs (**Fig.
 1b**) that are assumed to aid dispersal, although it is not always the case that they do [@RN1290].
@@
 Germination has been reported to occur at temperatures ranging from 5 to 30 ºC [@RN3273; @RN1229; @RN1261; @RN1230; @RN1292; @RN5024].
-Most Poaceae species have also been reported to germinate better in light than in darkness [@RN1271; @RN1248; @RN1230; @RN1221; @RN1172; @RN1292; @RN1280; @RN5024] and even to be inhibited to germinate by darkness [@RN1229].
+Most Poaceae species have also been reported to germinate better in light than in darkness [@RN1271; @RN1248; @RN1230; @RN1221; @RN1172; @RN1292; @RN1280; @RN5024] and even to be unable to germinate in darkness [@RN1229].
 Nonetheless, species of *Bromus* have been reported as germinating better in darkness [@RN1172] and *Cynosurus cristatus* as being indifferent to light/darkness [@RN1292; @RN1280].
-In *Poa trivialis*, germination was promoted by light but not by alternating temperatures [@RN1205], although alternating temperatures did encourage some germination in darkness [@RN1229].
+In *Poa trivialis*, germination is promoted by light but not by diurnal alternating temperatures (i.e.
+germination conditions where different temperatures are applied during the day and the night, in diurnal cycles) [@RN1205], although alternating temperatures do encourage some germination in darkness [@RN1229].
 Wild Poaceae ecotypes usually have a germination response to alternating temperatures [@RN1248; @RN1261; @RN1230; @RN1221; @RN1290; @RN1172; @RN1292; @RN1280], although this response is missing in some wild ecotypes and in the domesticated cultivars [@RN1261], and some species such as *Lolium perenne* have been reported as insensitive to temperature alternation [@RN1280; @RN1380].
@@
 Next in abundance to Poaceae are the legumes of the family Fabaceae.
-Fabaceae contribute to the nutritional value of meadow fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040].
+Fabaceae species contribute to the nutritional value of meadow fodder, as thanks to their N-fixating capabilities they have high N contents [@RN5028; @RN5040].
 The dispersal unit of most Fabaceae is the seed itself (**Fig.
@@
 1e**).
-Fabaceae are generally hard-seeded, i.e.
-they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e.
+Fabaceae are generally hard-seeded: they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e.
 physical seed dormancy) [@RN3214; @RN3273; @RN1215; @RN5016; @RN3073].
 In *Medicago*, seeds that have not reached full maturity can germinate before they become impermeable, but the completion of maturation imposes coat impermeability, and thereafter the seed must be scarified to allow water imbibition and germination [@RN652].
-Buried Fabaceae seeds are able to track the seasons, and in some species germination seems to be promoted by cold stratification and alternating temperatures [@RN780]; some of these species have been described as having combinational dormancy (i.e.
+Buried Fabaceae seeds can track the seasons, and in some species, germination seems to be promoted by cold stratification and alternating temperatures [@RN780]; some of these species have been described as having combinational dormancy (i.e.
 physical + physiological) [@RN5037].
@@
 
-Poaceae and Fabaceae are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the aesthetic value of the meadows [@RN5038; @RN5039].
+Poaceae and Fabaceae are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the high biodiversity of the system and increase the aesthetic value of the meadows as perceived by people [@RN5038; @RN5039].
 They also add nutritional scope, being richer than Poaceae and Fabaceae in specific elements [@RN5028; @RN5040].
@@
 1g**); hardened nutlets in *Prunella* (Lamiaceae, **Fig.
-1h**); hairy in *Knautia* (Dipsacaceae, **Fig.
+1h**); hairy and with an elaiosome in *Knautia* (Dipsacaceae, **Fig.
 1i**); surrounded by a perigynium which aids in dispersal by water in *Carex* (Cyperaceae, **Fig.
@@
 1l**), an indehiscent monospermic fragment of the fruit.
-In some other minor families the dispersal unit is the seed itself, such as in Caryophyllaceae (**Fig.
+In some other minor families, the dispersal unit is the seed itself, such as in Caryophyllaceae (**Fig.
 1m**), Plantaginaceae (**Fig.
@@
 1o**).
-In Asteraceae, high germination without pretreatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; @RN120; @RN963; @RN1428], *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403].
-*Taraxacum officinale* germinates between 5 and 30 ºC [@RN120; RN1428; @RN1276], and has higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428].
+In Asteraceae, high germination without previous treatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; @RN120; @RN963; @RN1428], *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403].
+*Taraxacum officinale* germinates between 5 and 30 ºC [@RN120; RN1428; @RN1276] and has higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428].
 In *Stachys officinalis* (Lamiaceae), seeds need either cold stratification, light or alternating temperatures to germinate [@RN416; @RN128].
@@
 morphological dormancy) are widespread in Ranunculaceae and Apiaceae [@RN3214; @RN1112].
-*Ranunculus repens* (Ranunculaceae) germinates between 10 and 25 ºC, but the germination percentages were reported to be low [@RN931]; the same species has been reported to respond to alternating temperatures, which can promote its germination even in darkness [@RN3180].
-In Polygonaceae, *Rumex acetosa* can germinate immediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not, this difference is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814].
+*Ranunculus repens* (Ranunculaceae) germinates between 10 and 25 ºC, but the germination percentages have been reported to be low [@RN931]; the same species has been reported to respond to alternating temperatures, which can promote its germination even in darkness [@RN3180].
+In Polygonaceae, *Rumex acetosa* can germinate immediately after dispersal and between 7 and 27 ºC, while the congeneric *Rumex acetosella* does not: this difference is due to the former being able to germinate at constant temperatures in the darkness [@RN814; @RN3273], while the latter has an absolute requirement for light [@RN814].
 In *Heracleum sphondylium* (Apiaceae), growth of the embryo only occurs below 10 ºC, in moist conditions [@RN1112].
-*Sanguisorba minor* (Rosaceae) increases its germination after abrasion of the seeds with bleach [@RN194; @RN226], although germination without pre-treatment has also been reported [@RN298; @RN226].
-Seeds of the hemiparasitic species *Rhinanthus angustifolius* and *Rhinanthus minor* (Orobanchaceae) require relatively long periods of cold stratification to germinate [@RN4014], and are able to germinate in the dark [@RN2236].
+*Sanguisorba minor* (Rosaceae) increases its germination after abrasion of the seeds with bleach [@RN194; @RN226], although germination without previous treatment has also been reported [@RN298; @RN226].
+Seeds of the hemiparasitic species *Rhinanthus angustifolius* and *Rhinanthus minor* (Orobanchaceae) require relatively long periods of cold stratification to germinate [@RN4014] and can germinate in the dark [@RN2236].
 
-Although a wealth of studies has accumulated, a synthesis of seed ecology in European mesic meadows is missing.
-In this article, we review for the first time this topic, combining newly-generated data on seed morphology and germination with records from existing databases [@RN5036; @RN4968; @RN2987; @RN2319].
+Although a wealth of studies has accumulated, a synthesis of the seed ecology of European mesic meadows is still missing.
+In this article, we review for the first time this topic, combining newly generated data on seed morphology and germination with records from existing databases [@RN5036; @RN4968; @RN2987; @RN2319].
 The resulting dataset contains `r MSrecords` germination records of `r MSspeciesN` plant species from `r MScountries` European countries.
-To test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light we apply Bayesian meta-analysis [@RN5000].
-Further, using well-preserved meadows of the Iberian Peninsula as a study system, we analyze the covariation between seed traits and species environmental preferences, and compare the germination ecology of mesic meadows with that of other herbaceous plant communities from the same area.
+We use this dataset to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light, applying Bayesian meta-analysis [@RN5000].
+Further, using well-preserved meadows of the Iberian Peninsula as a study system, we analyse the covariation between seed traits and species environmental preferences, and compare the germination ecology of mesic meadows with that of other herbaceous plant communities from the same geographic area.
 
-# Materials and methods
+# MATERIALS AND METHODS
 
@@
 Furthermore, meadows from the Pyrenees are closest to the European optimum of mesic meadow vegetation while the Portuguese ones are in suboptimal areas at the limit of the European distribution of temperate meadows [@RN4893].
-Finally, the traditional management of meadows is relatively well preserved in these three regions compared to their European context [@RN4998; @RN4999].
+Finally, the traditional management of meadows (i.e.
+mowing for haymaking once or twice per year plus light grazing) is relatively well preserved in these three regions compared to their European context [@RN4998; @RN4999].
 
-Study meadows had been maintained by traditional agricultural practices: mowing for hay-making once or twice per year plus light grazing.
-Each vegetation sampling plot was placed in a square area (25-100 m^2^ area) situated in the central part of a meadow, avoiding the margins.
+The studied meadows were maintained by traditional agricultural practices: mowing for haymaking once or twice per year plus light grazing.
+Each vegetation sampling plot was placed in a square area (25-100 m^2^ area) situated in the central part of the meadow, avoiding the margins.
 Vegetation sampling took place in 2016-2017, at the peak of plant development, just before mowing.
@@
 Finally, we rescaled the values of all species to a 1-100 scale to obtain the cumulative cover values.
-We removed 208 species with cumulative cover values below 2%, considering them to be rare species that might have been recorded by chance and may not represent the core mesic meadows flora.
-We used the remaining 116 species as the core list to retrieve seed germination data for this article.
-The majority of these species were hemicryptophytes (78%), with some therophytes (16%) and a few chamaephytes and geophytes (3% of each).
+From the resulting list of species, we removed 208 species with cumulative cover values below 2%, considering them to be transient species [@RN4977] that might have been recorded by chance and may not represent the core mesic meadows flora.
+We used the remaining 116 species as the core list of meadows species to retrieve seed data for this article.
+Most of these species were hemicryptophytes (78%), with some therophytes (16%) and a few chamaephytes and geophytes (3% of each).
 The family with more species in the list was Poaceae (22%), with another 17% belonging to Fabaceae, 15% to Asteraceae, 8% to Apiaceae, and the rest of the families representing less than 5% each.
 
-We also used the relevés as a basis to characterize the preferences of the selected species for three ecological drivers of mesic meadow diversity: cold, summer drought and soil reaction [@RN2980].
+## Species ecological requirements
+
+We also used the relevés as a basis to characterize the preferences of the selected species for three environmental factors (cold, summer drought and soil reaction) which have been found to be major ecological drivers of mesic meadow plant diversity [@RN2980].
 For cold and drought, we used the coordinates of the plots to retrieve from CHELSA [@RN4766] the bioclimatic variables bio06 (minimum temperature of the coldest month) and bio14 (precipitation of the driest month).
-For soil reaction, we took from each plot five soil samples from between 0 and 20 cm depth with a Dutch auger and combined them to make a bulk soil horizon, which we subsequently air-dried, crumbled, finely crushed and sieved with a 2 mm screen, to finally measure the pH in H~2~O with a glass electrode in a suspension of soil and water (1:2.5).
-For each of these three environmental variables (bio06, bio14, pH) measured at the plot level we calculated the species niche centroids (SNCs).
-The SNC for any given species and variable is the mean of the environmental variable in all the plots were the species occurs, weighted by species cover in each plot [@RN4709].
+For soil reaction (pH), we took from each plot five soil samples from between 0 and 20 cm depth with a Dutch auger and combined them to make a bulk soil horizon, which we subsequently air-dried, crumbled, finely crushed and sieved with a 2 mm screen, to finally measure the pH in H~2~O with a glass electrode in a suspension of soil:water (1:2.5).
+With each of these three environmental variables (bio06, bio14, pH) measured at the plot level we calculated the species niche centroids (SNCs).
+The SNC for any given species and variable is the mean of the environmental variable in all the plots where the species occurs, weighted by species cover in each plot [@RN4709].
 The list of core species, with their cumulative covers and SNCs, is available at GitHub (see Data Availability Statement).
@@
 We distributed the seeds on the transparent glass of the scanner, in a 10 x 10 grid.
-For each sample, and without moving the seeds, we repeated the scans with black and white backgrounds.
+For each sample, and without moving the seeds, we repeated the scan with black and white backgrounds.
 In the case of the black background, we covered samples with a black box to avoid interference from environmental light.
@@
 We processed the scanned images using *ImageJ*, an open-source image processing program designed for scientific multidimensional images [@RN4708].
-The program calculates several biometric parameters for each seed on the sample, and among these we choose seed length and width.
+The program calculates several biometric parameters for each seed on the sample, and among those we chose seed length and width.
 Additionally, we retrieved species values of seed mass from the Seed Information Database [@RN2987] and of seed number at the individual/ramet level from the LEDA database [@RN2319].
@@
 
-We germinated the collected seeds using three germination treatments to determine the germination response to temperatures that are representative of the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as the summer temperature, and 30/20 ºC as sun-heated soil, e.g.
+We also germinated the collected seeds using three germination treatments to determine the germination response to temperatures that are representative of the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as summer temperature, and 30/20 ºC as sun-heated soil, e.g.
 soil exposed to sun after hay cutting.
-Additionally, we compared, at each of these temperature regimes, the germination of fresh seeds versus seeds subjected to a dormancy-breaking treatment.
+Additionally, we compared, for each of these temperature regimes, the germination of fresh seeds versus seeds subjected to a dormancy-breaking treatment.
 In the case of Fabaceae and other families that might present physical dormancy [@RN3214], the treatment consisted in scarification by chipping the seed coat with a scalpel.
@@
 After four weeks, we cut the seeds that failed to germinate and examined them under a magnifying glass.
-We classified them as normal when the embryo was visible and firm, empty when they lacked an embryo, and contaminated when they were moldy.
-We only considered normal seeds to calculate germination proportions and conduct subsequent analyses.
+We classified them as normal when the embryo was visible and firm, empty when they lacked an embryo, and contaminated when they were mouldy.
+We only considered normal seeds when calculating germination proportions and conducting subsequent analyses.
 
-In addition to this experimental data, we retrieved seed germination records from ENSCOBASE [@RN5036], the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for the nemoral biome [@RN4968].
-The combined dataset, including our own experimental data and the records from ENSCOBASE and *SylvanSeeds*, contained `r MSrecords` germination records (i.e.
+In addition to this experimental germination data, we retrieved seed germination records for the list of study species from ENSCOBASE [@RN5036], the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for the nemoral biome [@RN4968].
+These new records included additional records of species in our experimental dataset, plus records of new species absent from our experimental dataset, and in all cases corresponded to seed lots originally collected within Europe.
+The combined dataset, including our own experimental data and the records from ENSCOBASE and *SylvanSeeds*, contained `r MSspeciesN` species (i.e.
+90% of the core list of meadow species was covered) and `r MSrecords` germination records (i.e.
 germination proportions for a given seed lot of a species, recorded in a set of laboratory experimental conditions) from `r MScountries` European countries.
-There were `r MSspeciesN` species in the dataset (i.e.
-90% of the core list of meadow species was covered).
 Overall, `r MSseedsN` seeds had been used in the experiments.
@@
 
-We conducted all analyses in R [@RN2315], and the code for analysis and creation of the figures and manuscript is available at GitHub (see Data Availability Statement)
+We conducted all analyses in R [@RN2315], and the code for analysis and creation of the figures and manuscript is available at GitHub (see Data Availability Statement).
 
-To test the effect of germination cues on seed germination proportions, we performed a meta-analysis [@RN5000] of the germination dataset by fitting binomial generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) [@RN5036; @RN4965] using the R package *MCMCglmm* [@RN4755].
-We fitted models to the entire dataset, and also separately for each of the three botanical groups of mesic meadows: Poaceae, Fabaceae, and the other families.
-To account for the effect of a shared phylogeny in species traits, models included as a random effect a reconstructed phylogenetic tree for the study species.
-To create the phylogeny, we used the R package *V.PhyloMaker* [@RN4753] which contains an updated mega-tree of the seed plants based on Smith & Brown [-@RN4754].
+To test the effect of germination treatments on seed germination proportions, we performed a meta-analysis [@RN5000] of the germination dataset by fitting binomial generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) [@RN5036; @RN4965] using the R package *MCMCglmm* [@RN4755].
+We fitted four models: (1) to the entire dataset; and separately for each of the three botanical groups of mesic meadows: (2) Poaceae, (3) Fabaceae, and (4) a third group including the rest of the families.
+To account for the effect of a shared phylogenetic history in species traits, models included as a random effect a reconstructed phylogenetic tree of the study species.
+We created the phylogeny using the R package *V.PhyloMaker* [@RN4753] which contains an updated mega-tree of the seed plants based on Smith & Brown [-@RN4754].
 We placed taxa absent from the mega-tree at the genus-level basal node.
 The phylogenetic tree is available at GitHub (see Data Availability Statement).
-Germination proportion was the response variable.
-Models had, as fixed effects, the experimental conditions (scarification, stratification, GA~3~, temperature, alternating temperature and light).
+The response variable of the models was the proportion of germinated seeds.
+The fixed effects were the germination treatments (scarification, stratification, GA~3~, temperature, alternating temperature and light).
 Random effects included the phylogenetic tree, species identity, seed lot and source of the data.
-In all models, response variables were scaled so their contribution to the effect sizes could be compared.
+In all models, variables were scaled so their contribution to the effect sizes could be compared.
 We used weakly informative priors in all models, with parameter-expanded priors for the random effects.
@@
 We estimated the significance of model parameters by examining CIs, considering parameters with CIs overlapping with zero as non-significant.
-To estimate the phylogenetic signal of seed germination over all variables, we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by De Villemereuil et al.
+To estimate the phylogenetic signal in the models we used Pagels’s lambda (λ) [@RN4757], estimated simultaneously with the models by calculating the mean of the posterior distribution and the 95% CI of λ as indicated by De Villemereuil et al.
 [-@RN4756].
 When λ = 0, related taxa are no more similar than expected by chance, while when λ = 1, the trait is evolving following a constant variance random walk or Brownian motion model; intermediate values of λ indicate a phylogenetic correlation in trait evolution that does not fully follow a Brownian motion model [@RN4757].
-Detailed results of the MCMCglmms are available at GitHub (see Data Availability Statement).
+The detailed output of the models is available at GitHub (see Data Availability Statement).
 
@@
 stratification, scarification, temperature, alternating temperatures and light).
-To do so, for each cue and species, we calculated a weighted average of the cue levels (in the case of temperature, cue levels were the temperature treatments; for the other cues the levels were 0 = absence and 1 = presence), weighting by the germination proportion at each level.
-This approach underrepresents the importance of the levels that were not tested for a given species, but can serve as a proxy of the response to the germination cues when visualized across the whole dataset.
-We also included seed mass and seed number in the ordination.
+To do so, for each cue and species, we calculated a weighted average of the cue levels (in the case of temperature, cue levels were the temperature treatments; for the other cues, the levels were 0 = absence and 1 = presence), weighting by the germination proportion at each level.
+This approach underrepresents the importance of the levels that had not been tested for a given species, but can serve as a proxy of the response to the germination cues when visualized across the whole dataset.
+We also included seed mass and seed number in the ordination (species average values).
 We left GA~3~ out of the PCA because its ecological interpretation is subordinated to stratification (as both cues break physiological seed dormancy).
@@
 
-Finally, to compare the germination of mesic meadow species with other herbaceous communities, we retrieved data from previous works on the seed germination ecology of bogs and fens [@RN3480; @RN3212], alpine and subalpine grasslands [@RN2371], and coastal plant communities of rocky cliffs and sand dunes [@RN2865].
-These additional germination records corresponded to the same experimental methodology as the one employed for some of the germination experiments of this study: recently-collected seeds, untreated for physiological dormancy (but scarified in the cases of families known of having physical dormancy), had been subjected to three germination thermoperiods (14/4 ºC, 22/12 ºC, 30/20 ºC) in the same conditions as described previously.
-All seeds had been collected in the Cantabrian Mountains of Spain and the neighboring coast.
-We added these records and the records with matching experimental conditions from the meadows germination dataset, and performed a PCA of the resulting dataset.
+Finally, to compare the germination of mesic meadow species with that of other herbaceous communities, we retrieved data from previous works on the seed germination ecology of bogs and fens [@RN3480; @RN3212], alpine and subalpine grasslands [@RN2371], and coastal plant communities of rocky cliffs and sand dunes [@RN2865].
+These additional germination records had been obtained using the same experimental methodology as the one employed for some of the germination experiments of this study: recently collected seeds, untreated for physiological dormancy (but scarified in the cases of families known for having physical dormancy), had been subjected to three germination thermoperiods (14/4, 22/12, 30/20 ºC).
+All seeds had been collected in the Cantabrian Mountains of Spain or the neighbouring coast.
+We combined these records with the records with matching experimental conditions from the meadows germination dataset, and performed a PCA of the resulting dataset.
 The germination records for the other plant communities are available at GitHub (see Data Availability Statement).
 
-# Results
+# RESULTS
 
@@
 2**).
-Seed shape also showed a divergence between Poaceae and Fabaceae, with seeds being elongated in the former and rounder in the latter (**Fig.
+Seed shape also showed a divergence between Poaceae and Fabaceae, with seeds being elongated in the former and round in the latter (**Fig.
 2**).
@@
 
-When considering the full set of mesic meadow species, all of the six studied germination cues had a significant effect on final germination proportions (**Fig.
+When considering the full set of mesic meadow species, all six studied germination cues had a significant effect on final germination proportions (**Fig.
 3**).
@@
 In Poaceae (**Fig.
-3**), no effect was found for stratification, GA~3~ or temperature.
+3**), no effect was found for scarification, GA~3~ or temperature.
 Stratification had a negative effect on germination.
@@
 The strongest effect was that of the phylogeny (mean = `r MSphylo`, CI = `r MSphylol` - `r MSphyloh`), followed by the source of the data (mean = `r MSsource`, CI = `r MSsourcel` - `r MSsourceh`) and the seed lot (mean = `r MSpopulation`, CI = `r MSpopulationl` - `r MSpopulationh`).
-The phylogenetic signal in the germination responses was relatively high (lambda = `r MSlambda`, CI = `r MSlambdal` - `r MSlambdah`).
+The phylogenetic signal in the germination responses was relatively high (λ = `r MSlambda`, CI = `r MSlambdal` - `r MSlambdah`).
 
-## Seed traits and ecological preferences
+## Seed traits and species ecological preferences
 
@@
 The first PCA axis explained `r round(pca$eig[1, 2], 0)`% of the variation and was related to environmental preferences.
-The variables with the largest contribution to this first axis were soil reaction (pH), winter cold (bio06) and summer drought (bio14).
-This horizontal axis separated (i, left) species with preferences for sites with warm winter temperatures from (ii, right) species with preferences for sites with high summer rainfall and alkaline soils.
+The variables with the largest contribution to this first axis were soil reaction (pH), winter cold (bio06) and summer rainfall (bio14).
+This horizontal axis separated (i, left) species with preferences for sites with warm winter temperatures from (ii, right) species with preferences for sites with high summer rainfall and less acidic soils.
 Axis 2 explained `r round(pca$eig[2, 2], 0)`% of the variability, mostly related to seed traits.
 The main contributing variables were seed mass, seed number and the germination response to scarification.
-This axis separated (iii, bottom) Poaceae species that produced many seeds with a positive germination response to alternating temperatures from (iv, top) Fabaceae species that produced heavier seeds with a positive response to scarification.
+This axis separated (iii, bottom) Poaceae species that produce many seeds with a positive germination response to alternating temperatures from (iv, top) Fabaceae species that produce heavy seeds with a positive response to scarification.
 
@@
 
-Ordination resulted in a separation of mesic meadow species from species belonging to natural herbaceous communities of the same region (**Fig.
+Ordination resulted in a separation of mesic meadow species from plant species belonging to natural herbaceous communities of the same region (**Fig.
 5**).
-Axis 1 explained 78% of the variance and was positively associated to high germination at all three temperature treatments.
+Axis 1 explained 78% of the variance and was positively associated with high germination at all three temperature treatments.
 This horizontal axis separated (i, left) species with low germination across treatments from (ii, right) species with high germinability across treatments.
-The second axis explained 17% of the variation and separated (iii, bottom) species that responded more to the cool temperature germination treatment from (iv, top) species that responded more positively to the warm germination treatment.
+The second axis explained 17% of the variation and separated (iii, bottom) species that responded more to the cool germination temperature from (iv, top) species that responded more to the warm germination treatment.
 Meadow species tended to be situated in the right sector of the ordination, indicating high germinability; and at the center of the vertical axis, indicating a neutral response to temperature.
@@
 
-# Discussion
+# DISCUSSION
 
-Our meta-analysis of germination records showed that, overall, the regeneration of mesic meadows is characterized by (1) high germinability when compared to other herbaceous plant communities; (2) low correspondence between seed trait variability and the natural environmental drivers of mesic meadow diversity; and (3) a strong phylogenetic separation between the two major families, Poaceae and Fabaceae.
+Our meta-analysis of germination records showed that, overall, the seed ecology of mesic meadows is characterized by (1) high seed germinability when compared to other herbaceous plant communities; (2) low correspondence between seed trait variability and the natural environmental drivers of mesic meadow diversity; and (3) a deep phylogenetic separation between the two major families, Poaceae and Fabaceae.
 
-We found that alternating temperatures and light are the two most influential factors triggering germination, especially in species from the family Poaceae and the remaining ones, with the exception of Fabaceae.
+We found that alternating temperatures and light are the two most influential factors triggering germination, especially in species from the family Poaceae and the minor families, with the exception of Fabaceae.
 Alternating temperatures and light are usually considered to be micro-environmental cues that indicate, at a fine scale, the existence of safe sites for regeneration [@RN4719].
-The diurnal alternation of temperatures decreases with burial depth in the soil, and the alternation is also reduced by vegetation cover [@RN1380; @RN4910; @RN4952].
-Thus, a positive germination response to alternating temperatures could detect the depth of seed burial, and prevent the germination of seeds that are buried too deep for seedlings to survive before emerging from the ground [@RN4898].
-Perhaps more importantly in this habitat, alternating temperatures could also indicate that the vegetation cover has been diminished by either sowing or grazing, and therefore mark an appropriate time for seedling regeneration, when the competition by the established plants is lessened.
-Similar functions can be ascribed to the germination response to light, as physiologically-active light in significant amounts only travels the uppermost milliliters of soil [@RN4909] and the quality of light will be affected by vegetation cover [@RN604].
-Therefore, the germination response to alternating temperatures and light indicate conditions that are found immediately after mowing for haymaking, a predictable perturbation that occurs approximately at the same time every year, and which forces meadow plants to establish, grow and reproduce in the time frame determined by this perturbation [@RN4887; @RN4888].
-In addition, these traits would also promote the formation of a persistent soil seed bank [@RN4718; @RN4929; @RN1292].
-On the other hand, it is worth mentioning that Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate meadows [@RN5014; @RN5013].
+The diurnal alternation of temperature decreases with burial depth in the soil, and the alternation is also reduced by vegetation cover [@RN1380; @RN4910; @RN4952].
+Thus, a positive germination response to alternating temperatures could detect the depth of seed burial and prevent the germination of seeds that are buried too deep for seedlings to survive before emerging from the ground [@RN4898].
+Perhaps more importantly in mesic meadows, alternating temperatures could also indicate that the vegetation cover has been diminished by either mowing or grazing, and therefore mark an appropriate time for seedling regeneration, when the competition by the established plants is lessened.
+Similar functions can be ascribed to the germination response to light, as physiologically-active light in significant amounts only travels the uppermost millimetres of soil [@RN4909] and the quality of light will be affected by vegetation cover [@RN604].
+Therefore, the germination response to alternating temperatures and light indicates conditions that are found immediately after mowing for haymaking, a predictable perturbation that occurs approximately at the same time every year, and which forces meadow plants to establish, grow and reproduce in the time frame determined by this perturbation [@RN4887; @RN4888].
+Theoretically, these germination traits would also promote the formation of a transient soil seed bank [@RN4929; @RN1292], but it is worth mentioning that Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate meadows [@RN5014; @RN5013].
 
@@
 In this family, physical dormancy as a result of an impermeable seed coat works as a mechanism to detect seasonal cycles of temperature and humidity [@RN780].
-It has been also proposed that hardseedness can avoid seed predation, by preventing the emission of olfactory cues that are elicited by seed imbibition and that can be perceived by seed predators [@RN2568].
+It has been also proposed that physical dormancy can avoid seed predation, by preventing the emission of olfactory cues that are elicited by seed imbibition and that can be perceived by seed predators [@RN2568].
 In the case of this study, Fabaceae seeds are amongst the heaviest, and they clearly follow a separate regeneration strategy from that of the other families.
-While other groups rely on alternating temperatures and light as gap-detecting mechanisms, Fabaceae seeds do not respond to these factors, and instead they are mainly regulated by scarification, and they also respond to cooler temperatures than the rest of species.
-This would indicate that Fabaceae seeds tend to germinate when temperatures are cooler, and thus before or after the summer hay-making season.
-Their larger size, and the related larger reserves, could allow Fabaceae seedlings to emerge at times when the competition from the established vegetation would be higher.
+While other groups rely on alternating temperatures and light as gap-detecting mechanisms, Fabaceae seeds do not respond to these factors, and they are regulated instead by scarification, while they also respond to cooler temperatures than the rest of the families.
+This suggests that Fabaceae seeds would tend to germinate when temperatures are cooler, and thus before or after the summer haymaking season.
+Their larger size, and the related larger reserves, could allow Fabaceae seedlings to emerge from greater depths [@RN4898], e.g.
+when the meadow grass is still high, before mowing.
+Furthermore, larger seed size also increases seed survival in cattle dung [@RN5050] and thus improves the capacity of being dispersed by animal depositions [@RN5051].
 
 The germination response to average temperatures and stratification is understood to detect cues related to seasonal cycling and macroclimatic variation [@RN3063; @RN5036].
-Apart from the aforementioned response to cool temperatures in Fabaceae, the rest of the families did not show a response to average temperatures, further highlighting their reliance on micro-habitat cueing.
+Apart from the response to cool temperatures in Fabaceae, the rest of the families did not show a response to average temperatures, further highlighting their reliance on micro-habitat cueing.
 Moreover, the dominant family Poaceae showed a negative response to cold stratification.
@@
 
-The positive phylogenetic signal found in the models highlights the phylogenetic clustering of regeneration strategies that we have described so far: (1) the dominant family Poaceae, showing a lack of responses to seasonal cues (average temperature and cold stratification), and relying instead on large amounts of propagules and on detecting micro-niche cues (alternating temperatures, light) that can be associated to the yearly perturbation of mowing, (2) the second most-dominant family, Fabaceae, which do not respond to the micro-niche cues, and regulate their germination timing through scarification and cooler germination temperatures, possibly because their larger size and reserves allow their seeds to decouple their emergence timing from the mowing disturbance, and finally (3) the minor families that respond to micro-cues in a similar way than Poaceae, but which are differentiated from them by showing a positive response to cold stratification and G~3~, indicating that they rely on physiological seed dormancy.
-This third strategy would allow to fine-tune their germination with the local environmental conditions and the inter-annual climatic variation.
+The strong phylogenetic signal in germination responses highlights the phylogenetic clustering of regeneration strategies that we have described so far: (1) the dominant family Poaceae, showing a lack of response to seasonal cues (average temperature and cold stratification), and relying instead on large amounts of propagules and on detecting micro-niche cues (alternating temperatures, light) that can be associated to the yearly perturbation of mowing, (2) the second most-dominant family, Fabaceae, which do not respond to the micro-niche cues, and regulate their germination timing through scarification and cooler germination temperatures, possibly because their larger size and reserves allow their seeds to decouple their emergence timing from the mowing disturbance, and finally (3) the minor families that respond to micro-cues in a similar way than Poaceae, but which are differentiated from them by showing a positive response to cold stratification and G~3~, indicating that they rely on physiological seed dormancy to track seasonal cycles.
+This third strategy would allow to fine-tune their germination to the micro-environmental conditions plus the inter-annual climatic variation.
 Phylogenetic clustering is also apparent regarding seed shape, mass and number, with Poaceae producing many elongated and light seeds, and Fabaceae producing fewer, rounder and heavier seeds.
-The shape of the Poaceae seeds could make them particularly successful in being dispersed attached to hay, and Poaceae seeds tend to be overrepresented in seed mixes obtained via hay-making [@RN5046; @RN4894].
+The shape of Poaceae seeds could make them particularly successful in attaching to hay and dispersing with it, and Poaceae seeds tend to be overrepresented in seed mixes obtained via haymaking [@RN5046; @RN4894].
 Asteraceae and Apiaceae, some of the most abundant among the minor families, also tend to have shapes resembling those of Poaceae.
 
-As we have seen, this regeneration strategy of high germinability appears to be unrelated to environmental patterns in hay meadows, and seems more adapted to the predictable perturbation that is yearly mowing.
-In an abandoned meadow of the Prealps, resuming mowing promoted germination and emergence, although natural regeneration was limited due to a lack of seed dispersal to the site [@RN3073].
+As we have seen, the regeneration strategy of high germinability appears to be general in meadow species, unrelated to regional environmental gradients, and to be more responsive to the predictable perturbation that is yearly mowing.
+In an abandoned meadow of the Swiss Prealps, resuming mowing promoted germination and emergence, although natural regeneration was limited due to a lack of seed dispersal to the site [@RN3073].
 Mowing also promoted autumn germination in a dry grassland of northern Germany [@RN5025].
@@
 Ten Brink et al.
-[-@RN3570], when comparing congeneric herbaceous species from open and forest habitats found that species from open habitats needed less cold stratification.
+[-@RN3570], when comparing congeneric herbaceous species from open and forest habitats, found that species from open habitats needed less cold stratification.
 Similarly, seeds collected from a hay meadow germinated better when left untreated, rather than when being exposed to dormancy-breaking treatments [@RN4894].
-These high-germinability strategy has the consequence of greatly limiting the long-term seed bank of mesic meadows.
-Rather, the soil bank is transient in hay meadows and related grasslands [@RN5026; @RN5041; @RN5042].
-In Poland, the soil bank of a hay meadow was dominated by forbs of arable and weedy species, with low representation of Poaceae and Fabaceae [@RN5032].
+These high-germinability strategy has the practical consequence of greatly limiting the long-term seed bank persistence of mesic meadow species.
+Rather, the soil seed bank has been reported as being transient in hay meadows and related grasslands [@RN5026; @RN5041; @RN5042].
+In Poland, the soil bank of a hay meadow was dominated by arable and weedy forbs, with low representation of Poaceae and Fabaceae [@RN5032].
 
-Overall, this indicated that mesic meadows are quite sensitive to land use change, and their inability to form seed banks leads to the known dispersal limitation to natural regeneration [@RN3073], highlighting the importance of active actions of meadow restoration via seed supply.
-Human activities have shaped the regeneration of hay meadows, leading to a loss of seed dormancy and seasonal tracking, as has been found in many domesticated species [@RN4315].
+Overall, our results suggest that mesic meadows are quite sensitive to land use change.
+Their inability to form persistent seed banks leads to dispersal limitations to passive natural regeneration [@RN3073], highlighting the importance of active actions of meadow restoration via seed supply.
+For centuries, human activities have shaped the regeneration of hay meadows, leading to a loss of seed dormancy and decoupling seeds from seasonal cycles, as has been found in many domesticated species [@RN4315].
 The same anthropic processes that have shaped semi-natural mesic meadows have left them dependent on continued human intervention for their regeneration.
@@
 
-# Acknowledgements
+# FUNDING
 
-E.F.P.
-received financial support from the Government of Asturias and the FP7 – Marie Curie - COFUND programme of the European Commission (Grant ‘Clarín’ ACB17-19).
-This research was supported by project Interreg SUDOE “SOS PRADERAS” SOE1/P5/E037 (https://www.sospraderas.eu/en/).
+This work was supported by the European Regional Development Fund via Interreg Europe [SOS PRADERAS - SOE1/P5/E037]; the Government of Asturias and the FP7 – Marie Curie - COFUND programme of the European Commission [CLARÍN-ACB17-19 to E.F.P]; and the Jardín Botánico Atlántico [SV-20-GIJON-JBA to E.F.P.].
 
-# Author contributions
+# AUTHOR CONTRIBUTIONS
 
-All authors contributed data. E.F.P. conceived the study and performed the analyses. E.F.P. wrote the manuscript with help from A.C. All authors revised the manuscript, and approved the final version.
+All authors contributed data. E.F.P. conceived the study and performed the analyses. E.F.P. wrote the manuscript with help from A.C. All authors revised the manuscript and approved the final version.
 
-# Data availability
+# DATA AVAILABILITY
 
@@
 
-# References
+# LITERATURE CITED
 
@@
 
-# Figures
+# FIGURES
 
@@
 Principal Component Analysis ordination of mesic meadow species considering their environmental preferences and their seed traits.
-Each point is a species, colored by the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families.
+Each point is a species, coloured by the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families.
 Labels indicate the contribution of the variables to the axes: grey-background labels for environmental preferences, and white-background labels for seed traits.
@@
 The seed traits are seed mass and the germination relative indices for the response to scarification, stratification, average germination temperature (temperature), alternating temperature (alternating) and light.
-All environmental preferences were aligned to the first axis, while the seed traits were aligned to the second axis, showing that environment and seed traits are separate axis of variation in mesic meadows.
-Stratification and average germination temperatire showed very low variation, in accordance with their small effect on Figure 2."}
+All environmental preferences aligned to the first axis, while seed traits aligned to the second axis, showing that environment and seed traits are separate axis of variation in mesic meadows.
+Stratification and average germination temperature showed very low variation, in accordance with their small effect on Figure 2."}
 knitr::include_graphics("../results/figures/Fig4.png")
@@
 Labels indicate the contribution of the germination treatments to the axes.
-All species had been collected in herbaceous vegetation types of the Cantabrian Mountains of Spain and the neighboring coast.
-All seeds were recently collected and untreated, except for scarification, which was applied routinely to all botanical families presenting it.
+All species had been collected in herbaceous vegetation types of the Cantabrian Mountains of Spain and the neighbouring coast.
+All seeds were untreated, except for scarification, which was applied routinely to all botanical families presenting physical dormancy.
 Germination treatments consisted in 12/12 h periods with a warmer phase in light and a cooler phase in darkness.

### 2021-07-16 src/figures.R

@@ -381,6 +381,19 @@ ggplot(pcaInds, aes(x = Dim.1, y = Dim.2)) +
   annotate("text", x = -.55, y = -0.6, label = "Less\ngermination", color = "grey50") +
   scale_color_manual(values = c("#fdda62",  "saddlebrown", "#3dc6fa", "#5fc765")) -> f5;f5
 
-
 ggsave(f5, file = "results/figures/Fig5.png", 
        path = NULL, scale = 1, width = 173, height = 140, units = "mm", dpi = 600)
+
+# For publication
+
+ggsave(f2, file = "results/figures/Fig2.jpg", 
+       path = NULL, scale = 1, width = 173, height = 110, units = "mm", dpi = 300)
+
+ggsave(f3, file = "results/figures/Fig3.jpg", 
+       path = NULL, scale = 1, width = 173, height = 65, units = "mm", dpi = 300)
+
+ggsave(f4, file = "results/figures/Fig4.jpg", 
+       path = NULL, scale = 1, width = 173, height = 140, units = "mm", dpi = 300)
+
+ggsave(f5, file = "results/figures/Fig5.jpg", 
+       path = NULL, scale = 1, width = 173, height = 140, units = "mm", dpi = 300)

### 2021-07-16 src/numbers.R

@@ -57,12 +57,18 @@ read.csv("../data/germination.csv") %>%
             Stratification = weighted.mean(Stratification, w = Germination),
             GA3 = weighted.mean(GA3, w = Germination)) %>%
   merge(read.csv("../data/traits.csv"), by = "Taxon") %>%
-  select(Taxon, Family, Temperature:Stratification, Seed.mass) %>%
+  mutate(Family = ifelse(Family %in% c("Leguminosae", "Poaceae"), Family, "Others"),
+         Family = ifelse(Family %in% "Leguminosae", "Fabaceae", Family)) %>%
+  mutate(Family = fct_relevel(Family, c("Poaceae", "Fabaceae", "Others"))) %>%
+  select(Taxon, Family, Temperature:Stratification, Seed.mass, Seed.number, bio06:pH) %>%
   na.omit() -> traits
 
 ## Do PCA
 
 traits %>%
+  gather(Trait, Value, Temperature:Seed.mass) %>%
+  mutate(Value = log(0.01+ Value)) %>%
+  spread(Trait, Value) %>%
   select(-c(Taxon, Family)) %>%
   FactoMineR::PCA(graph = FALSE) -> pca
 

### 2021-10-14 README.Rmd
AoB 1st revision 🔎

--- 2021-04-30 README.Rmd
@@
 
-# Seed germination ecology of European mesic grasslands
+# Seed ecology of European mesic meadows
 
-This repository stores all the information related to the manuscript *Seed germination ecology of European mesic grasslands*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
+This repository stores all the information related to the manuscript *Seed ecology of European mesic meadows*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
 
@@
 
-* `data` Data files including the seed germination dataset, the species traits, and the phylogenetic tree.
+* `data` Data files including the seed trait datasets (European, Iberian and Cantabrian) and the phylogenetic tree.
 * `doc` Files to create the manuscript using `Rmarkdown`.

### 2021-10-14 doc/cover-letter.Rmd
AoB 1st revision 🔎

---
title: "Manuscript 21508 - Response to review"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output: word_document
bibliography: refs.bib
csl: annals-of-botany.csl
link-citations: yes
always_allow_html: true
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear Editor,

Please find attached our revision of manuscript 21508. We would like to thank the handling editor Mick Hanley and the three reviewers for their constructive comments. Here is a summary of the main changes we have introduced into the text:

* We have expanded the ecological context of the study, addressing why seed traits and species ecological requirements may not correlate in these semi-natural habitats.
* We discuss how the stressors in the other habitats could explain the differential germination syndromes.
* We have revised the discussion to make it more explicit that the "domestication" is an interpretation we make after synthesizing the results (and previous research).
* We have expanded the conclusions as related to restoration, introducing the dilemma about sourcing seed from "wild" vs. "domesticated" populations.
* We have used a European-level list of mesic meadow species to get data (and conduct analysis), addressing the concern of the article conclusions being limited only to Iberia.
* We have revised the methods, indicating exactly which (and how many) data is used for each of the tests.

Below is a point-by-point response to the issues found by the reviewers. We have also included a track-changes version of the manuscript as supplementary material.

## Referee #1

> **Despite a lot of great information provided, I believe the manuscript could be improved in several key ways before publication. First, I think the authors ought to make a stronger argument for why this study was conducted in the introduction. As written, a clear justification is lacking thus results seem to read like a case study rather than an investigation. Second, I ask the authors to more clearly state what was done and why in the methods. As written, it is tough to follow which data are being used in which analyses. Finally, I believe the authors could more clearly state why their results are novel and important throughout the discussion (i.e., why do they think traits do not map onto climate variables, what are the implications for restoration/conservation). With these improvements, I trust this will be a valuable contribution to our understanding of how seed traits can inform plant community ecology and its applications. **

Please see below comments on these various points.

> **Abstract: Overall the aim remains unclear. The methods provide more details, but why was this study done is still missing from the "background and aims" section. Consider adding a "however..." or "but..." (... data are unavailable, knowledge is lacking) statement after "...urban greening and rewilding" or a "to" statement after medic meadows". Also, the conclusion seems tangential to the results - clearer links between these two sections in the abstract would strengthen the key take-home messages of the study.**

We have added the "however" to the aims ("However, limited knowledge of species germination traits is a bottleneck to the development of a competitive native seed industry").

> **Page 2, lines 19 - 20: Consider changing "their species-rich seed mixes" to "these species-rich communities" as "seed-mix" has a specific meaning in restoration ecology.**

Changed to "These species-rich communities can be a source of seed mixes for ecological restoration".

> **Page 3, line 2: change to active voice: "Here, we synthesize..."**

Done.

> **Page 2; line 18: High germination of all seeds or poacea?**

High average germination across the species in the community, as represented by Figure 5.

> **Introduction: I think a lot of the specific details about Poaceae, Fabacea, and other family germination can be made more succinct. What are the key messages of these paragraphs - that there is high variability in germination strategies? Consider including topic sentences and conclusion sentences with selected details to make the paragraphs on germination of these groups more streamlined. **

Please see a response to a similar comment below.

> **Also, despite all of the great information provide, a compelling reason for why this study was conducted is missing. More focus on the why of the study would strengthen the introduction. In addition, hypotheses about what relationships the authors are expecting to observe would clarify for the reader why specific data were collected and methods used. Finally, hypotheses would allow the authors to more directly relate results to expectations and other research in the discussion. **

We do believe the "why" is developed in the paragraphs about the importance of meadows for European nature, and the importance of a knowledge basis for seed-based restoration. We have revised the end of the introduction to stress this. We could further rephrase the last part of the introduction into directional hypotheses, but we feel this would be untrue to the "research in context" nature of the manuscript; our goal was to provide an overview about the seed ecology of mesic meadow plants, synthesising data and published works into a coherent narrative, rather than to tests some specific hypotheses as would be the objective in a regular research paper.

> **Page 3, line 12: Consider a new sentence beginning with "during".**

Done.

> **Page 3, line 17: real and imaged? This is confusing... **

Changed "imagined" to "fictional".

> **Page 4, line 10: Great "however" statement. Can this be made more succinct and added to the abstract? **

We have done as suggested.

> **Page 4, line 16: "framework" not needed**

Removed.

> **Page 4, line 19: it is unclear whether the biomass is of individual poacea plants or of the whole community **

Changed to "the biomass of the community".

> **Page 5, lines 6 - 9: Are all of these references necessary? **

We agree that the references are very numerous and perhaps reiterative, however, we decided to be comprehensive because of the nature of the article as a *research in context* review hybrid. We would like potential readers to be able to access a list of references which is as complete as possible.

> **Page 4, line 10: comparatively to...?**

To stored seeds. We added "than stored seeds".

> **Page 7, line 10: nutritional scope for which animals - insects, granivores, ungulates, all?**

We have added "for livestock".

> **Materials and Methods: I am a bit confused about why specific treatment were done to specific groups. Justification of these methods is warranted. Also, additional clarification of specific models that were run for the different groups is needed (i.e., what are the specific fixed vs. random effects in each). Also, it remains unclear about what communities/meadows these analyses relate to - are they all for the meadows that were sampled by the authors, or are additional meadows included? **

Please see responses below. We have reformed the methods to make clearer which data is used for each test.

> **Page 10, line 19 - 22; Page 11 lines 10 -14: These sounds more like results than materials/methods13:**

These lines contain a description of part of the datasets used in the analyses. Since this description is not related to seed ecology (but to species composition), we feel it is more appropriately placed in the materials and methods.

> **Page 13, lines 1 - 5: How reliable are values obtained from these databases? Were all values for a species used, or were they filtered/limited in some way?**

The Seed Information Database provides an average value per species. For the LEDA database, we averaged all records per species. These values are reliable for broad-scale species comparisons, and are regularly used for such research.

> **Page 13; lines 12 - 16: Could these treatments affect results, especially since the same treatment was not done to every species? A comment to justify why these methods are appropriate is warranted.**

The purpose of the treatment was to measure the germination of non-dormant seeds. Since different taxonomic groups have different dormancy types, we had to adjust the treatment to the taxonomic group. We have added the sentence "This treatment was adjusted to the most probable dormancy type per taxonomic group".

> **Page 16; line 3: Are these fixed effects for every model (poacea, fabacea, other)? From the methods, it does not seem like all treatments were applied to all groups?**

The effects were the same for every model. Each of the three taxonomic groups had records of responses to each of the treatments. The treatments mentioned by the reviewer refer only to the Iberian germination dataset that we generated ourselves, but the models were applied to an extended dataset (our dataset + ENSCOBASE + SylvanSeeds). We have reformed the materials and methods to make clearer which data was used for each test.

> **Page 16; line 5: which variables were scaled? **

The fixed effect variables (the germination treatments). We now indicate this in the text.

> **Page 17: A clear list of which traits (and the data source) are and are not included in the PCA would be helpful.**

We have reformed the materials and methods to make clearer which data was used for each test.

> **Page 17, lines 16 - 22: How close are these comparison communities to the ones the authors sampled? Are the authors controlling for spatial auto-correlation in any way? I am suspicious about how informative these comparisons are if other factors drive germination patterns (i.e., geography).**

All communities are from the same 75 km radius region, and the original sampled locations are interspersed, i.e. not clustered together by community (except of course the coastal communities which occur along the coast). No control was made for spatial auto-correlation, we would like to stress that this part of the analyses is only a visual analysis to further support the high germinability of meadow seeds which is found in the rest of the results (as well as the literature). See also responses regarding comments about the stressors of the different vegetation types influencing the germination patterns.

> **Discussion: I find the discussion could be improve with more narrative about why the authors findings are novel and important (see comments in introduction section). In addition, I find the results of seed traits not being related to climate quite interesting (Fig 4), but this is hardly discussed. **

We have expanded the discussion to comment on the lack of a relationship between trait variability and environment, and also on the restoration dilemma on whether to source "domesticated" or "wild" populations of meadow plants.

> **Page 22; Lines 11- 16: This ought to be broken up into two sentences to improve clarity. **

Done.

> **Page 23; lines 4 - 6 & 12 - 17 & 16 - 20: This discussion of species relationship to cold temperatures via. stratification seems contradictory to the results which found high separation of seed traits and groups from climate variables. The fact that stratification is important but seed traits were not related to climate traits makes me wonder if a mismatch exists in the traits and climate variables the authors chose to investigate? Could it be that the wrong traits were looked at or that the traits investigated align with other (unmeasured) climate variables? Additional clarification is needed here in the discussion (and in the introduction) for *why* the authors suspect these specific traits and climate variables to be related.**

One thing is that the seeds show responses to environmental cues that relate to the macroclimate of the biome in which they occur, and a different thing is that this response shows variation along climatic gradients within the biome. For example, the species of an alpine flora might all respond to cold stratification, but those species from higher elevations might require longer periods of cold stratification. We do not see the contradiction. The relationships between germination cues and climate are based on numerous references, several of which we mention in the discussion. We have expanded the discussion to comment this lack of relationship.

> **Page 23; lines 12 - 14: This sentence is confusing.**

We have changed the sentence to "Seed germination responses to average temperatures and stratification are understood to be a mechanism to track seasonal climatic cycling ".

> **Page 24; lines 3 - 14: I recommend breaking this up into separate sentences to improve clarity.**

We have made it so.

> **Page 25; line 3: Because responses to yearly mowing were not actually tested, this remains speculative. I recommend the authors make this distinction and mention that this is a possible "selective force" in these systems, but that additional work is needed to test this hypothesis. **

We have made the suggested change.

> **Page 25; lines 9 - 12: These sentences are confusing.**

We have rephrased the sentences to attempt to make them clearer.

> **Page 25; line 21: I disagree with this statement, as this was not tested in the present study.**

This statement is based on the interpretation of our results plus the conclusions of several of the articles cited in the discussion. We have revised the paragraph to make this clearer.

> **Figure 5: Listing the three germination treatments here in the caption would be helpful.**

We have done so.

## Referee #2

> **Several genera of Poaceae that grow in European meadows show seed heteromorphism (see e.g. Aegilops, Avena, Dasypirum, Poa). They produce different seed morphs characterized by different seed traits in terms of dormancy, longevity, morphology, phenology, tolerance to abiotic stresses and susceptibility to predation. This is connected to bet hedging ecological strategies, typical of ruderal/disturbed habitats. I wonder if this can be relevant also in characterizing the seed ecology of the grass species of European mesic meadows analysed by the authors. This is true also for other plant families mentioned in the paper, mainly Asteraceae.**

The reviewer raises an interesting point, we have added this to the description of seed morphology in grasses.

> **I would suggest to the authors to add brief definitions of some specialistic terms (e.g. hemiparasitic, mesic, mesophilous grasses) for the benefit of the non specialistic reader.**

We have tried to do as suggested, but the only way we have found would be to insert sentences in parentheses in the middle of other sentences, in parts of the text that are already quite dense (and one of these parts is the critical first sentence). But we think that any interested reader can find the definition of such terms easily on the internet.

> **The connection of human activities with seed ecology in mesic meadows is really interesting but I do not feel that is fully demonstrated by the results presented by the authors, appearing to me more as a speculation. The paper nicely characterizes the seed ecology of these plant communities, with comparisons with natural herbaceous communities of the same geographical area. This is to me the main focus of this study, while the supposed "domestication syndrome" of mesic meadow species appears to me as an interesting speculation. I would suggest to the authors to clarify this point and perhaps consider changing the title of the paper to better fine-tune it with the results obtained by their analysis. **

We have changed the discussion (and the title) to make clearer that the domestication is an interpretation that we derive from our results and previous articles, and that it will be necessary to do future research with the specific goal of testing it.

> **The authors rightly state that a deeper knowledge of germination ecology in mesic meadows is fundamental for a better management and conservation of this habitat and also that the dependence of these meadows on human activities highlights the importance of active restoration practices. I was wondering if the authors, based on their results, could provide more practical implications/advices for the management and seed-based restoration of this habitat that is drawing lots of attention in rewilding projects in the last few years.**

We have expanded the discussion to address how germinability affects the ease to use meadow seeds for restoration, and also the dilemma on whether to source seeds from "domesticated" meadows or from "wild" populations of meadow species.

## Referee #3

> **Most of the work is based on data from Iberian Peninsula and consequently, I have doubts about the possibility to extrapolate results to the rest of European meadows. For instance, one of the main conclusions is the high seed germinability of mesic meadows when compared to other herbaceous plant communities. However, this comparison is made exclusively from data of Cantabrian Mountains. Other main result is the low correspondence between seed traits and species ecological preferences, but if I have not misunderstood, both species and ecological preferences were studied only in the Iberian Peninsula. Finally, the last main result, a clear phylogenetic difference between Poaceae and Fabaceae, is not new. Authors should explain much better why their results can be extrapolated outside the Iberian Peninsula. [] One of my concerns refers to selection of species for the analyses since it was based only on relevés from Iberian Peninsula, two countries from the 32 countries included in this study. I wonder if these species are really representative of all European meadows. Maybe, authors can add some species inventories for other countries to justify the species selection.**

In the previous version, we used an Iberian species list to collect seed germination data from across Europe. In this new version, we have used a European species list to collect the data, and therefore we feel it is more extrapolable outside the Iberian Peninsula. We have repeated the germination models with the new dataset based on the European species list, and updated the results consequently (the only change is that stratification is now a not significant effect, in all models except for a negative effect in Poaceae).

> **Authors should explain how many species were analysed by the own authors and how many species were included from published databases. Furthermore, authors should give the complete reference for one of the used databases (Carta et al. 2021) (see p. 28, ln 15). A summary table with all analysed species as well as their responses to treatments, the provenance of the seeds and the corresponding reference should be included as supplementary material.**

We have included the numbers of species as requested. We have also removed the Carta reference since it is yet unpublished, and substituted it by a link to ENSCOBASE. We think the supplementary table is not necessary since the actual datasets are provided at GitHub (see the data availability section in the manuscript).

> **In addition, Poaceae and Fabaceae were considered as the two main families and the analyses and discussion were focussed on them. However, while Fabaceae included 17 % of the species, Asteraceae included 15 % (p.11 ln. 12-13) and consequently, Asteraceae should be incorporated in the study as an additional relevant plant family and not in the third minor group with the rest of the species. That is, authors should differentiate 4 botanical groups (p. 15 ln. 17). **

We see the reviewer's point, and as a matter of fact we had a separate Asteraceae group in the earliest drafts. However we decided to merge them with the rest for three reasons. (1) Even if the Asteraceae make up 15% of the species in the original list, their contribution to the meadow biomass (as seen by the relevé covers) is much smaller. (2) Furthermore, the classification in three groups follows what is traditional in grassland studies as the three groups contribute differentially to livestock nutrition (Collins, M.; Nelson, C.J.; Moore K.J. Barnes, R.B. 2020. Forages, Volume 1: An Introduction to Grassland Agriculture, 7th Edition. Wiley-Blackwell Publishing. 968 pp.). (3) And finally, separating Asteraceae would mean that the Asteraceae group would have a comparatively low record number, and make the model outputs less reliable.

### 2021-10-14 doc/manuscript.Rmd
AoB 1st revision 🔎

@@
 ---
-title: "RESEARCH IN CONTEXT: Seed ecology of European mesic meadows: the domestication of plant regeneration strategies"
+title: "RESEARCH IN CONTEXT: Seed ecology of European mesic meadows"
 output: 
@@
 * **Background and Aims** European mesic meadows are semi-natural open habitats of high biodiversity and an essential part of European landscapes.
-Their species-rich seed mixes can be a source of plant material for ecological restoration, urban greening and rewilding.
-Here, we make a synthesis of the seed ecology of mesic meadows.
+These species-rich communities can be a source of seed mixes for ecological restoration, urban greening and rewilding.
+However, limited knowledge of species germination traits is a bottleneck to the development of a competitive native seed industry.
+Here, we synthesize the seed ecology of mesic meadows.
 * **Methods** We combined our own experimental data with data obtained from databases to create a combined dataset containing `r MSrecords` germination records of `r MSspeciesN` plant species from `r MScountries` European countries.
@@
 The mesic meadows of Europe are semi-natural open habitats that occupy clearings created by human intervention over a temperate wooded landscape [@RN5033; @RN5045], in sites with moderately fertile and well-drained soils [@RN4698].
-Once, these were novel plant communities, with a composite flora made up of species from different ecological backgrounds; during historical times, their species composition further evolved following changes in human events and management practices [@RN5035; @RN5045].
+Once, these were novel plant communities, with a composite flora made up of species from different ecological backgrounds.
+During historical times, their species composition further evolved following changes in human society and management practices [@RN5035; @RN5045].
 Mesic meadows are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034].
-Nowadays, mesic meadows maintained by agricultural practices are an essential aspect of European landscapes, both real [@RN4996] and imagined [@RN5029].
+Nowadays, mesic meadows maintained by agricultural practices are an essential aspect of European landscapes, both real [@RN4996] and fictional [@RN5029].
 
@@
 Understanding seed germination is part of the practical scientific framework needed to tackle large-scale ecological restoration challenges [@RN3507] and to maintain *ex situ* collections of plant genetic resources [@RN3763].
-When such a knowledge framework is available, ambitious regional schemes of seed-based landscape restoration can be designed [@RN4666].
+When such knowledge is available, ambitious regional schemes of seed-based landscape restoration can be designed [@RN4666].
 
-Mesic meadows are dominated by mesophilous grasses of the family Poaceae, which make up most of their biomass and define the structure of the vegetation.
+Mesic meadows are dominated by mesophilous grasses of the family Poaceae, which make up most of the biomass in the community and define the structure of the vegetation.
 The dispersal unit in Poaceae is generally the floret (**Fig.
@@
 1b**) that are assumed to aid dispersal, although it is not always the case that they do [@RN1290].
-Meadow Poaceae seeds have been reported as having physiological seed dormancy in various degrees [@RN5018; @RN3214; @RN5024], but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273].
-Freshly harvested seeds are comparatively more dormant, but dormancy tends to disappear quickly in dry storage [@RN5024; @RN5018].
+Several genera of Poaceae that grow in European meadows show seed heteromorphism (see e.g.
+*Aegilops*, *Avena*, *Dasypirum*, *Poa*) [@RN5063].
+They produce different seed morphs characterized by different seed traits in terms of dormancy, longevity, morphology, phenology, tolerance to abiotic stresses and susceptibility to predation.
+This is connected to bet hedging ecological strategies, typical of ruderal/disturbed habitats.
+Regarding germination, meadow Poaceae seeds have been reported as having physiological seed dormancy in various degrees [@RN5018; @RN3214; @RN5024], but germinability is usually high even without treating the seeds with cold stratification [@RN1205; @RN1260; @RN1271; @RN1229; @RN372; @RN789; @RN680; @RN1290; @RN398; @RN223; @RN334; @RN1281; @RN5024; @RN3273].
+Freshly harvested seeds are comparatively more dormant than stored seeds, but dormancy tends to disappear quickly in dry storage [@RN5024; @RN5018].
 Germination has been reported to occur at temperatures ranging from 5 to 30 ºC [@RN3273; @RN1229; @RN1261; @RN1230; @RN1292; @RN5024].
@@
 As in Poaceae, seeds of Fabaceae have been reported to germinate in high numbers across a range of temperatures from 5 to 25 ºC [@RN652; @RN3273].
-Fabaceae seeds have been described as not responsive to light and capable of germinating in darkness [@RN1339; @RN3273].
+Fabaceae seeds have been described as not being responsive to light and as capable of germinating in darkness [@RN1339; @RN3273].
 
 Poaceae and Fabaceae are accompanied by a diversity of other families which, even if present in lower abundances, contribute to the high biodiversity of the system and increase the aesthetic value of the meadows as perceived by people [@RN5038; @RN5039].
-They also add nutritional scope, being richer than Poaceae and Fabaceae in specific elements [@RN5028; @RN5040].
+They also add nutritional scope for livestock, being richer than Poaceae and Fabaceae in specific elements [@RN5028; @RN5040].
 In many of these families (e.g.
@@
 In Asteraceae, high germination without previous treatments has been reported in *Taraxacum officinale* [@RN1276; @RN194; @RN120; @RN963; @RN1428], *Hypochaeris radicata* [@RN194; @RN1403] and *Achillea millefolium* [@RN1403].
-*Taraxacum officinale* germinates between 5 and 30 ºC [@RN120; RN1428; @RN1276] and has higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428].
+*Taraxacum officinale* germinates between 5 and 30 ºC [@RN120; @RN1428; @RN1276] and has higher germination in light [@RN1015; @RN963; @RN1172] and in alternating temperatures [@RN1428].
 In *Stachys officinalis* (Lamiaceae), seeds need either cold stratification, light or alternating temperatures to germinate [@RN416; @RN128].
@@
 Although a wealth of studies has accumulated, a synthesis of the seed ecology of European mesic meadows is still missing.
-In this article, we review for the first time this topic, combining newly generated data on seed morphology and germination with records from existing databases [@RN5036; @RN4968; @RN2987; @RN2319].
+In this article, we review for the first time this topic to provide an overview of traits related to plant regeneration by seed, as a knowledge basis to expand our understanding of meadow communities and assist in their management by conservation and restoration practitioners.
+We combine newly generated data on seed morphology and germination with records from existing databases [@RN4968; @RN2987; @RN2319].
 The resulting dataset contains `r MSrecords` germination records of `r MSspeciesN` plant species from `r MScountries` European countries.
@@
 
-## Selection of mesic meadow species
+## Data sources
 
-To create a list of representative mesic meadow species for inclusion in our analysis, we used a dataset comprising 118 vegetation relevés (i.e.
-records of plants species co-occurring in sampling plots) from three Western European regions with well-maintained mesic meadows: 43 relevés from Northern Portugal, 25 from the Cantabrian Mountains of Spain and 50 from the Pyrenees.
-These relevés contain a sample of mesic meadow diversity, as they were recorded along a major stress gradient related to summer drought [@RN2980]: the Pyrenees and Cantabrian Mountains have a temperate macroclimate, whereas Northern Portugal is transitional between the temperate and Mediterranean macroclimates.
+### Selection of mesic meadow species
+
+To create a list of representative European mesic meadow species for inclusion in our analysis, we used the list of species provided by Chytrý et al.
+[-@RN5061] as dominant, constant and diagnostic species of the EUNIS habitats "low and medium altitude hay meadows" and "mountain hay meadows".
+This list included 120 species and was used to test the response to germination cues at the European level (see below).
+
+To conduct further analyses at the Iberian level, we constructed a second species list based on a dataset comprising 118 vegetation relevés (i.e.
+records of plants species co-occurring in sampling plots) from three Iberian regions with well-maintained mesic meadows: 43 relevés from Northern Portugal, 25 from the Cantabrian Mountains of Spain and 50 from the Pyrenees.
+These relevés were used to assess variation in seed responses along environmental gradients (i.e.
+the relationship between species seed traits and species ecological requirements, see below), as they were recorded along a major stress gradient related to summer drought [@RN2980]: the Pyrenees and Cantabrian Mountains have a temperate macroclimate, whereas Northern Portugal is transitional between the temperate and Mediterranean macroclimates.
 Furthermore, meadows from the Pyrenees are closest to the European optimum of mesic meadow vegetation while the Portuguese ones are in suboptimal areas at the limit of the European distribution of temperate meadows [@RN4893].
@@
 
-The studied meadows were maintained by traditional agricultural practices: mowing for haymaking once or twice per year plus light grazing.
+The Iberian meadows were maintained by traditional agricultural practices: mowing for haymaking once or twice per year plus light grazing.
 Each vegetation sampling plot was placed in a square area (25-100 m^2^ area) situated in the central part of the meadow, avoiding the margins.
@@
 From the resulting list of species, we removed 208 species with cumulative cover values below 2%, considering them to be transient species [@RN4977] that might have been recorded by chance and may not represent the core mesic meadows flora.
-We used the remaining 116 species as the core list of meadows species to retrieve seed data for this article.
+We used the remaining 116 species as the core list of Iberian meadow species.
 Most of these species were hemicryptophytes (78%), with some therophytes (16%) and a few chamaephytes and geophytes (3% of each).
@@
 
-## Species ecological requirements
+### Species ecological requirements
 
-We also used the relevés as a basis to characterize the preferences of the selected species for three environmental factors (cold, summer drought and soil reaction) which have been found to be major ecological drivers of mesic meadow plant diversity [@RN2980].
+We also used the Iberian relevés as a basis to characterize the preferences of the selected species for three environmental factors (cold, summer drought and soil reaction) which have been found to be major ecological drivers of mesic meadow plant diversity [@RN2980].
 For cold and drought, we used the coordinates of the plots to retrieve from CHELSA [@RN4766] the bioclimatic variables bio06 (minimum temperature of the coldest month) and bio14 (precipitation of the driest month).
@@
 The SNC for any given species and variable is the mean of the environmental variable in all the plots where the species occurs, weighted by species cover in each plot [@RN4709].
-The list of core species, with their cumulative covers and SNCs, is available at GitHub (see Data Availability Statement).
 
-## Seed morphology and germination dataset
+### Seed morphology
 
@@
 
-We also germinated the collected seeds using three germination treatments to determine the germination response to temperatures that are representative of the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as summer temperature, and 30/20 ºC as sun-heated soil, e.g.
+### Seed germination
+
+We constructed a germination dataset by combining the results of our own germination experiments with Iberian seeds, and records of seed germination of meadow seeds from across Europe.
+
+We germinated the collected Iberian seeds using three germination treatments to determine the germination response to temperatures that are representative of the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as summer temperature, and 30/20 ºC as sun-heated soil, e.g.
 soil exposed to sun after hay cutting.
 Additionally, we compared, for each of these temperature regimes, the germination of fresh seeds versus seeds subjected to a dormancy-breaking treatment.
+This treatment was adjusted to the most probable dormancy type per taxonomic group [@RN3214].
 In the case of Fabaceae and other families that might present physical dormancy [@RN3214], the treatment consisted in scarification by chipping the seed coat with a scalpel.
@@
 
-In addition to this experimental germination data, we retrieved seed germination records for the list of study species from ENSCOBASE [@RN5036], the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for the nemoral biome [@RN4968].
+In addition to this experimental germination data, we retrieved European seed germination records for the European list of study species from ENSCOBASE, the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for the nemoral biome [@RN4968].
 These new records included additional records of species in our experimental dataset, plus records of new species absent from our experimental dataset, and in all cases corresponded to seed lots originally collected within Europe.
@@
 subjected to previous incubation in dormancy-breaking conditions, including treatments of wet incubation under cold, warm and combinations of cold and warm conditions) in `r MSstratY` records.
-Finally, there were `r MSgaY` records where GA~3~ had been applied, and `r MSscarY` records where seeds had been scarified.
-The full germination dataset is available at GitHub (see Data Availability Statement).
+There were `r MSgaY` records where GA~3~ had been applied, and `r MSscarY` records where seeds had been scarified.
+
+Finally, to compare the germination of mesic meadow species with that of other herbaceous communities, we retrieved data from previous works on the seed germination ecology of bogs and fens [@RN3480; @RN3212], alpine and subalpine grasslands [@RN2371], and coastal plant communities of rocky cliffs and sand dunes [@RN2865].
+These additional germination records had been obtained using the same experimental methodology as the one employed for some of the germination experiments of this study: recently collected seeds, untreated for physiological dormancy (but scarified in the cases of families known for having physical dormancy), had been subjected to three germination thermoperiods (14/4, 22/12, 30/20 ºC).
+All seeds had been collected in the Cantabrian Mountains of Spain or the neighbouring coast.
+We combined these records with the records with matching experimental conditions from the meadows germination dataset to create a Cantabrian dataset.
 
@@
 
-To test the effect of germination treatments on seed germination proportions, we performed a meta-analysis [@RN5000] of the germination dataset by fitting binomial generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) [@RN5036; @RN4965] using the R package *MCMCglmm* [@RN4755].
+### Seed responses to germination cues
+
+We used the European germination dataset (available at GitHub, see Data Availability Statement) to test the effect of germination treatments on seed germination proportions.
+The used dataset included 2,005 germination records of 90 species; of these, 222 records came from our own experimental data and the rest from the databases.
+We performed a meta-analysis [@RN5000] of the germination dataset by fitting binomial generalized mixed models with Bayesian estimation (Markov Chain Monte Carlo generalized linear mixed models, MCMCglmms) [@RN4965] using the R package *MCMCglmm* [@RN4755].
 We fitted four models: (1) to the entire dataset; and separately for each of the three botanical groups of mesic meadows: (2) Poaceae, (3) Fabaceae, and (4) a third group including the rest of the families.
@@
 Random effects included the phylogenetic tree, species identity, seed lot and source of the data.
-In all models, variables were scaled so their contribution to the effect sizes could be compared.
+In all models, fixed effect variables were scaled so their contribution to the effect sizes could be compared.
 We used weakly informative priors in all models, with parameter-expanded priors for the random effects.
@@
 
-To check whether seed traits and plant ecological preferences were related, we did a Principal Component Analysis (PCA) of seed traits and species SNCs for cold, drought and pH.
+### Seed traits vs. species environmental preferences
+
+We used the Iberian dataset (available at GitHub, see Data Availability Statement) to check whether seed traits and plant ecological preferences were related.
+The used dataset contained 76 species with complete trait data for seed germination, seed morphology and ecological requirements.
+We did a Principal Component Analysis (PCA) of seed traits and species SNCs for cold, drought and pH.
 We performed the PCA ordination at the species level, i.e.
@@
 
-Finally, to compare the germination of mesic meadow species with that of other herbaceous communities, we retrieved data from previous works on the seed germination ecology of bogs and fens [@RN3480; @RN3212], alpine and subalpine grasslands [@RN2371], and coastal plant communities of rocky cliffs and sand dunes [@RN2865].
-These additional germination records had been obtained using the same experimental methodology as the one employed for some of the germination experiments of this study: recently collected seeds, untreated for physiological dormancy (but scarified in the cases of families known for having physical dormancy), had been subjected to three germination thermoperiods (14/4, 22/12, 30/20 ºC).
-All seeds had been collected in the Cantabrian Mountains of Spain or the neighbouring coast.
-We combined these records with the records with matching experimental conditions from the meadows germination dataset, and performed a PCA of the resulting dataset.
-The germination records for the other plant communities are available at GitHub (see Data Availability Statement).
+### Germination in meadows vs. other habitats
+
+We used the Cantabrian dataset (available at GitHub, see Data Availability Statement) to compare the germination of mesic meadows with that of other herbaceous communities from the same region.
+The Cantabrian dataset included 131 plant species.
+We analysed this dataset by PCA.
 
@@
 
-When considering the full set of mesic meadow species, all six studied germination cues had a significant effect on final germination proportions (**Fig.
+When considering the full set of mesic meadow species, all the studied germination cues except for stratification had a significant effect on final germination proportions (**Fig.
 3**).
-The germination of meadow seeds was positively associated with scarification, stratification, GA~3~, alternating temperatures, and light.
+The germination of meadow seeds was positively associated with scarification, GA~3~, alternating temperatures, and light.
 Average temperature had a negative effect, indicating a trend towards higher germination at lower temperatures.
@@
 In Fabaceae (**Fig.
-3**), the largest positive effect on germination was produced by scarification, with no effect of stratification, GA~3~ or light.
-Temperature had a negative effect, and in fact Fabaceae species had higher germination proportion at temperatures under 20 ºC.
-Fabaceae seed germination also responded negatively to alternating temperatures.
+3**), the largest positive effect on germination was produced by scarification, with no effect of stratification, GA~3~, alternating temperatures or light.
+Temperature had a negative effect, and in fact Fabaceae species had higher germination proportions at temperatures under 20 ºC.
 Finally, in the remaining families (**Fig.
-3**), the main cues having a positive effect on germination were stratification, GA~3~, alternating temperatures and light.
-These species did not respond to scarification nor to average temperature.
+3**), the main cues having a positive effect on germination were GA~3~, alternating temperatures and light.
+These species did not respond to scarification, stratification or average temperatures.
 
@@
 The second axis explained 17% of the variation and separated (iii, bottom) species that responded more to the cool germination temperature from (iv, top) species that responded more to the warm germination treatment.
-Meadow species tended to be situated in the right sector of the ordination, indicating high germinability; and at the center of the vertical axis, indicating a neutral response to temperature.
+Meadow species tended to be situated in the right sector of the ordination, indicating high germinability; and at the centre of the vertical axis, indicating a neutral response to temperature.
 On the other hand, species from the other communities were positioned at the left of the horizontal axis, indicating a lower germinability; and were more separated along the vertical axis, indicating a preference for either warmer (bogs and fens) or cooler (alpine grasslands, coastal communities) germination treatments.
@@
 
-Our meta-analysis of germination records showed that, overall, the seed ecology of mesic meadows is characterized by (1) high seed germinability when compared to other herbaceous plant communities; (2) low correspondence between seed trait variability and the natural environmental drivers of mesic meadow diversity; and (3) a deep phylogenetic separation between the two major families, Poaceae and Fabaceae.
+Our meta-analysis of germination records showed that the seed ecology of mesic meadows is characterized by (1) high germinability when compared to other herbaceous plant communities; (2) low correspondence between seed trait variability and the natural environmental drivers of mesic meadow diversity; and (3) a deep phylogenetic separation between the two major families, Poaceae and Fabaceae.
 
-We found that alternating temperatures and light are the two most influential factors triggering germination, especially in species from the family Poaceae and the minor families, with the exception of Fabaceae.
+We found that alternating temperatures and light are the two most influential factors triggering germination in all taxonomic groups except for Fabaceae.
 Alternating temperatures and light are usually considered to be micro-environmental cues that indicate, at a fine scale, the existence of safe sites for regeneration [@RN4719].
-The diurnal alternation of temperature decreases with burial depth in the soil, and the alternation is also reduced by vegetation cover [@RN1380; @RN4910; @RN4952].
+The diurnal alternation of temperature decreases with burial depth in the soil, and by the presence of vegetation cover [@RN1380; @RN4910; @RN4952].
 Thus, a positive germination response to alternating temperatures could detect the depth of seed burial and prevent the germination of seeds that are buried too deep for seedlings to survive before emerging from the ground [@RN4898].
-Perhaps more importantly in mesic meadows, alternating temperatures could also indicate that the vegetation cover has been diminished by either mowing or grazing, and therefore mark an appropriate time for seedling regeneration, when the competition by the established plants is lessened.
+Perhaps more importantly in mesic meadows, alternating temperatures could also indicate that the vegetation cover has been diminished by either mowing or grazing, marking an appropriate time for seedling regeneration, when the competition by the established plants is lessened.
 Similar functions can be ascribed to the germination response to light, as physiologically-active light in significant amounts only travels the uppermost millimetres of soil [@RN4909] and the quality of light will be affected by vegetation cover [@RN604].
-Therefore, the germination response to alternating temperatures and light indicates conditions that are found immediately after mowing for haymaking, a predictable perturbation that occurs approximately at the same time every year, and which forces meadow plants to establish, grow and reproduce in the time frame determined by this perturbation [@RN4887; @RN4888].
-Theoretically, these germination traits would also promote the formation of a transient soil seed bank [@RN4929; @RN1292], but it is worth mentioning that Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate meadows [@RN5014; @RN5013].
+Therefore, the germination response to alternating temperatures and light indicates conditions that are found immediately after mowing for haymaking, a predictable perturbation that occurs approximately at the same time every year, and which forces meadow plants to establish, grow and reproduce in a regular time frame [@RN4887; @RN4888].
+Theoretically, these germination traits would also promote the formation of a transient soil seed bank [@RN4929; @RN1292].
+But it is worth mentioning that Poaceae, which show the germination response to alternating temperatures and light, also have the more elongated seed shape, a trait that is thought to reduce the capacity of seeds to enter the soil seed bank in temperate meadows [@RN5014; @RN5013].
 
@@
 In the case of this study, Fabaceae seeds are amongst the heaviest, and they clearly follow a separate regeneration strategy from that of the other families.
-While other groups rely on alternating temperatures and light as gap-detecting mechanisms, Fabaceae seeds do not respond to these factors, and they are regulated instead by scarification, while they also respond to cooler temperatures than the rest of the families.
+While other groups rely on alternating temperatures and light as gap-detecting mechanisms, Fabaceae seeds are regulated instead by scarification, while they also respond to cooler temperatures than the rest of the families.
 This suggests that Fabaceae seeds would tend to germinate when temperatures are cooler, and thus before or after the summer haymaking season.
 Their larger size, and the related larger reserves, could allow Fabaceae seedlings to emerge from greater depths [@RN4898], e.g.
-when the meadow grass is still high, before mowing.
+when the meadow grass is still high before mowing.
 Furthermore, larger seed size also increases seed survival in cattle dung [@RN5050] and thus improves the capacity of being dispersed by animal depositions [@RN5051].
 
-The germination response to average temperatures and stratification is understood to detect cues related to seasonal cycling and macroclimatic variation [@RN3063; @RN5036].
-Apart from the response to cool temperatures in Fabaceae, the rest of the families did not show a response to average temperatures, further highlighting their reliance on micro-habitat cueing.
+Seed germination responses to average temperatures and stratification are understood to be a mechanism to track seasonal climatic cycles [@RN3063].
+Apart from the response to cool temperatures in Fabaceae, the rest of the families did not show a response to average temperatures, further highlighting their reliance on micro-habitat and short-term cueing.
 Moreover, the dominant family Poaceae showed a negative response to cold stratification.
-The less frequent families, however, showed a divergence from the Poaceae strategy in their positive response to stratification.
-These non-dominant species appear to require a period of cold stratification, indicating that overwintering has occurred, before they can germinate [@RN3214].
-These minor families are also the only group of species that showed a positive response to GA~3~, a phytohormone which can work as a substitute of cold stratification to overcome physiological dormancy [@RN3368].
+The less frequent families, however, showed a divergence from the Poaceae strategy: their positive response to GA~3~.
+GA~3~ is a phytohormone which can work as a substitute of cold stratification to overcome physiological dormancy [@RN3368].
+The response to GA~3~ indicates that the non-dominant species have a degree of physiological dormancy, which could postpone their germination until overwintering has occurred [@RN3214].
 
-The strong phylogenetic signal in germination responses highlights the phylogenetic clustering of regeneration strategies that we have described so far: (1) the dominant family Poaceae, showing a lack of response to seasonal cues (average temperature and cold stratification), and relying instead on large amounts of propagules and on detecting micro-niche cues (alternating temperatures, light) that can be associated to the yearly perturbation of mowing, (2) the second most-dominant family, Fabaceae, which do not respond to the micro-niche cues, and regulate their germination timing through scarification and cooler germination temperatures, possibly because their larger size and reserves allow their seeds to decouple their emergence timing from the mowing disturbance, and finally (3) the minor families that respond to micro-cues in a similar way than Poaceae, but which are differentiated from them by showing a positive response to cold stratification and G~3~, indicating that they rely on physiological seed dormancy to track seasonal cycles.
+The strong phylogenetic signal in germination responses highlights the phylogenetic clustering of regeneration strategies that we have described so far: (1) The dominant family Poaceae shows a lack of response to seasonal cues (average temperature and cold stratification).
+Instead, Poaceae rely on large amounts of propagules and on detecting micro-niche cues (alternating temperatures, light) that can be associated to the yearly perturbation of mowing.
+(2) The second most-dominant family, Fabaceae, does not respond to micro-niche cues, and regulates germination timing through scarification and cooler germination temperatures.
+Possibly, because their larger size and reserves allow their seeds to decouple their emergence timing from the mowing disturbance.
+(3) The minor families respond to micro-cues in a similar way than Poaceae, but are differentiated from them by showing a positive response to GA~3~, indicating that they rely on physiological seed dormancy to track seasonal cycles.
 This third strategy would allow to fine-tune their germination to the micro-environmental conditions plus the inter-annual climatic variation.
@@
 
-As we have seen, the regeneration strategy of high germinability appears to be general in meadow species, unrelated to regional environmental gradients, and to be more responsive to the predictable perturbation that is yearly mowing.
+As we have seen, the regeneration strategy of high germinability appears to be general in meadow species, across regional environmental gradients.
+Seed traits usually show variation along environmental gradients; some examples include warm-cued germination in species from colder sites [@RN2365] or smaller seeds in species from higher elevations [@RN4912].
+Such relationships are missing in the hay meadows of the Iberian Peninsula, highlighting the semi-natural character of this vegetation, conditioned by human management as much as by environmental constraints.
+To some extent, the lower germinability that we found in coastal habitats, mires and alpine grasslands could be a result of the stressors imposed by salinity, waterlogging and temperature extremes.
+At the same time, the seed ecology of meadow species could be a response to the potential selective force that is the predictable perturbation by yearly mowing.
+Past research supports the relationship between mowing and germination timing.
 In an abandoned meadow of the Swiss Prealps, resuming mowing promoted germination and emergence, although natural regeneration was limited due to a lack of seed dispersal to the site [@RN3073].
 Mowing also promoted autumn germination in a dry grassland of northern Germany [@RN5025].
-These results agree with the lack of dormancy and the high germinability of meadows seeds that we have found, and which has also been highlighted by previous studies.
-For example, when comparing populations of *Poa trivialis*, grassland seeds were less dormant than seeds collected in an arable field [@RN1229].
+Past studies also agree with the lack of dormancy and the high germinability of meadow seeds that we have found.
+For example, meadow seeds of *Poa trivialis* were less dormant than seeds of the same species collected in an arable field [@RN1229].
 Ten Brink et al.
-[-@RN3570], when comparing congeneric herbaceous species from open and forest habitats, found that species from open habitats needed less cold stratification.
+[-@RN3570], when comparing congeneric herbaceous species from open grasslands or forest habitats, found that species from the open grasslands needed less cold stratification.
 Similarly, seeds collected from a hay meadow germinated better when left untreated, rather than when being exposed to dormancy-breaking treatments [@RN4894].
-These high-germinability strategy has the practical consequence of greatly limiting the long-term seed bank persistence of mesic meadow species.
-Rather, the soil seed bank has been reported as being transient in hay meadows and related grasslands [@RN5026; @RN5041; @RN5042].
+The high-germinability syndrome as a response to artificial selection by mowing is indeed an intriguing hypothesis, and future work should be designed to test it.
+
+The high-germinability strategy, however, has the practical consequence of greatly limiting the long-term seed bank persistence of mesic meadow species: the soil seed bank has been reported as being transient in hay meadows and related grasslands [@RN5026; @RN5041; @RN5042].
 In Poland, the soil bank of a hay meadow was dominated by arable and weedy forbs, with low representation of Poaceae and Fabaceae [@RN5032].
+The inability to form persistent seed banks leads to dispersal limitations to passive natural regeneration [@RN3073], making mesic meadows quite sensitive to land use change, and stressing the importance of active actions of meadow restoration via seed supply.
 
-Overall, our results suggest that mesic meadows are quite sensitive to land use change.
-Their inability to form persistent seed banks leads to dispersal limitations to passive natural regeneration [@RN3073], highlighting the importance of active actions of meadow restoration via seed supply.
-For centuries, human activities have shaped the regeneration of hay meadows, leading to a loss of seed dormancy and decoupling seeds from seasonal cycles, as has been found in many domesticated species [@RN4315].
-The same anthropic processes that have shaped semi-natural mesic meadows have left them dependent on continued human intervention for their regeneration.
 Understanding the germination requirements of the different plant groups that coexist in mesic meadows can help to manage, conserve and restore their biological diversity.
+For centuries, human activities have shaped the regeneration of hay meadows, and apparently this has led to a loss of seed dormancy, decoupling seeds from seasonal cycles, as has been found in many domesticated species [@RN4315].
+The same anthropic processes that have shaped semi-natural mesic meadows may have left them dependent on continued human intervention for their regeneration.
+The high germinability of meadow seeds makes them relatively easy to use in restoration projects.
+But it also creates a fascinating dilemma to the restoration practitioner: instead of sourcing seed from existing – and seemingly domesticated – meadows, it could be advisable to also source seeds from related wild populations of meadow species, which may hold the genetic variability and phenotypic plasticity to cope with the threats posed by new environmental challenges [@RN3766].
 
@@
 ```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Higher germinability in mesic meadows compared to other herbaceous plant communities.
-Principal Component Analysis ordination of species responses to three seed germination treatments.
+Principal Component Analysis ordination of species responses to three seed germination treatments (fresh seeds germinated at 30/20, 22/12 and 14/4 ºC in a 12/12h light-darkness photoperiod).
 Species are grouped by their habitat, with points showing the centroid for each habitat group, and rays linking the centroid with the position of each species in the group.

### 2021-10-14 src/cleaning.R

@@ -1,5 +1,12 @@
 library(tidyverse); library(openxlsx)
 
+# Read meadow species list (https://doi.org/10.1111/avsc.12519)
+
+read.csv("data/species.csv") %>%
+  separate(Taxon, into = c("Genus", "Species"), sep = " ") %>%
+  mutate(Taxon = paste(Genus, Species), sep = " ") %>%
+  select(Taxon) -> spplist
+
 # Clean species data
 
 read.xlsx("../#data/plots/data/sospraderas/Inventarios SOS PRADERAS privacidad.xlsx", sheet = 1) %>% # These are the relevés in the DB
@@ -133,93 +140,6 @@ read.csv("../#data/germination/data/sospraderas/Muestras de semillas.csv") %>% #
                 Country) ->
   seedlots
 
-# Clean morphometric analysis data
-
-read.csv("../#data/germination/data/sospraderas/Accesiones morfo.csv") %>% # Seedlot information for the morphological analysis
-  filter(Tipo.de.espécies == "Espécies silvestres") %>% # Commercial seed lots of the project, are of no interest here
-  dplyr::select(ID, Label) %>% 
-  merge(read.csv("../#data/germination/data/sospraderas/Análise morfométrica.csv"), by = "Label") %>% # Results of the morphometric analysis as provided by Madalena Vaz
-  dplyr::select(-c(Label, X.1)) %>%
-  rename(Seedlot = ID) %>% # This is the seedlot ID code as used in the final report of the project
-  merge(seedlots, by = "Seedlot") %>%
-  rename(Taxon = Species) %>%
-  group_by(Taxon) %>%
-  select(Taxon, Country, Seedlot, Height, Width) %>%
-  rename(Population = Seedlot, Length = Height) %>%
-  filter(Width < 2*Length) -> morpho
-
-# Clean seed mass
-
-read.csv("../#data/germination/data/enscobase meadows/Species_domimance_in_ENSCOBASE.csv") %>%
-  mutate(Taxon = gsub("_", " ", animal)) %>%
-  select(Taxon, mass) %>% 
-  rename(Seed.mass = mass) %>%
-  unique -> mass1
-
-read.csv("../#data/germination/data/sospraderas/SID Seed Mass.csv") %>%
-  rename(Taxon = Species,
-         Seed.mass = Seed.mass.mg) -> mass2
-
-read.csv("../#data/seedmass/results/seedmass.csv") %>%
[…]

### 2021-10-14 src/figures.R

@@ -4,7 +4,7 @@ library(tidyverse)
 
 ## Plot figure A
 
-read.csv("data/traits.csv") %>%
+read.csv("data/iberian.csv") %>%
   select(Taxon, Family,
          Seed.mass, Seed.number) %>%
   mutate(Family = ifelse(Family %in% c("Leguminosae", "Poaceae"), Family, "Others"),
@@ -32,18 +32,14 @@ read.csv("data/traits.csv") %>%
 
 ## Plot figure B
 
-read.csv("data/traits.csv") %>%
-  select(Taxon, Family) %>%
-  merge(read.csv("data/morphometrics.csv"), by = "Taxon") %>%
+read.csv("data/morphometrics.csv") %>%
   select(Taxon, Family) %>%
   unique %>%
   group_by(Family) %>%
   tally() %>%
   pull(Family) -> fams
 
-read.csv("data/traits.csv") %>%
-  select(Taxon, Family) %>%
-  merge(read.csv("data/morphometrics.csv"), by = "Taxon") %>% 
+read.csv("data/morphometrics.csv") %>% 
   filter(Family %in% fams) %>%
   mutate(Family = ifelse(Family %in% "Compositae", "Asteraceae", Family),
          Family = ifelse(Family %in% "Leguminosae", "Fabaceae", Family),
@@ -167,21 +163,10 @@ ggsave(f3, file = "results/figures/Fig3.png",
 
 ## Prepare data frame
  
-read.csv("data/germination.csv") %>%
-  mutate(Germination = Germinated / Germinable) %>%
-  group_by(Taxon) %>%
-  summarise(Temperature = weighted.mean(Tmean, w = Germination),
-            Alternating = weighted.mean(Alternating, w = Germination),
-            Light = weighted.mean(Light, w = Germination),
-            Scarification = weighted.mean(Scarification, w = Germination),
-            Stratification = weighted.mean(Stratification, w = Germination),
-            GA3 = weighted.mean(GA3, w = Germination)) %>%
-  merge(read.csv("data/traits.csv"), by = "Taxon") %>%
+read.csv("data/iberian.csv") %>%
   mutate(Family = ifelse(Family %in% c("Leguminosae", "Poaceae"), Family, "Others"),
[…]

### 2021-10-14 src/models.R

@@ -6,6 +6,8 @@ phangorn::nnls.tree(cophenetic(ape::read.tree("data/meadowstree.tree")),
                     ape::read.tree("data/meadowstree.tree"), rooted = TRUE) -> 
   nnls_orig
 
+nnls_orig$node.label <- NULL
+
 # Set number of iterations
 
 nite = 500000
@@ -23,8 +25,7 @@ list(R = list(V = 1, nu = 50),
 
 # Read data
 
-read.csv("data/germination.csv") %>%
-  merge(read.csv("data/traits.csv"), by = "Taxon") %>%
+read.csv("data/european.csv") %>%
   mutate(ID = gsub(" ", "_", Taxon), animal = ID) %>% 
   select(ID, animal, Family,
          Reference:Population, Scarification:GA3,

### 2021-10-14 src/numbers.R

@@ -1,7 +1,7 @@
 # Read data
 
-read.csv("../data/germination.csv") -> germination
-read.csv("../data/traits.csv") -> traits
+read.csv("../data/european.csv") -> germination
+read.csv("../data/iberian.csv") -> traits
 load("../results/models/allspp.RData")
 
 # Calculate numbers
@@ -47,16 +47,7 @@ summary(allspp)$Gcovariances[2, 3] %>% round(2) -> MSsourceh
 
 # PCA numbers
 
-read.csv("../data/germination.csv") %>%
-  mutate(Germination = Germinated / Germinable) %>%
-  group_by(Taxon) %>%
-  summarise(Temperature = weighted.mean(Tmean, w = Germination),
-            Alternating = weighted.mean(Alternating, w = Germination),
-            Light = weighted.mean(Light, w = Germination),
-            Scarification = weighted.mean(Scarification, w = Germination),
-            Stratification = weighted.mean(Stratification, w = Germination),
-            GA3 = weighted.mean(GA3, w = Germination)) %>%
-  merge(read.csv("../data/traits.csv"), by = "Taxon") %>%
+read.csv("../data/iberian.csv") %>%
   mutate(Family = ifelse(Family %in% c("Leguminosae", "Poaceae"), Family, "Others"),
          Family = ifelse(Family %in% "Leguminosae", "Fabaceae", Family)) %>%
   mutate(Family = fct_relevel(Family, c("Poaceae", "Fabaceae", "Others"))) %>%

### 2021-10-20 doc/manuscript.Rmd
AoB 2nd revision 🍷

@@
 The mesic meadows of Europe are semi-natural open habitats that occupy clearings created by human intervention over a temperate wooded landscape [@RN5033; @RN5045], in sites with moderately fertile and well-drained soils [@RN4698].
-Once, these were novel plant communities, with a composite flora made up of species from different ecological backgrounds.
-During historical times, their species composition further evolved following changes in human society and management practices [@RN5035; @RN5045].
-Mesic meadows are thus the result of a process of habitat domestication that is characteristic of Europe's natural history [@RN5034].
-Nowadays, mesic meadows maintained by agricultural practices are an essential aspect of European landscapes, both real [@RN4996] and fictional [@RN5029].
+These novel anthropogenic plant communities, with a composite flora made up of species from different ecological backgrounds, have constantly changed in response to shifts in human society and associated management practices [@RN5035; @RN5045], but nonetheless characterise Europe's natural history [@RN4996].
 

### 2021-10-22 doc/manuscript.Rmd
ms accepted ☘️

@@
 ---
-title: "RESEARCH IN CONTEXT: Seed ecology of European mesic meadows"
-output: 
-  bookdown::word_document2:
-    reference_docx: template.docx
-    number_sections: false
+title: "Seed ecology of European mesic meadows"
+output: github_document
 bibliography: refs.bib
@@
 
-Running title: Seed ecology of European mesic meadows
-
 ^+^ Correspondence: IMIB – Biodiversity Research Institute, University of Oviedo, Edificio de Investigación – 5ª planta, C/ Gonzalo Gutiérrez Quirós s/n
@@
 
-# ABSTRACT (300 words)
+# ABSTRACT
 
@@
 Wild Poaceae ecotypes usually have a germination response to alternating temperatures [@RN1248; @RN1261; @RN1230; @RN1221; @RN1290; @RN1172; @RN1292; @RN1280], although this response is missing in some wild ecotypes and in the domesticated cultivars [@RN1261], and some species such as *Lolium perenne* have been reported as insensitive to temperature alternation [@RN1280; @RN1380].
+
+```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 1** Diversity of dispersal units in plant species from mesic meadows: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadensis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (o) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
+knitr::include_graphics("../results/figures/Fig1.png")
+```
 
@@
 
+```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 2** Morphology of mesic meadow seeds.
+The two panels on the left show the probability densities of species values for seed mass and seed number, log-transformed for ease of visualization.
+The three horizontal lines within the probability densities represent the first quartile, the median and the third quartile of the values.
+The panel on the right shows values of seed length and width obtained by image analysis, with each point being a seed.
+In all cases, data is divided between the grasses (Poaceae), legumes (Fabaceae) and the other plant families."}
+knitr::include_graphics("../results/figures/Fig2.png")
+```
+
 ## Seed germination
@@
 
+```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 3** Germination cues in mesic meadows.
+Effect of germination cues simulated in the laboratory over the final germination proportions of mesic meadow seeds.
+Dots indicate the posterior mean of the effect size for each cue, and whiskers the 95% credible interval of the effect size.
+The line of zero-effect is shown: when a credible interval overlaps with the zero-effect line, the effect can be regarded as non-significant.
+In separate panels, the figure shows the results of a general model including data for all species, plus specific models for the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families."}
+knitr::include_graphics("../results/figures/Fig3.png")
+```
+
 To describe the effects of the random factors, we will refer only to the full model that included all species in the dataset.
@@
 
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 4** Environment and seed traits are separate axis of variation in mesic meadows.
+Principal Component Analysis ordination of mesic meadow species considering their environmental preferences and their seed traits.
+Each point is a species, coloured by the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families.
+Labels indicate the contribution of the variables to the axes: grey-background labels for environmental preferences, and white-background labels for seed traits.
+Environmental preferences were calculated as species niche centroids (SNCs) for the minimal temperature of the coldest month (bio06), precipitation of the driest month (bio14), and soil pH.
+To calculate the SNCs, a vegetation dataset of mesic meadows of the Iberian Peninsula was used.
+The seed traits are seed mass and the germination relative indices for the response to scarification, stratification, average germination temperature (temperature), alternating temperature (alternating) and light.
+All environmental preferences aligned to the first axis, while seed traits aligned to the second axis, showing that environment and seed traits are separate axis of variation in mesic meadows.
+Stratification and average germination temperature showed very low variation, in accordance with their small effect on Figure 2."}
+knitr::include_graphics("../results/figures/Fig4.png")
+```
+
 ## Comparison with other habitats
@@
 On the other hand, species from the other communities were positioned at the left of the horizontal axis, indicating a lower germinability; and were more separated along the vertical axis, indicating a preference for either warmer (bogs and fens) or cooler (alpine grasslands, coastal communities) germination treatments.
+
+```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "**Figure 5** Higher germinability in mesic meadows compared to other herbaceous plant communities.
+Principal Component Analysis ordination of species responses to three seed germination treatments (fresh seeds germinated at 30/20, 22/12 and 14/4 ºC in a 12/12h light-darkness photoperiod).
+Species are grouped by their habitat, with points showing the centroid for each habitat group, and rays linking the centroid with the position of each species in the group.
+Labels indicate the contribution of the germination treatments to the axes.
+All species had been collected in herbaceous vegetation types of the Cantabrian Mountains of Spain and the neighbouring coast.
+All seeds were untreated, except for scarification, which was applied routinely to all botanical families presenting physical dormancy.
+Germination treatments consisted in 12/12 h periods with a warmer phase in light and a cooler phase in darkness.
+The position of mesic meadows compared to the other habitats indicates that their species tended to have higher germinability even if untreated (i.e.
+less seed dormancy) and were more neutral in their thermal requirements."}
+knitr::include_graphics("../results/figures/Fig5.png")
+```
 
@@
 The original data, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/meadows.
-Upon publication, a version of record of the repository will be deposited in Zenodo.
+A version of record of the repository is deposited in Zenodo.
 
@@
 :::
-
-# FIGURES
-
-```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Diversity of dispersal units in plant species from mesic meadows: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadensis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (o) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
-knitr::include_graphics("../results/figures/Fig1.png")
-```
-
-```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Morphology of mesic meadow seeds.
-The two panels on the left show the probability densities of species values for seed mass and seed number, log-transformed for ease of visualization.
-The three horizontal lines within the probability densities represent the first quartile, the median and the third quartile of the values.
-The panel on the right shows values of seed length and width obtained by image analysis, with each point being a seed.
-In all cases, data is divided between the grasses (Poaceae), legumes (Fabaceae) and the other plant families."}
-knitr::include_graphics("../results/figures/Fig2.png")
-```
-
-```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Germination cues in mesic meadows.
-Effect of germination cues simulated in the laboratory over the final germination proportions of mesic meadow seeds.
-Dots indicate the posterior mean of the effect size for each cue, and whiskers the 95% credible interval of the effect size.
-The line of zero-effect is shown: when a credible interval overlaps with the zero-effect line, the effect can be regarded as non-significant.
-In separate panels, the figure shows the results of a general model including data for all species, plus specific models for the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families."}
-knitr::include_graphics("../results/figures/Fig3.png")
-```
-
-```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Environment and seed traits are separate axis of variation in mesic meadows.
-Principal Component Analysis ordination of mesic meadow species considering their environmental preferences and their seed traits.
-Each point is a species, coloured by the three main botanical groups of mesic meadows: grasses (Poaceae), legumes (Fabaceae) and the other families.
-Labels indicate the contribution of the variables to the axes: grey-background labels for environmental preferences, and white-background labels for seed traits.
-Environmental preferences were calculated as species niche centroids (SNCs) for the minimal temperature of the coldest month (bio06), precipitation of the driest month (bio14), and soil pH.
-To calculate the SNCs, a vegetation dataset of mesic meadows of the Iberian Peninsula was used.
-The seed traits are seed mass and the germination relative indices for the response to scarification, stratification, average germination temperature (temperature), alternating temperature (alternating) and light.
-All environmental preferences aligned to the first axis, while seed traits aligned to the second axis, showing that environment and seed traits are separate axis of variation in mesic meadows.
-Stratification and average germination temperature showed very low variation, in accordance with their small effect on Figure 2."}
-knitr::include_graphics("../results/figures/Fig4.png")
-```
-
-```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Higher germinability in mesic meadows compared to other herbaceous plant communities.
-Principal Component Analysis ordination of species responses to three seed germination treatments (fresh seeds germinated at 30/20, 22/12 and 14/4 ºC in a 12/12h light-darkness photoperiod).
-Species are grouped by their habitat, with points showing the centroid for each habitat group, and rays linking the centroid with the position of each species in the group.
-Labels indicate the contribution of the germination treatments to the axes.
-All species had been collected in herbaceous vegetation types of the Cantabrian Mountains of Spain and the neighbouring coast.
-All seeds were untreated, except for scarification, which was applied routinely to all botanical families presenting physical dormancy.
-Germination treatments consisted in 12/12 h periods with a warmer phase in light and a cooler phase in darkness.
-The position of mesic meadows compared to the other habitats indicates that their species tended to have higher germinability even if untreated (i.e.
-less seed dormancy) and were more neutral in their thermal requirements."}
-knitr::include_graphics("../results/figures/Fig5.png")
-```

### 2021-10-22 README.Rmd
readme update ☘️

--- 2021-04-30 README.Rmd
@@
 
-# Seed germination ecology of European mesic grasslands
+# Seed ecology of European mesic grasslands
 
-This repository stores all the information related to the manuscript *Seed germination ecology of European mesic grasslands*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
+This repository stores all the information related to the manuscript *Seed ecology of European mesic grasslands*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
 
@@
 
-Please cite the dataset as `CITATION`
+Please cite the dataset as `Fernández-Pascual E, Vaz M, Morais B, Reiné R, Ascaso J, Afif Khouri E, Carta A. Seed germination ecology of European mesic grasslands. Annals of Botany, in press.`

### 2021-10-23 README.Rmd
updated readme ☘️

--- 2021-10-14 README.Rmd
@@
 
-This repository stores all the information related to the manuscript *Seed ecology of European mesic meadows*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
+This repository stores all the information related to the manuscript [*Seed ecology of European mesic meadows*](https://dx.doi.org/10.1093/aob/mcab135), including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
 
@@
 
-* `data` Data files including the seed trait datasets (European, Iberian and Cantabrian) and the phylogenetic tree.
-* `doc` Files to create the manuscript using `Rmarkdown`.
+* `data` Data files including the [European](https://github.com/efernandezpascual/meadows/blob/master/data/european.csv), [Iberian](https://github.com/efernandezpascual/meadows/blob/master/data/iberian.csv) and [Cantabrian](https://github.com/efernandezpascual/meadows/blob/master/data/cantabrian.csv) datasets; the [seed morphometrics](https://github.com/efernandezpascual/meadows/blob/master/data/morphometrics.csv) and the [phylogenetic tree](https://github.com/efernandezpascual/meadows/blob/master/data/meadowstree.tree).
+* `doc` Files to create the [manuscript](https://github.com/efernandezpascual/meadows/blob/master/doc/manuscript.md) using `Rmarkdown`.
 * `results` Output of the `R` scripts, including figures and models.
@@
 
+## Abstract
+
+* **Background and Aims** European mesic meadows are semi-natural open habitats of high biodiversity and an essential part of European landscapes.
+These species-rich communities can be a source of seed mixes for ecological restoration, urban greening and rewilding.
+However, limited knowledge of species germination traits is a bottleneck to the development of a competitive native seed industry.
+Here, we synthesize the seed ecology of mesic meadows.
+* **Methods** We combined our own experimental data with data obtained from databases to create a combined dataset containing 2,005 germination records of 90 plant species from 31 European countries.
+We performed a Bayesian meta-analysis of this dataset to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light.
+We also used multivariate ordination to check the relationship between seed traits (germination and morphology) and species ecological preferences, and to compare the seed ecology of mesic meadows with that of other herbaceous plant communities from the same area.
+* **Key Results** The seed ecology of mesic meadows is characterized by (1) high seed germinability when compared to other herbaceous plant communities; (2) low correspondence between seed traits and species ecological preferences; and (3) a deep phylogenetic separation between the two major families, Poaceae and Fabaceae.
+Poaceae produce many light seeds which respond to gap-detecting germination cues (alternating temperatures and light); Fabaceae produce fewer heavy seeds, which need scarification to break their physical dormancy.
+* **Conclusions** High germinability of meadow seeds will reduce their capacity to form persistent seed banks, resulting in dispersal limitations to passive regeneration.
+For centuries, human activities have shaped the regeneration of meadows, leading to a loss of seed dormancy and decoupling seeds from seasonal cycles, as has been found in many domesticated species.
+The same anthropic processes that have shaped semi-natural mesic meadows have left them dependent on continued human intervention for their regeneration, highlighting the importance of active restoration via seed supply.
+
 ## Citation
 
-Please cite the dataset as `CITATION`
+Please cite the repository, datasets and article as:
+
+> Fernández-Pascual E, Vaz M, Morais B, Reiné R, Ascaso J, Afif Khouri E, Carta A.
+[Seed germination ecology of European mesic meadows.](https://dx.doi.org/10.1093/aob/mcab135) *Annals of Botany*, in press.

### 2022-02-14 README.Rmd
volume and pages
zenodo link

@@
 ```
+
+[![DOI](https://zenodo.org/badge/349511614.svg)](https://doi.org/10.5281/zenodo.5602741)
 
@@
 
-> Fernández-Pascual E, Vaz M, Morais B, Reiné R, Ascaso J, Afif Khouri E, Carta A.
-[Seed germination ecology of European mesic meadows.](https://dx.doi.org/10.1093/aob/mcab135) *Annals of Botany*, in press.
+> Fernández-Pascual E, Vaz M, Morais B, Reiné R, Ascaso J, Afif Khouri E, Carta A. [Seed ecology of European mesic meadows.](https://dx.doi.org/10.1093/aob/mcab135) *Annals of Botany* 129, 121-133.
